\documentclass[11pt, a4paper]{report}
\usepackage{Custom}
\pagestyle{headings}
\makeindex[intoc]

\begin{document}

\reversemarginpar
\setcounter{tocdepth}{2}

\title{Compendium of Mathematics \& Physics}
\author{Nicolas Dewolf}
\maketitle

\tableofcontents

\chapter{Introduction}

    This compendium originated out of the necessity of a compact summary of important theorems and formulas during physics and mathematics classes at university. When the interest in more (and more exotic) subjects grew, this collection lost its compactness and became the chaos it now is. Although there should exist some kind of overall structure, it was not always possible to keep every section self-contained or respect the order of the chapters.

    It should definitely not be used as a formal introduction to any subject. It is neither a complete work nor a fact-checked one, so the usefulness and correctness is not guaranteed. However it should be used as lookup table for theorems and formulas and as a guide to the literature. To this end each chapter begins with a list of useful references. At the same time only a small number of statements are proven in the text (or appendices). This was done to keep the text as short as possible. However, in some cases the major ideas underlying the proofs are given.

\section{Conventions}

    Definitions, properties and formulas marked by a dagger symbol $^\dag$ are explained and/or derived in one of the appendices. This has been done such that the summary itself contains only core notions and theorems. Sections and statements that require more advanced concepts, in particular concepts from later chapters or (higher) category theory, will be labelled by a \textit{clubs} symbol $\clubsuit$.

    Definitions of words in the middle of a text will be indicated by the use of \textbf{bold font}. Terminology that has been defined in the past but that receives a new meaning/nuance will be indicated by \dotuline{a dotted line}. Notions that have not been defined in this summary but that are relevant are indicated by \textit{italic text}. Names of authors will also be written in \textit{italic}.

    Objects from a general category will be denoted by a lower case letter (depending on the context we might also use upper case for clarity), functors will be denoted by upper case letters and the categories themselves will be denoted by symbols in \textbf{bold font}. In the later chapters on physics we will often adopt specific conventions for the different species of vectors. Vectors in Euclidean space will be denoted by a bold font letter with an arrow above, e.g. $\vector{a}$. Vectors in Minkowski space (4-vectors) and differential forms will be written without the arrow, e.g. $\mathbf{a}$. Matrices and tensors will always be represented by capital letters and dependent on the context a specific font will be adopted.

\part{Set Theory \& Algebra}
\input{Set/Set}

\part{Category Theory}
\input{Set/Cat}
\input{Set/HomAlg}
\input{Set/Topoi}
\input{Set/Type}

\part{Topology}
\input{Set/Topology}
\input{Set/MetricSpace}
\input{Set/AlgTop}
\input{Set/Sheaf}
\input{Set/Model}
\input{Set/AlgGeom}

\part{Calculus}
\input{Calculus/Calculus}
\input{Calculus/ComplexCalculus}
\input{Calculus/Lebesgue/Lebesgue}
\input{Calculus/Distributions}
\input{Calculus/OrdinaryDiffEq}
\input{Calculus/PartialDiffEq}

\part{Linear Algebra}
\input{Vector/LinearAlgebra}
\input{Vector/Tensors}
\input{Vector/Rep}
\input{Vector/Normed}
\input{Vector/Operators}
\input{Vector/Clifford}
\input{Vector/QA}
\input{Vector/HDA}

\part{Differential Geometry}\label{part:diffgeom}
\input{Diff/Surface}
\input{Diff/Manifolds}
\input{Diff/Lie}
\input{Diff/Bundles}
\input{Diff/Connections}
\input{Diff/Integration}
\input{Diff/Riemann}
\input{Diff/Symplectic}
\input{Diff/Contact}
\input{Diff/ComplexBundles}
%\input{Diff/Infinite}
\input{Diff/Var}
\input{Diff/K}
\input{Diff/Synth}

\part{Probability Theory \& Statistics}
\input{Probability}
\input{Info}
\input{Statistics}
\input{Data}
\input{Set/Fuzzy}
%\input{Stochastic}

\part{Classical Mechanics}
\input{Physics/Mechanics}
\input{Physics/Lagrange}
\input{Physics/Optics}
\input{Physics/Astronomy}

\part{Electromagnetism}
\input{Physics/Electricity}
\input{Physics/Maxwell}

\part{Relativity Theory}
\input{Physics/Relativity}

\part{Quantum Mechanics}
\input{Physics/QM/WaveMatrix}
\input{Physics/QM/Formalism}
\input{Physics/QM/AngularMomentum}
\input{Physics/QM/Dirac}
\input{Physics/QM/Perturbation}
\input{Physics/QM/Scattering}
\input{Physics/QM/Computing}

\part{Quantum Field Theory}
\input{Physics/Field}
\input{Physics/QM/KG}
\input{Physics/QM/YM}
%\input{Physics/QM/QCD}
\input{Physics/QM/CFT}
\input{Physics/QM/SUSY}
\input{Physics/QM/Entanglement}
\input{Physics/QM/AxiomaticQFT}

\part{Thermodynamics \& Statistical Mechanics}
\input{Physics/Thermodynamics}
\input{Physics/StatisticalMechanics}

\part{Condensed Matter Physics}
\input{Physics/Condensed/SolidState}
\input{Physics/Condensed/TenNet}

\chapter{Epilogue}

    Since one of the main goals of this compendium was to keep me motivated to read papers and watch lectures about the various subjects under consideration, I should also add some of the seminal works that I hope to be able to understand  in the future, or at least be a able to skim through:

    \begin{center}
        \begin{tabular}{|c|c|}
            \hline
            Author&Title\\
            \hline
            Brown-Henneaux&Central charges in the canonical realization of ...\\
			D'Auria-Fr\'e&Geometric supergravity in $D=11$ and its hidden supergroup\\
			D'Auria-Fr\'e-Castellani&Supergravity and Superstrings: A Geometric Perspective\\
            \textit{Friends}&Master's (and potentially PhD) theses\\
            Fuchs-Runkel-Schweigert&TFT construction of RCFT correlators\\
            Grothendieck&Pursuing Stacks\\
            Lurie&Higher Topos Theory (\textit{book})\\
            Lurie&On the Classification of Topological Field Theories\\
            Schreiber&Differential cohomology in a cohesive $\infty$-topos\\
            Schreiber&Higher prequantum geometry\\
            Schreiber&Local prequantum field theory\\
            \hline
        \end{tabular}
    \end{center}

\part{Appendices}
\begin{appendices}
\input{DerivMath}
\input{Diff/GStruct}
\input{Physics/deriv1}
\input{Scrib}
\end{appendices}

\nomenclature[O_zsymisisom]{$\cong$}{is isomorphic to}
\nomenclature[O_zsymisapprox]{$\approx$}{is approximately equal to}
\nomenclature[O_zsymmap]{$\mapsto$}{mapsto}
\nomenclature[O_zsymisincl]{$\hookrightarrow$}{is included in}
\nomenclature[O_zsymangle]{$\sphericalangle(v, w)$}{Angle between the vectors $v$ and $w$.}
\nomenclature[O_zsymbincart]{$X\times Y$}{Cartesian product of the sets $X$ and $Y$.}
\nomenclature[O_zsymids]{$\mathbbm{1}_X$}{Identity map on the set $X$.}
\nomenclature[O_e]{$e$}{Identity element of a group.}
\nomenclature[S_zsyminto]{$]a, b[$}{Open interval}
\nomenclature[S_zsymintc]{$[a, b]$}{Closed interval}
\nomenclature[S_zsymempty]{$\emptyset$}{Empty set}
\nomenclature[S_Sn]{$S^n$}{Standard $n$-sphere}
\nomenclature[S_Tn]{$T^n$}{Standard $n$-torus. Cartesian product of $n$ times $S^1$.}
\nomenclature[S_Dn]{$D^n$}{Standard $n$-disk}

%Print symbol list
\printnomenclature

\nocite{*}
\bibliographystyle{unsrt}
\bibliography{Biblio}

\printindex

\end{document}
