\chapter{Notes}

    This chapter contains notes written down during lunch talks, summer courses and conferences.

\section{``Noether's Theorem and Gauge-Gravity Duality'' by S. De Haro}

    Date \& location: October 6 2018, London\\
    Conference: \textit{The Philosophy and Physics of Noether's Theorems}

\subsection{Pseudotensors}

    Maxwell theory has a Noether stress-energy tensor of the form
    \begin{gather}
        T^\mu_{\ \nu} := F^{\mu\lambda}\partial A_\lambda - \frac{1}{4}\delta^\mu_\nu F^{\lambda\kappa}F_{\lambda\kappa}
    \end{gather}
    and an associated weak conservation law
    \begin{gather}
        \partial_\mu T^{\mu\nu}\approx 0.
    \end{gather}
    Through the Schwarz theorem \ref{calculus:schwarz_theorem}, this tensor can be enlarged to a conserved \textit{Belinfante tensor}
    \begin{gather}
        \overline{T}^\mu_{\ \nu} := T^\mu_{\ \nu} + \partial_\lambda U^{[\mu\lambda]}_{\ \ \ \ \nu}.
    \end{gather}
    Using such an extension, the standard form
    \begin{gather}
        T^\mu_{\ \nu} := F^{\mu\lambda}F_{\nu\lambda} - \frac{1}{4}\delta^\mu_\nu F^{\lambda\kappa}F_{\lambda\kappa} + \text{eom}
    \end{gather}
    can be obtained.

    Now, consider the theory coupled to gravity. In this case partial derivatives have to be replaced by covariant derivatives and the (weak) conservation law becomes:
    \begin{gather}
        \nabla_\mu T^{\mu\nu}.
    \end{gather}
    However, this identity only contains the stress-energy tensor of matter, not of gravity itself. A possible solution, due to \textit{Einstein}, was to construct a stress-energy (pseudo)tensor\footnote{Pseudotensor here just means an object that does not transform tensorially.} from the Christoffel symbols, since these are already responsible for the coupling to gravity in the conservation law above:
    \begin{gather}
        \partial_\mu(\sqrt{g}T^\mu_{\ \nu}+t^\mu_{\ \nu}) = 0.
    \end{gather}
    As above this leads to a \textbf{superpotential}:
    \begin{gather}
        \sqrt{g}T^\mu_{\ \nu}+t^\mu_{\ \nu} = \partial_\lambda s^{\mu\lambda}_{\ \ \ \nu},
    \end{gather}
    where $\partial_\mu\partial_\lambda s^{\mu\lambda}_{\ \ \ \nu}=0$. All conservation laws will be preserved if $s^{\mu\lambda}_{\ \ \ \nu}=s^{[\mu\lambda]}_{\ \ \ \ \nu}$.

    The main issues with this new ``stress-energy tensor'' are:
    \begin{itemize}
        \item It is not a tensor.
        \item There are an infinite number of possibilities.
    \end{itemize}
    However, the superpotential can be related to boundary conditions and, therefore, has a physical intepretation. Consider the Hamiltonian
    \begin{gather}
        \overline{H}(n) := \int_\Sigma n^\mu H_\mu + \oint_{\partial\Sigma}B(n),
    \end{gather}
    where $n$ is the ADM-like shift vector that generates tangential motion along the spacelike hypersurface $\Sigma$ and $B(n)\sim n^\nu s^{\mu\lambda}_{\ \ \nu}$. Noether's theorem implies that $H_\mu$ is proportional to the EOM, which implies that $H(n)$ is a boundary term determined by the superpotential. It determines the quasi-local energy.

    A different approach is the \textbf{Brown-York pseudotensor}. Here, a bounded spacetime region $M$ is considered with two spacelike boundaries, the initial and final slices $\Sigma_\pm$, and a timelike hypersurface $N$. The matter-coupled gravitational action is given by:
    \begin{gather}
        S = \frac{1}{2\kappa}\int_Md^4x\sqrt{-g}R + \frac{1}{\kappa}\int_{\Sigma_\pm}d^3x\sqrt{h}K - \frac{1}{\kappa}\int_Nd^3x\sqrt{-\gamma}\Theta + S_\mathrm{matter} + S_\mathrm{ref},
    \end{gather}
    where $S_\mathrm{ref}$ is a reference action to regularize the action. Variation with respect to the metric gives:
    \begin{gather}
        \delta S = -\int_Md^4xE^{\mu\nu}\delta g_{\mu\nu} + \frac{1}{2}\int_{\Sigma_\pm}d^3x\sqrt{h}P_{ij}\delta h_{ij} + \frac{1}{2}\int_Nd^3x\sqrt{\gamma}\tau^{ij}_\mathrm{BY}\delta\gamma_{ij} +\text{matter terms}.
    \end{gather}
    The first term gives the Einstein field equations and vanishes on-shell. The other terms define the conjugate momenta. As the notation implied, the object $\tau^{ij}_\mathrm{BY}$ is the (quasilocal) Brown-York stress-energy tensor. For pure gravity this can be written as follows:
    \begin{gather}
        \tau^{ij}_\mathrm{BY} = -\frac{1}{\kappa\sqrt{\gamma}}(\Theta\gamma^{ij}-\Theta^{ij}) +\text{regularizing terms}.
    \end{gather}

\subsection{AdS-CFT duality}

    Now, consider an anti-de Sitter spacetime $M$. In the CFT description the stress-energy tensor can be obtained by a functional derivative of the partition function with respect to some fixed background metric $g_{(0)}$:
    \begin{gather}
        \langle T_{ij}(x) \rangle = \frac{2}{\sqrt{g_{(0)}}}\frac{\delta W[g_{(0)}]}{\delta g^{ij}_{(0)}(x)}.
    \end{gather}
    In the gravitational description the fixed metric rerepsents the asymptotic metric (up to conformal factors). A theorem by \textit{Fefferman-Graham} says that near the boundary, the metric admits a local Poincar\'e form:
    \begin{gather}
        ds^2 = \frac{l^2}{r^2}(dr^2 + g_{ij}(r,x)dx^idx^j),
    \end{gather}
    with $g_{ij}(0,x) = g_{(0)ij}(x)$. Holographic duality then shows that the qausilocal Brown-York tensor is equal to the holographic stress-energy tensor.

\section{``A generalization of Noether's theorem and the information-theoretic approach to the study of symmetric dynamics'' by R. Spekkens}

    Date \& location: October 6 2018, London\\
    Conference: \textit{The Philosophy and Physics of Noether's Theorems}

    The general idea of this talk is that (quantum) entanglement can be considered from a \textit{resource theory} perspective. In this approach quantum entanglement is a \textbf{resource} and LOCC operations are the free operations, i.e. given sufficient entanglement they can be used to perform any quantum operation.