\chapter{Operator Algebras}\label{chapter:operator_algebras}

    The main reference for this chapter is \cite{blackadar}.

\section{\texorpdfstring{$C^*$-}{C-star }algebras}
\subsection{Involutive algebras}

    \newdef{Involutive algebra}{\index{algebra!involutive}\index{$*$-algebra|see{algebra, involutive}}
        An involutive algebra is an associative algebra $A$ over a commutative involutive ring $(R,\overline{\,\cdot\,}\,)$ together with an algebra involution $\cdot^*:A\rightarrow A$ such that:
        \begin{enumerate}
            \item $(a + b)^* = a^* + b^*$,
            \item $(ab)^* = b^*a^*$, and
            \item $(\lambda a)^* = \overline\lambda a^*$,
        \end{enumerate}
        for all $a,b\in A$ and $\lambda\in R$. These algebras are also sometimes called \textbf{$\ast$-algebras}.
    }

    \newdef{$C^*$-algebra}{\index{C$^*$-algebra}\label{operators:c_star}
        A $C^*$-algebra is an involutive Banach algebra \ref{functional:banach_space} such that the \textbf{$C^*$-identity}
        \begin{gather}
            \|a^*a\| = \|a\|\|a^*\|
        \end{gather}
        is satisfied.
    }

    The Artin-Wedderburn theorem \ref{algebra:artin_wedderburn} implies the following decomposition theorem:
    \begin{theorem}
        Let $C$ be a finite-dimensional $C^*$-algebra. There exist unique integers $N,d_1,\ldots,d_N$ such that
        \begin{gather}
            C\cong\bigoplus_{i=1}^NM_{d_i}(K).
        \end{gather}
    \end{theorem}
    This implies that every $C^*$-algebra can be represented using block matrices.

    \begin{example}[Bounded operators]
        Let $\mathcal{H}$ be a finite-dimensional Hilbert space. The space of bounded operators $\mathcal{B}(\mathcal{H})$ is a $C^*$-algebra.
    \end{example}

    \begin{property}
        Every norm-closed $\ast$-subalgebra of a $C^*$-algebra is a $C^*$-algebra.
    \end{property}

    \newdef{$H^*$-algebra}{\index{H-!$\ast$-algebra}
        A Hilbert space $\mathcal{H}$ equipped with a unital $\ast$-algebra structure that satisfies the following conditions for all $a,b,c\in\mathcal{H}$:
        \begin{enumerate}
            \item $\langle ab|c \rangle = \langle b|a^*c \rangle$, and
            \item $\langle ab|c \rangle = \langle a|cb^* \rangle$.
        \end{enumerate}
    }
    \begin{example}[Linear operators]\index{Hilbert-Schmidt!norm}\label{operators:hilbert_schmidt_inner_product}
        The canonical example of $H^*$-algebras is given by the algebra of linear operators on a Hilbert space $\mathcal{H}$, where the involution is given by taking adjoints and the inner product is the Hilbert-Schmidt inner product induced by the norm \ref{linalgebra:hilbert_schmidt_norm} (up to a factor $k>0$):
        \begin{gather}
            \langle f|g \rangle_\text{HS} := k\,\tr(f^*g).
        \end{gather}
        The resulting space is denoted by $L^2(\mathcal{H},k)$. A result, analogous to the Artin-Wedderburn theorem \ref{algebra:artin_wedderburn}, states that every $H^*$-algebra can be decomposed as an orthogonal direct sum of finitely many algebras of the form $L^2(\mathcal{H}_i,k_i)$.
    \end{example}

\subsection{Positive maps}

    \newdef{Positive element}{\index{positive}
        A self-adjoint element of a $C^*$-algebra $A$ for which its spectrum is contained in $[0,\infty[$. The cone of all positive elements in $A$ is often denoted by $A^+$.
    }
    \begin{property}[Positive decomposition]
        Every positive element $a$ can be written as $a=b^*b$ for some element $b$. Moreover, every positive element admits a positive square-root.
    \end{property}

    \newdef{Cuntz algebra}{\index{Cuntz algebra}
        The $n^{th}$ Cuntz algebra $\mathcal{O}_n$ is defined as the (universal) unital $C^*$-algebra generated by $n$ isometric elements $s_i$ under the additional relation
        \begin{gather}
            \sum_{i=1}^ns_i^*s_i = 1,
        \end{gather}
        where 1 is the unit element.
    }

    \newdef{Positive map}{\label{operators:positive_map}
        A morphism of $C^*$-algebras is called positive if every positive element is mapped to a positive element.
    }
    \newdef{Completely positive map}{\index{positive!completely}\label{operators:cp_map}
        \nomenclature[A_CP]{CP}{completely positive}
        A morphism of $C^*$-algebras $T:A\rightarrow B$ is called completely positive if for all $k\in\mathbb{N}$ the following map is positive:
        \begin{gather}
            \mathbbm{1}_k\otimes T:\mathbb{C}^{k\times k}\otimes A\rightarrow \mathbb{C}^{k\times k}\otimes B.
        \end{gather}
        If $T$ satisfies this condition only up to an integer $n$, it is said to be \textbf{$n$-positive}.
    }

    \newdef{State}{\index{state}
        A positive linear functional of unit norm on a $C^*$-algebra.
    }

    \newdef{Positivity-improving map}{
        A positive map $\phi$ that satisfies
        \begin{gather}
            a\geq0,a\neq0\implies\phi(a)>0.
        \end{gather}
    }
    \newdef{Ergodic map}{\index{ergodic}
        A positive map $\phi$ that satisfies
        \begin{gather}
            \forall a>0:\exists t\in\mathbb{R}_0:\exp(t_a\phi)a>0.
        \end{gather}
    }

    \begin{property}[Convexity]\index{state!pure}
        The set of states is a convex set. The extreme points of this set are called \textbf{pure states}, all other elements are called \textbf{mixed states}.
    \end{property}

    \begin{example}[Function algebras]
        Let $X$ be a locally compact Hausdorff space and consider its algebra of functions with compact support $C_c(X)$. By the Riesz-Markov theorem \ref{distributions:riesz_markov} the positive linear functionals on this algebra correspond to Radon measures on $X$. The subspace of states then correspond exactly to the probability measures on $X$. Under this identification the pure states correspond to Dirac measures (evaluation functionals).
    \end{example}

\subsection{Traces}

    \newdef{Trace}{\index{trace}
        Let $A$ be a $C^*$-algebra. A trace on $A$ is a linear functional that satisfies the following conditions:
        \begin{enumerate}
            \item\textbf{Positivity}: $\tr(A^+)\geq 0$, and
            \item\textbf{Tracial}: $\tr(ab)=\tr(ba)$ for all $a,b\in A$.
        \end{enumerate}
    }
    \remark{The tracial property could have been replaced by the following equivalent, but seemingly weaker, condition: $\tr(a^*a)=\tr(aa^*)$ for all $a\in A$.}

    \newdef{Adjoint map}{\index{adjoint!operator}
        Assume that a trace functional $\tr$ is given on a $C^*$-algebra and consider a continuous linear map $\phi$ defined on the Schatten class $\mathcal{I}_p$. One can define the adjoint map $\phi^*$ on $\mathcal{I}_q$ whenever $p,q$ are H\"older conjugate. This adjoint is given by the following equation:
        \begin{gather}
            \tr\Big((\phi^*(A))^*B\Big) = \tr\Big(A^*\phi(B)\Big),
        \end{gather}
        where $A\in\mathcal{I}_q,B\in\mathcal{I}_p$.
    }
    \newdef{Trace-preserving map}{\label{operators:trace_preserving}
        A map $\phi$ is said to be trace-preserving if it satisfies
        \begin{gather}
            \tr(\phi(A)) = \tr(A)
        \end{gather}
        for all trace-class elements $A$. Using the above definition it is easily seen that on a unital $C^*$-algebra this is equivalent to
        \begin{gather}
            \phi^*(1) = 1.
        \end{gather}
    }

    \begin{property}
        A completely positive, trace-preserving map $\phi$ satisfies:
        \begin{gather}
            \|\phi\|_1 = 1,
        \end{gather}
        where the subscript $1$ indicates that this operator is defined on trace-class elements.
    \end{property}

    \begin{property}[States]\index{state}
        Whenever a $C^*$-algebra is commutative, every state defines a trace and, possibly after a suitable normalization, every trace defines a state. However, for noncommutative $C^*$-algebras only the latter implication holds.
    \end{property}

\subsection{Representations}

    \newdef{$C^*$-algebra representation}{\index{representation!of an algebra}
        A representation of a $C^*$-algebra $A$ is a unital $\ast$-morphism $A\rightarrow\mathcal{B}(\mathcal{H})$.
    }

    \newdef{Normal state}{\index{state!normal}\label{operators:normal_state}
        Consider a Hilbert space $\mathcal{H}$ and a $C^*$-representation $\pi:A\rightarrow\mathcal{B}(\mathcal{H})$. A normal state $\omega$ is a state such that there exists a trace-class operator $\rho\in\mathcal{B}_1(\mathcal{H})$ with the following property:
        \begin{gather}
            \omega(a) = \frac{\tr(\rho\pi(a))}{\tr(\rho)}.
        \end{gather}
        In case where $A=\mathcal{B}(\mathcal{H})$, the normal states are exactly the $\sigma$-weakly continuous states \ref{functional:sigma_weak_topology}.
    }

    \newdef{Cyclic vector}{\index{cyclic!vector}
        A cyclic vector for a $C^*$-algebra representation $\rho:A\rightarrow\mathcal{B}(\mathcal{H})$ is a vector $\xi\in\mathcal{H}$ such that $\{\rho(a)\xi\mid a\in A\}$ is (norm) dense in $\mathcal{H}$.
    }

    \begin{theorem}[Stinespring]\index{Stinespring}\label{operators:stinespring}
        Consider a linear operator $\Phi:\mathcal{B}(\mathcal{H}_1)\rightarrow\mathcal{B}(\mathcal{H}_2)$. It is completely positive if and only if there exists a $C^*$-representation $\pi:\mathcal{B}(\mathcal{H}_2)\rightarrow\mathcal{B}(\mathcal{K})$ and a bounded linear operator $V:\mathcal{H}_1\rightarrow\mathcal{K}$ such that
        \begin{gather}
            \Phi^\dagger(a) = V^\dag\pi(a)V
        \end{gather}
        or, equivalently,
        \begin{gather}
            \Phi(\rho) = \pi^\dagger(V\rho V^\dagger).
        \end{gather}
        Moreover, $V$ is an isometry if and only if $\Phi$ is trace-preserving.
    \end{theorem}
    \begin{remark}
        Because the adjoint of a completely positive map is again completely positive, the above two characterizations can be used interchangeably. The former is the mostly used by mathematicians, while the latter is better known in the physics literature.

        The most common situation is where $\mathcal{H}:=\mathcal{H}_1=\mathcal{H}_2$ and $\pi:\mathcal{B}(\mathcal{H})\rightarrow\mathcal{B}(\mathcal{H}\otimes\mathcal{K}):a\mapsto a\otimes b$, for some density operator $B\in\mathcal{B}(\mathcal{K})$. The adjoint to tensoring by a density operator is taking the partial trace $\tr_\mathcal{K}$. This way one obtains the following expressions:
        \begin{gather}
            \Phi^\dagger(a) = V^\dag(a\otimes b)V
        \end{gather}
        and, equivalently,
        \begin{gather}
            \Phi(\rho) = \tr_\mathcal{K}(V\rho V^\dagger).
        \end{gather}
    \end{remark}

    \newdef{Hilbert module}{\index{Hilbert!module}
        Let $A$ be a $C^*$-algebra and $H$ a vector space. $H$ is a (right) pre-Hilbert $A$-module if there exists a map
        \begin{gather}
            \langle\cdot\mid\cdot\rangle_A:H\times H\rightarrow A
        \end{gather}
        satisfying the following conditions:
        \begin{enumerate}
            \item\textbf{Conjugate linearity}: For all $x,y,z\in H$:
                \begin{gather}
                    \langle x|y+z \rangle_A = \langle x|y \rangle_A+\langle x|z \rangle_A \qquad\qquad \langle x|y \rangle_A = \langle y|x \rangle_A^*.
                \end{gather}
            \item\textbf{Nondegeneracy}: For all $x\in H$:
                \begin{gather}
                    \langle x|x \rangle_A\geq0 \qquad\qquad \langle x|x \rangle_A = 0 \iff x = 0.
                \end{gather}
            \item\textbf{Equivariance}: For all $x,y\in H$ and $a\in A$:
                \begin{gather}
                    \langle x|y \rangle_Aa = \langle x|ya \rangle_A.
                \end{gather}
        \end{enumerate}
        A left module is obtained by requiring linearity and (left-)equivariance in the first argument. The completion of a pre-Hilbert module with respect to the norm $x\mapsto\sqrt{\|\langle x|x \rangle_A\|}$ is called a \textbf{Hilbert $A$-module}.
    }

    \newdef{Fredholm module}{\index{Fredholm!module}\label{operators:fredholm_module}
        Let $A$ be unital a $C^*$-algebra with a $\ast$-representation $\pi:A\rightarrow\End(\mathcal{H})$ on a Hilbert space $\mathcal{H}$. When equipped with an operator $F\in\End(\mathcal{H})$ this gives an (\textbf{odd}) Fredholm module if
        \begin{gather}
            F=F^*\qquad\text{and}\qquad F^2=\mathbbm{1}_\mathcal{H}\qquad\text{and}\qquad[F,\pi(A)]\subseteq\mathcal{B}_0(\mathcal{H}).
        \end{gather}
        The Fredholm operator is said to be \textbf{even} if $\mathcal{H}$ is a super-Hilbert space with grading $\Gamma\in\End(\mathcal{H})$ satisyfing
        \begin{gather}
            \Gamma^*=\Gamma\qquad\text{and}\qquad\Gamma^2=\mathbbm{1}_\mathcal{H}\qquad\text{and}\qquad[\Gamma,\pi(A)]=0\qquad\text{and}\qquad\{\Gamma,F\}_+=0.
        \end{gather}
        $F$ has the form
        \begin{gather}
            \begin{pmatrix}
                0&F_+\\F_-&0
            \end{pmatrix},
        \end{gather}
        where $F_+$ is Fredholm.
    }
    \begin{remark}[Kasparov $K$-theory]\index{K-theory!Kasparov}
        If $A$ is not unital, one should relax the first two relations up to compact operators, i.e. $F=F^*$ and $F^2=\mathbbm{1}_\mathcal{H}$ in the Calkin algebra. This gives a clear relation to \textit{Kasparov's $K$-theory}, where the homology complex $KK_\bullet$ is defined in terms of such generalized Fredholm modules. However, it can be shown that these induce the same $KK$-classes.
    \end{remark}

    \begin{example}
        Consider $A=\mathbb{C}$ and let $\pi:\mathbb{C}\rightarrow\End(\mathcal{H})$ be the unique unital representation. Fredholm modules with this data correspond to (essentially self-adjoint) Fredholm operators on $\mathcal{H}$.
    \end{example}

\subsection{Gel'fand duality}\index{Gel'fand!duality}

    \newdef{Gel'fand spectrum}{\index{Gel'fand!spectrum}
        Consider a commutative $C^*$-algebra $A$ (in fact any involutive algebra suffices). Its set of characters, i.e. the algebra morphisms $A\rightarrow\mathbb{C}$, can be equipped with a locally compact Hausdorff topology (the weak-* topology \ref{functional:weak_star_topology}). This space is compact if and only if the algebra is unital.
    }
    \newdef{Gel'fand representation}{
        Consider a $C^*$-algebra $A$ and let $\Phi_A$ denote its Gel'fand spectrum. The Gel'fand transformation of an element $a\in A$ is defined as the morphism $\hat{a}:\Phi_A\rightarrow\mathbb{C}$ given by the following formula:
        \begin{gather}
            \hat{a}(\lambda) = \langle\lambda,a\rangle
        \end{gather}
        where $\langle\cdot,\cdot\rangle$ denotes the pairing between $A$ and $\Phi_A$. By definition of the topology on the Gel'fand spectrum the functional $\hat{a}$ is continuous for all $a\in A$. The mapping $a\mapsto\hat{a}$ is called the Gel'fand representation of $A$.
    }

    \begin{theorem}[Gel'fand-Naimark: Commutative case]\index{Gel'fand-Naimark}
        Let $A$ be a commutative $C^*$-algebra. The Gel'fand representation gives an isometric $\ast$-isomorphism between $A$ and the set $C_0(\Phi_A)$ of continuous complex-valued functions that vanish at infinity on its Gel'fand spectrum.
    \end{theorem}
    \begin{remark}
        In fact, the Gel'fand-Naimark theorem gives an equivalence between the category of locally compact Hausdorff spaces and the category of commutative $C^*$-algebras.
    \end{remark}

    \begin{property}
        A compact Hausdorff space is connected if and only if its algebra of continuous functions has no nontrivial projections.
    \end{property}
    \begin{property}
        A commutative $C^*$-algebra is separable if and only if its Gel'fand spectrum is metrizable.
    \end{property}

    \begin{formula}[Spectrum]
        Recall Definition \ref{functional:spectrum} of the spectrum of an operator. This is related to the spectrum of a unital, commutative $C^*$-algebra as follows:
        \begin{gather}
            \sigma(a) = \{a(x)\mid x\in\Phi_A\},
        \end{gather}
        for all $a\in A$.
    \end{formula}

    \begin{construct}[Gel'fand-Naimark-Segal]\index{Gel'fand-Naimark-Segal construction}\label{operators:gns}
        \nomenclature[A_GNS]{GNS}{Gel'fand-Naimark-Segal}
        Let $A$ be a $C^*$-algebra. Given a state $\omega$ on $A$ there exists a $C^*$-representation $\rho:A\rightarrow\mathcal{B}(D)$, where $D\subset\mathcal{H}$ is a dense subspace of a Hilbert space $\mathcal{H}$, such that the following conditions are satisfied:
        \begin{itemize}
            \item There exists a distinguished cyclic unit vector $\xi$ such that $D = \{\rho(a)\xi\mid a\in A\}$.
            \item For all elements $a\in A$ the following equality holds:
                \begin{gather}
                    \omega(a) = \langle\rho(a)\xi|\xi\rangle.
                \end{gather}
        \end{itemize}

        ?? COMPLETE CONSTRUCTION ??
    \end{construct}

    \begin{theorem}[Gel'fand-Naimark: General case]\index{Gel'fand-Naimark}\label{operators:gelfand_naimark}
        Every $C^*$-algebra is isometrically $\ast$-isomorphic to a norm closed ($C^*$-)algebra of bounded operators on a Hilbert space $\mathcal{H}$.
    \end{theorem}

\section{von Neumann algebras}

    \newdef{von Neumann algebra}{\index{von Neumann!algebra}
        A $*$-subalgebra of a $C^*$-algebra equal to its double commutant: $M = M''$.
    }
    \newdef{Concrete von Neumann algebra}{
        A weakly closed unital $*$-algebra of bounded operators on some Hilbert space.
    }
    \begin{theorem}[Double Commutant theorem\footnotemark]
        \footnotetext{Often called \textbf{von Neumann's double commutant theorem}.}
        The above definitions are equivalent.
    \end{theorem}

    \newdef{Projection}{\index{projection}\label{operators:projection}
        An element $p$ of a von Neumann algebra is called a projection if it satisfies
        \begin{gather}
            p = p^2 = p^*.
        \end{gather}
        This terminology reflects the property that if a von Neumann algebra is regarded as an algebra of bounded operators, the projections are exactly the operators associated to an orthogonal projection.
    }
    \begin{property}
        Any von Neumann algebra is generated by its projections.
    \end{property}

    \newdef{Murray-von Neumann equivalence}{\index{equivalence!Murray-von Neumann}
        Two closed subspaces are said to be Murray-von Neumann equivalent if one is mapped isomorphically onto the other by a partial isometry. In terms of projections this means that $p\sim q$ if and only if there exists a partial isometry $u$ such that $p=uu^*$ and $q=u^*u$.
    }

    \newdef{Finite projection}{\index{projection!finite}
        The collection of projections inherits the structure of a partial order from the partial order on the corresponding subspaces. A projection $p$ is said to be finite if there exists no smaller projection $q$ that is equivalent to $p$.
    }

\subsection{Factors}

    \newdef{Factor}{\index{factor}
        Consider a von Neumann algebra $M$. A $*$-subalgebra $A$ is called a factor of $M$ if its center $Z(A)$ is given by the scalar multiples of the idenity.
    }

    \newdef{Type $\mathrm{I}$ factor}{
        A factor is of type $\mathrm{I}$ if it contains a \textit{minimal projection}.
    }
    \begin{property}[Type $\mathrm{I}_n$ factors]
        Any type $\mathrm{I}$ factor is isomorphic to the algebra of all bounded operators on a Hilbert space. To indicate the dimension $n$ of this Hilbert space (which may be $\infty$) one sometimes uses the subclassification of type $\mathrm{I}_n$ factors.
    \end{property}

    \newdef{Powers index}{\index{index!Powers}
        Consider a Hilbert space $\mathcal{H}$ together with its von Neumann algebra of bounded operators $\mathcal{B}(\mathcal{H})$. A unital $*$-endomorphism $\alpha$ is said to have Powers index $n\in\mathbb{N}$ if the space $\alpha(\mathcal{B}(\mathcal{H}))$ is isomorphic to a type $\mathrm{I}_n$ factor.
    }

    \newdef{Type $\mathrm{II}$ factor}{
        A factor is of type $\mathrm{II}$ if it contains nonzero \textit{finite projections} but no minimal ones. If the identity is finite, the factor is sometimes said to be of type $\mathrm{II}_1$, otherwise it is of type $\mathrm{II}_\infty$.
    }

    \newdef{Type $\mathrm{III}$ factor}{
        A factor is of type $\mathrm{III}$ if it does not contain any nonzero finite projections.
    }

\subsection{Projection-valued measures}\label{section:PVM}

    This section focuses on the algebra of bounded operators $\mathcal{B}(\mathcal{H})$ on a (complex) Hilbert space $\mathcal{H}$.

    \begin{property}[Closed subspaces]
        There exists a bijection between the set of closed subspaces of $\mathcal{H}$ and the set of projections in $\mathcal{B}(\mathcal{H})$. Furthermore, if the projection $p$ corresponds to a subspace $\mathcal{H}_p$, the projection $\mathbbm{1}_\mathcal{H}-p$ corresponds to the orthogonal complement $\mathcal{H}^\perp_p$.
    \end{property}

    \newdef{Projection-valued measure}{\index{measure!projection-valued}\index{spectral!measure}
        \nomenclature[A_PVM]{PVM}{projection-valued measure}
        Consider a topological space $X$ and let $\Sigma$ be a $\sigma$-algebra \ref{set:sigma_algebra} on $X$. A projection-valued measure (PVM) or \textbf{spectral measure} on $X$ is a map $P:\Sigma\rightarrow\mathcal{B}(\mathcal{H})$ satisfying the following conditions:
        \begin{enumerate}
            \item $P_E$ is a projection for all $E\in\Sigma$,
            \item $P_X = \mathbbm{1}_\mathcal{H}$,
            \item $P_AP_B = P_{A\cap B}$, and
            \item for all disjoint $\seq{E}\subset\Sigma$:
                \begin{gather}
                    \sum_{n\in\mathbb{N}}P_{E_n} = P_{\cup_{n\in\mathbb{N}}E_n}.
                \end{gather}
        \end{enumerate}
    }
    \begin{property}
        For every two elements $v,w\in\mathcal{H}$ the map
        \begin{gather}
            E\mapsto\mu^P_{v,w}(E):=\langle v|P_Ew \rangle
        \end{gather}
        defines a (complex) measure $\mu^P_{v,w}$ on $X$. The square of the norm of an element $v\in\mathcal{H}$ is then simply given by $\mu^P_{v,v}(X)$ for any PVM $P$ due to the second condition above.
    \end{property}

    \begin{property}
        Let $f:X\rightarrow\mathbb{C}$ be a measurable function on a measurable space $(X,\Sigma)$. Given a PVM $P$ on $X$, one defines $\Delta_f$ to be the set of all $v\in\mathcal{H}$ for which $f\in L^2(X,\mu^P_{v,v})$. The operator $\int_Xf(\lambda)\,dP(\lambda):\Delta_f\rightarrow\mathcal{H}$ defined by
        \begin{gather}
            \left\langle v\left|\int_Xf(\lambda)\,dP(\lambda)w\right.\right\rangle = \int_Xf(\lambda)\,d\mu^P_{v,w}(\lambda)
        \end{gather}
        is closed and normal. Furthermore, it satisfies the following two equalities:
        \begin{align}
            \left(\int_Xf(\lambda)\,dP(\lambda)\right)^* &= \int_X\overline{f(\lambda)}\,dP(\lambda)\\
            \left\|\int_Xf(\lambda)\,dP(\lambda)v\right\|^2 &= \int_X|f(\lambda)|^2d\mu^P_{v,v}(\lambda).
        \end{align}
        If $f$ is bounded, the above operator is bounded by the supremum norm of $f$:
        \begin{gather}
            \left\|\int_Xf(\lambda)dP(\lambda)\right\|\leq\|f\|_\infty.
        \end{gather}
    \end{property}

    \begin{theorem}[Spectral decomposition]
        Let $A$ be a self-adjoint operator on a Hilbert space $\mathcal{H}$. There exists a unique projection-valued measure $P_A:\mathcal{B}(X)\rightarrow\mathcal{B}(\mathcal{H})$ on the Borel $\sigma$-algebra of the real line such that
        \begin{gather}
            A = \int_\mathbb{R}\lambda\,dP_A(\lambda).
        \end{gather}
        If $\mathbb{R}$ is replaced by $\mathbb{C}$, this theorem also holds for normal operators. If $A$ is compact, the measure becomes purely atomic, i.e. there exists an orthonormal basis of $\mathcal{H}$ consisting of eigenvectors of $A$ and, moreover, the eigenvalues are necessarily real.
    \end{theorem}
    \begin{property}[Spectrum and support]
        The spectrum of a self-adjoint operator $A$ coincides with the support of its associated spectral measure $P_A$. A number $\lambda\in\mathbb{R}$ belongs to the point spectrum of $A$ if and only if the PVM associated to $A$ does not vanish on $\{\lambda\}$. A number $\lambda\in\mathbb{R}$ belongs to the continuous spectrum of $A$ if the associated PVM vanishes on $\{\lambda\}$ but is nonvanishing on any open set containing $\lambda$.
    \end{property}

    \newdef{Singular value}{\index{singular!value}
        Let $A$ be a compact operator. The singular values of $A$ are given by the square roots of the eigenvalues of the self-adjoint (and compact) operator $A^*A$. If $A$ is self-adjoint, its singular values and eigenvalues coincide.
    }

    The above property allows to compose self-adjoint operators with (measurable) functions similar to how one can compute $f(X)$ for finite-dimensional operators by applying $f$ to the eigenvalues of $X$:
    \begin{formula}[Function of operators]
        Let $f:\sigma(A)\rightarrow\mathbb{C}$ be a measurable function (with respect to the restriction of the Borel algebra on $\mathbb{R}$) and let $g:\mathbb{R}\rightarrow\mathbb{C}$ be any other measurable function that coincides with $f$ on $\sigma(A)$.
        \begin{gather}
            f(A) := \int_{\sigma(A)}f(\lambda)\,dP_A(\lambda) = \int_\mathbb{R}g(\lambda)\,dP_A(\lambda) =: g(A).
        \end{gather}
    \end{formula}

    \newdef{Dixmier ideal}{\index{Dixmier!ideal}
        For every compact operator one can sort the singular values in decreasing order. One can then define the operator
        \begin{gather}
            \sigma_n(A) := \sum_{i=1}^ns_i(A).
        \end{gather}
        The Dixmier ideal is defined as follows:
        \begin{gather}
            \mathfrak{D}(\mathcal{H}):=\{A\in\End_0(\mathcal{H})\mid\sigma_n(A)=O(\ln(n))\}.
        \end{gather}
    }
    The following functional gives an alternative for the ordinary trace on operators that are not trace-class.
    \newdef{Dixmier trace}{\index{Dixmier!trace}\index{measurable!operator}\label{operators:dixmier_trace}
        For every element in $A\in\mathfrak{D}(\mathcal{H})$, the Dixmier trace is defined as follows:
        \begin{gather}
            \tr_\mathfrak{D}(A) := \omega\left(\left\{\frac{\sigma_n(A)}{1+\log(n)}\mid n\in\mathbb{N}\right\}\right),
        \end{gather}
        where $\omega:\ell^\infty(\mathbb{N})$ is a linear functional satisfying the conditions
        \begin{enumerate}
            \item\textbf{Positivity}: $\omega(s)\geq0$ if $s_n\geq0$ for all $n\in\mathbb{N}$.
            \item\textbf{Convergence}: $\omega(s)=\lim_{n\rightarrow\infty}s_n$ if $s$ is convergent.
            \item\textbf{Dilation-invariant}: Let $D:\ell^\infty(\mathbb{N})\rightarrow\ell^\infty(\mathbb{N})$ be the following operator:
            \begin{gather}
                D:(s_1,s_2,\ldots)\mapsto(s_1,s_1,s_2,s_2,\ldots).
            \end{gather}
            The state $\omega$ is dilation-invariant if $\omega(s)=\omega(D(s))$ for all $s\in\ell^\infty(\mathbb{N})$.
        \end{enumerate}
        If for an operator $A\in\mathfrak{D}(\mathcal{H})$ the Dixmier trace is independent of the chosen state, it is said to be \textbf{measurable}.
    }
    \begin{property}\index{singular!trace}
        The Dixmier trace has the following properties:
        \begin{itemize}
            \item $\tr_\mathfrak{D}(A+B)=\tr_\mathfrak{D}(A) + \tr_\mathfrak{D}(B)$ for all $A,B\in\mathfrak{D}(\mathcal{H})$.
            \item If $A$ is bounded and $B$ is an element of the Dixmier ideal, then $\tr_\mathfrak{D}(AB)=\tr_\mathfrak{D}(BA)$.
            \item The Dixmier ideal vanishes on trace-class operators (in particular, it is \textbf{singular}, i.e. it vanishes on finite-rank operators):
                \begin{gather}
                    A\in\mathcal{B}_1(\mathcal{H})\implies\tr_\mathfrak{D}(A)=0.
                \end{gather}
        \end{itemize}
    \end{property}