\chapter{Derivations: Mathematics}

\section{Group theory}
\subsection{Proof of Property \ref{group:transitive_action_property}}\label{proof:stabilizer}

    Choose an element $x\in X$. The stabilizer of $x$ with respect to $G$ is the set \[S_x := \{g\in G\mid g\cdot x = x\}.\] Due to the transitivity of the group action one obtains that \[\forall x, y\in X: \exists h\in G: h\cdot x = y.\] So, for every $z\in X$ one can choose a group element $g_z$ such that $g_z\cdot x = z$. For all elements in the coset $g_zS_x = \{g_zs\in G\mid s\in S_x\}$ the following equality is satisfied: \[(g_zs)\cdot x = g_z\cdot (s\cdot x) = g_z\cdot x = z.\] This implies that the map $\Phi:G/S_x \rightarrow X$ is surjective.

    Now, one needs to prove that $\Phi$ is also injective. A proof by contradiction is given. Choose two distinct cosets $gS_x$ and $hS_x$. There exist two elements $G,H\in X$ such that $g\cdot x = G$ and $h\cdot x = H$. Assume that $G = H$. This means that
    \begin{align*}
        &g\cdot x = h\cdot x\\
        \iff&(h^{-1}g)\cdot x = x\\
        \iff&h^{-1}g\in S_x\\
        \iff&hS_x\ni h(h^{-1}g) = g.
    \end{align*}
    This would imply that $gS_x = hS_x$, which is in contradiction with the assumptions. It follows that $G\neq H$ and, hence, that $\Phi$ is injective.\qed

\section{Calculus}
\subsection{Proof of Method \ref{calculus:borel_transform}}

    The function $F$ is defined as follows:
    \begin{gather}
        F(x) := \\sum_{n=0}^\infty\frac{a_n}{n!}x^n.
    \end{gather}
    Perform a Borel transform:
    \begin{align*}
        \int_0^\infty F(xt)e^{-t}dt&=\sum_{n=0}^N\int_0^\infty\frac{a_n}{n!}x^nt^ne^{-t}dt\\
        &=\sum_{n=0}^N\frac{a_n}{n!}x^n\int_0^\infty t^ne^{-t}dt\\
        &=\sum_{n=0}^N\frac{a_n}{n!}x^n\Gamma(n+1)\\
        &=\sum_{n=0}^Na_nx^n,
   \end{align*}
   where the definition of the Gamma function \ref{calculus:gamma_function} was used on line 3 and the relation between the factorial function and the Gamma function \eqref{calculus:gamma_factorial_relation} was used on line 4.\qed

\section{Linear algebra}
\subsection{Proof of equivalence of Definitions \ref{vector:exterior_algebra} and \ref{vector:adef_exterior_algebra}}

    \begin{gather}
        (u+v)\otimes(u+v) - u\otimes u - v\otimes v = u\otimes v + v\otimes u
    \end{gather}
    The left-hand side is an element of the ideal $I$ generated by $\{v\otimes v\mid v\in V\}$. Using the ideal generated by elements of the form of the right-hand side gives the usual definition of the exterior algebra based on the wedge product as defined in \ref{vector:wedge_product} because it imposes the relation $u\wedge v = -v\wedge u$.

    However, one should pay attention to one little detail. As mentioned in \ref{vector:adef_exterior_algebra} the general definition uses the ideal $I$ to construct the quotient space. The other construction is only equivalent when working over a field with characteristic different from 2. This follows from the fact that one has to divide by 2 when trying to obtain the ideal $I$ from the RHS when setting $u=v$.

\section{Manifolds and bundles}
\subsection{Proof of equivalence of Definitions \ref{manifold:tangent_vector_partial} and \ref{manifold:alternative_definition}}

    Let $(U,\varphi)$ be a chart around the point $p\in M$. Using the first definition of a tangent vector \ref{manifold:tangent_vector_partial}, i.e. \[\left.\pderiv{}{q^i}\right|_{p}:\mathcal{F}_p(M, \mathbb{R})\rightarrow\mathbb{R}:f\mapsto \pderiv{}{q^i}\left(f\circ\varphi^{-1}\right)(\varphi(p))\] one can rewrite Equation \eqref{manifold:tangent_vector_chain_rule} \[v_p(f) = \pderiv{(f\circ\varphi^{-1})}{q^i}(\varphi(p))\deriv{q^i}{t}(0)\] as follows: \[v_p(f) = \left.\pderiv{f}{q^i}\right|_p\deriv{q^i}{t}(0).\] Because the partial derivatives as defined in \ref{manifold:tangent_vector_partial} form a basis for the tangent space (by construction), one can see that this equation is in fact an expansion of the tangent vector $v_p$ in terms of that basis. It follows that vectors tangent to curves\footnote{More precisely, representatives of equivalence classes of vectors tangent to curves.} are also tangent vectors according to the first definition.

    To prove the other direction we have to show that the partial derivative operators can be constructed as vectors tangent to curves.

    A tangent vector can be expressed, according to the first construction, in the following way: \[v_p = v^i\left.\pderiv{}{q^i}\right|_p,\] where the definition $v = (v^1,\ldots,v^n)$ was used. One can then construct the curve $\gamma:t\mapsto\varphi^{-1}(q_0+vt)$. It is obvious that the tangent vector $v_p$ is tangent to the curve $\gamma$. From this it follows that there exists an isomorphism between the tangent vectors from the first definition and the equivalence classes of vectors tangent to curves from the second definition.\qed

    Although the previous equivalence implies that the tangent space construction using germs of curves gives a vector space, one could also check the vector space axioms directly. First, one should prove that the sum of vectors tangent to the curves $\gamma$ and $\delta$ is again a vector tangent to some curve $\chi:\mathbb{R}\rightarrow M$. To this end, define the curve \[\chi(t)\equiv\varphi^{-1}\circ\Big(\varphi\circ\gamma(t) + \varphi\circ\delta(t) - \varphi(p)\Big),\] where $\varphi$ is again the coordinate map in some chart $(U,\varphi)$ around $p\in M$. Using Equation \eqref{manifold:tangent_vector_chain_rule} one can find
    \begin{align*}
        v_{p,\chi}(f) &= \pderiv{(f\circ\varphi^{-1})}{q^i}(\varphi(p))\deriv{(\varphi^i\circ\chi)}{t}(0)\\
        &=\pderiv{(f\circ\varphi^{-1})}{q^i}(\varphi(p))\deriv{}{t}\left(\varphi^i\circ\gamma + \varphi^i\circ\delta - \varphi^i(p)\right)\\
        &=\pderiv{(f\circ\varphi^{-1})}{q^i}(\varphi(p))\left(\deriv{(\varphi^i\circ\gamma)}{t} + \deriv{(\varphi^i\circ\delta)}{t}\right)\\
        &=v_{p,\gamma}(f) + v_{p,\delta}(f).
    \end{align*}
    The constant term $-\varphi(p)$ in the definition of $\chi$ is necessary to make sure that $\chi(0) = \gamma(0) = \delta(0) = p$. The axiom of scalar multiplication by a number $\lambda\in K$ can be proven similarly by defining the curve \[\chi(t) = \varphi^{-1}\circ\left[\lambda\Big(\varphi\circ\gamma(t)\Big)\right].\]

\subsection{Explanation for Example \ref{bundle:lie_derivative_function}}\index{Landau!little-o notation}

    In this derivation Landau's little-o notation $o(t)$ is used:
    \begin{gather}
        \lim_{t\rightarrow0}\frac{o(t)}{t} = 0
    \end{gather}
    Now, assume that $X$ is a smooth vector field and $f$ is a smooth function. Because the Lie derivative is a local operation one can work in a local chart such that $\gamma$ is (again locally) equivalent to a curve\footnote{The vector field $X(p) = (p,Y(p))$, where $Y$ is a smooth vector field on $\mathbb{R}^n$, can also be identified with $Y$ itself. This is implicitly done in the derivation by using the notation $X$ for both vector fields.} $\beta_p:U\rightarrow\mathbb{R}^n$ and such that one can expand $\beta_p(t)$ around $p\in U$:
    \begin{align}
        \mathcal{L}_Xf(p) &= \lim_{t\rightarrow0}\left[\frac{f(\beta_p(0) + t\beta_p'(0) + o(t)) - f(p)}{t}\right]\nonumber\\
        &= \lim_{t\rightarrow0}\left[\frac{f(p + tX(p) + o(t)) - f(p)}{t}\right]\nonumber\\
        &= \lim_{t\rightarrow0}\left[\frac{f(p) + tDf(p)\cdot X(p) + o(t) - f(p)}{t}\right]\nonumber\\
        &= \sum_k\pderiv{f}{x^k}(p)X_k(p) + \lim_{t\rightarrow0}\frac{o(t)}{t}\nonumber\\
        &= \sum_k\pderiv{f}{x^k}(p)X_k(p),\label{diff:lie_derivative_calc}
    \end{align}
    where the defining condition \ref{bundle:integral_curve} for integral curves was used on the second line. If this equation is rewritten as an operator equality, one obtains
    \begin{gather}
        \mathcal{L}_X = \sum_kX_k\pderiv{}{x^k}.
    \end{gather}

\subsection{Explanation for Formula \ref{bundle:lie_derivative_vector_field}}

    For vector fields one cannot just take the difference at two different points because the tangent spaces generally do not coincide. This can be resolved by using the flow \ref{bundle:flow}:
    \begin{gather}
        \label{derivmath:lie_derivative_vector_fields}
        \mathcal{L}_XY = \lim_{t\rightarrow0}\frac{(T\sigma_t)^{-1}X(\gamma_p(t)) - X(p)}{t},
    \end{gather}
    where $T\sigma_t$ is the differential \ref{bundle:differential} of the flow, which satisfies $(T\sigma)^{-1} = T\sigma_{-t}$. To see that this definition makes sense, one has to show that $(T\sigma_t)^{-1}[X(\gamma_p(t))]\in T_pM$. This goes as follows:
    \begin{align*}
        (T\sigma_t)^{-1}X(\gamma_p(t))(f) &= T\sigma_{-t}X(\gamma_p(t))(f)\\
        &= X(\sigma_{-t}\circ\gamma_p(t))(f\circ\sigma_{-t})\\
        &= X(\sigma_{-t}\circ\sigma_t(p))(f\circ\sigma_{-t})\\
        &= X(p)(f\circ\sigma_{-t})\\
        &\in T_pM
    \end{align*}
    for all $f\in C^k(M,\mathbb{R})$. On the third line the definition of the flow \ref{bundle:flow} was used. One can also rewrite the second term in the numerator of \eqref{derivmath:lie_derivative_vector_fields} using the flow: \[X(p) = X(\sigma_0(p)) = T\sigma_0(X).\] Using the definition of the pushforward of vector fields \eqref{bundle:pushforward}, the Lie derivative can be rewritten as follows:
    \begin{align*}
        \mathcal{L}_XY &= \lim_{t\rightarrow0}\frac{\sigma_{-t*}X(\gamma_p(t)) - \sigma_{0*}X(\gamma_p(0))}{t}\\
        &= \left.\deriv{}{t}(\sigma_{-t*}X)(\gamma_p(t))\right|_{t=0}.
    \end{align*}
    Finally, by using the relation between pushforward and pullback \eqref{bundle:pullback_pushforward} this becomes
    \begin{gather}
        \mathcal{L}_XY = \left.\deriv{}{t}(\sigma_t^*X)(\gamma_p(t))\right|_{t=0}.
    \end{gather}

\subsection{Explanation of Remark \ref{bundle:vector_calculus}}

    Looking at Equation \eqref{bundle:function_derivative} for the exterior derivative of a smooth function and remembering the definition of the gradient \ref{vector:gradient}, one can see that these two definitions look very similar. The major difference lies in the fact that $\nabla f$ is a vector in $\mathbb{R}^3$ and $df$ is a covector in $\mathbb{R}^{*3}$. However, there exists an isomorphism between these spaces and so one can identify $\nabla f$ and $df$.

    Similar relations hold for the divergence \ref{vector:divergence} and rotor \ref{vector:rotor}. However, here one has to use a different construction because the spaces $\Lambda^1$ and $\Lambda^2$ have to be used. The Hodge star \ref{vector:hodge_star} can be used to obtain the correct dimensions.

    Consider a vector $\vector{f} = (f_1,f_2,f_3)$ where the $f_i$ are smooth. Using these functions $f_i$ one can construct a 1-form $\alpha = f_1dx_1 + f_2dx_2 + f_3dx_3$ and a 2-form $\omega = f_1dx_2\wedge dx_3 + f_2dx_3\wedge dx_1 + f_3 dx_1\wedge dx_2$. After applying the exterior derivative (in the corresponding spaces) one obtains
    \begin{align}
        d\alpha &= \left(\pderiv{f_3}{x_2} - \pderiv{f_2}{x_3}\right)dx_2\wedge dx_3 + \left(\pderiv{f_1}{x_3} - \pderiv{f_3}{x_1}\right)dx_3\wedge dx_1 + \left(\pderiv{f_2}{x_1} - \pderiv{f_1}{x_2}\right)dx_1\wedge dx_2\nonumber\\
        d\omega &= \left(\pderiv{f_1}{x_1} + \pderiv{f_2}{x_2} + \pderiv{f_3}{x_3}\right)dx_1\wedge dx_2\wedge dx_3.\nonumber
    \end{align}
    Using Corollary \ref{vector:hodge_star_vectorcalculus} and the canonical isomorphism $\sim\ :\mathbb{R}^{3*}\rightarrow\mathbb{R}^3$ one can rewrite this as
    \begin{align}
        \sim df &= \nabla f \\
        \sim (\ast d\alpha) &= \nabla\times\vector{f} \\
        \ast d\omega &= \nabla\cdot\vector{f}.
    \end{align}