\chapter{General Topology}\label{chapter:topology}
\section{Topological spaces}

    \newdef{Topology}{\index{topology}
        \nomenclature[S_Top]{$\mathbf{Top}$}{category of topological spaces}
        Let $X$ be a set and consider a collection of subsets $\tau\subseteq 2^X$. The set $\tau$ is a topology on $X$ if it satisfies the following axioms:
        \begin{enumerate}
            \item $\emptyset\in\tau$ and $X\in\tau$,
            \item $\forall\,\mathcal{F}\subseteq\tau: \bigcup_{V\in\mathcal{F}}V \in \tau$, and
            \item $\forall\,U,V\in\tau: U\cap V\in\tau$.
        \end{enumerate}
        The elements of $\tau$ are called \textbf{open sets} and the couple $(X,\tau)$ is called a \textbf{topological space}. \textbf{Closed sets} are defined as the sets that have an open complement. Because complements are uniquely defined, one could just as well define a topology in terms of closed subsets.
    }

    \begin{property}[\difficult{Category of opens}]
        \nomenclature[S_Open]{$\mathbf{Open}(X)$}{category of open subsets of a topological space $X$}
        Consider a topological space $(X,\tau)$ and let $U\subseteq V\in\tau$. The topology $\tau$ together with the collection of inclusion maps $U\hookrightarrow V$ forms a poset and, by extension, a small category $\mathbf{Open}(X)$.
    \end{property}

    \newdef{Pointed topological space}{\label{topology:pointed_space}\index{pointed!topological space}
        A topological space $X$ with a designated \textbf{base point} $x_0\in X$.
    }

    \begin{example}[Relative topology\footnotemark]\label{topology:relative_topology}
        \footnotetext{Sometimes called the \textbf{subspace} or \textbf{trace topology}.}
        Any subset $Y$ of a topological space $(X,\tau)$ can be turned into a topological space by equipping it with the following topology:
        \begin{gather}
            \tau_{\text{rel}} := \{U_i\cap Y\mid U_i\in\tau\}\,.
        \end{gather}
    \end{example}
    \begin{example}[Discrete topology]\index{discrete!topology}
        The topology in which every subset is open (and thus also closed).
    \end{example}
    \begin{example}[Indiscrete topology]
        The topology in which only the empty set and the space itself are open.
    \end{example}

    \newdef{Interior}{\index{interior}
        \nomenclature[O_zint]{$X^\circ,\overset{\circ}{X}$}{interior of a topological space $X$}
        The interior $Y^\circ$ of a subset $Y$ of a topological space $X$ is defined as the union of all open subsets of $Y$. Elements of the interior are called \textbf{interior points} of $Y$.
    }
    \newdef{Closure}{\index{closure}
        \nomenclature[O_zclos]{$\overline{X}$}{closure of a topological space $X$}
        The closure $\overline{Y}$ of a subset $Y$ of a topological space $X$ is defined as the intersection of all closed sets containing $Y$.
    }
    \newdef{Boundary}{\index{boundary}
        \nomenclature[O_zbound]{$\partial X$}{boundary of a topological space $X$}
        The boundary $\partial Y$ of a subset $Y$ of a topological space $X$ is defined as $\overline{Y}\backslash Y^\circ$.
    }

    \newdef{Borel algebra}{\index{Borel!set}\label{topology:borel_set}
        The $\sigma$-algebra \ref{set:sigma_algebra} generated by all open subsets of a topological space. The elements of this $\sigma$-algebra are called Borel sets.
    }
    \begin{property}
        The closed subsets of a topological space also generate the Borel $\sigma$-algebra.
    \end{property}

    \newdef{Topological group}{\index{group!topological}
        A group equipped with a topology such that both the multiplication and inversion morphisms are continuous.
    }

    \newdef{$G_\delta$-set}{\index{G$_\delta$-set}\label{top:g_delta}
        A countable intersection of open sets.
    }
    \newdef{$F_\sigma$-set}{\index{F$_\sigma$-set}
        A countable union of closed sets.
    }

\subsection{Neighbourhoods}

    \newdef{Neighbourhood}{\index{neighbourhood}
        A set $N\subseteq X$ is a neighbourhood of a point $x\in X$ if there exists an open set $U$ such that $x\in U\subseteq N$.
    }

    Although the following two notions are often treated as synonyms in the literature, they can be given a separate meaning:
    \newdef{Limit point}{\index{limit!point}
        Let $Y$ be a subset of $X$. A point $x\in X$ is called a limit point of $Y$ if every neighbourhood of $x$ contains at least one point of $Y$ different from $x$.
    }
    \newdef{Adherent point}{\index{adherent point}
        Let $Y$ be a subset of $X$. A point $x\in X$ is called an adherent point of $Y$ if every neighbourhood of $x$ contains at least one point of $Y$. A point $x$ is an adherent point of $Y$ if and only if it is an element of the closure $\overline{Y}$.
    }

    \newdef{Accumulation point\footnotemark}{\index{accumulation point|seealso{limit point}}
        \footnotetext{Sometimes called a \textbf{cluster point}.}
        Let $x\in X$ be a limit point of $Y$. It is called an accumulation point of $Y$ if every open neighbourhood of $x$ contains infinitely many points of $Y$.
    }

    \newdef{Basis}{\index{basis}
        A collection $\mathcal{B}\subseteq\tau$ of open subsets of a topological space $(X,\tau)$ is a basis for $\tau$ if every $U\in\tau$ can be written as
        \begin{gather}
            U = \bigcup_{V\in\mathcal{F}}V\,,
        \end{gather}
        where $\mathcal{F}\subseteq\mathcal{B}$.
    }
    \newdef{Local basis}{
        A collection $\mathcal{B}_x$ of open neighbourhoods of a point $x\in X$ is a local basis of $x$ if every neighbourhood of $x$ contains at least one element in $\mathcal{B}_x$.
    }

    \newdef{First-countable space}{\index{countability axiom}
        A topological space $(X,\tau)$ for which for every point $x\in X$ there exists a countable local basis.
    }
    \begin{property}[Decreasing basis]
        Let $x\in X$. If there exists a countable local basis for $x$, there also exists a countable decreasing local basis for $x$.
    \end{property}

    \newdef{Second-countable space}{
        A topological space $(X,\tau)$ for which there exists a countable (global) basis.
    }

    \begin{property}[Closure]\label{topology:closure}
        Let $X$ be a topological space. The closure of a subset $Y\subseteq X$ is given by
        \begin{gather}
            \overline{Y} = \{x\in X\mid\exists\text{ a net }\net{x}\text{ in } X:x_\alpha\longrightarrow x\}.
        \end{gather}
        This implies that the topology on $X$ is completely determined by the convergence of nets {\cref{set:net}}.
    \end{property}

    \newdef{Fr\'echet--Urysohn space}{\index{Fr\'echet--Urysohn space}\index{sequential!space}
        A topological space for which the closure of every subset is equal to its sequential closure, i.e.~the subset obtained as in \cref{topology:closure}, but with nets replaced by sequences.

        Fr\'echet--Urysohn spaces form an important subclass of \textbf{sequential spaces}, i.e.~topological spaces where the topology is uniquely determined by the convergence of sequences: a subset of a sequential space is closed if and only if every convergent sequence converges to a point in the set.
    }
    The following property is of great practical importance.
    \begin{property}\label{topology:first_countable_sequential}
        Every first-countable space is Fr\'echet-Urysohn and, therefore, only convergent sequences have to be considerd in these spaces.
    \end{property}

    \newdef{Germ}{\index{germ}\label{topology:germ}
        Let $X$ be a topological space and let $Y$ be a set. Consider two functions $f,g:X\rightarrow Y$. If there exists a neighbourhood $N$ of a point $x\in X$ such that \[f(u) = g(u)\qquad\qquad\forall u\in N\,,\] this property defines an equivalence relation denoted by $f\sim_x g$ and the equivalence classes are called germs at $x$.
    }

\subsection{Separation axioms}\index{separation axioms}

    \newdef{Irreducible space}{\index{irreducible!space}
        A topological space that is not the union of two proper closed subsets or, equivalently, if the intersection of two nonempty open subsets is again nonempty.
    }

    \newdef{$T_0$-space}{\index{distinguishable}\index{Kolmogorov!topology}
        A topological space such that for every two distinct points at least one of them has a neighbourhood not containing the other, i.e.~they are \textbf{topologically distinguishable}. $T_0$-spaces are also said to carry a \textbf{Kolmogorov topology}.
    }

    \newdef{$T_1$-space}{\index{separated}\index{Fr\'echet!topology}
        A topological space such that for every two distinct points $x,y$ there exists neighbourhood $N,N'$ of $x$ and $y$ respectively such that $y\not\in N$ and $x\not\in N'$, i.e.~they can topologically \textbf{separated}. $T_1$-spaces are also said to carry a \textbf{Fr\'echet topology} (not to be confused with \textit{Fr\'echet spaces} from functional analysis).
    }

    \newdef{Hausdorff space}{\index{Hausdorff!space}\label{topology:hausdorff}
        A topological space $X$ is called a Hausdorff space or $T_2$-space if it satisfies the following condition:
        \begin{gather}
            \forall x,y\in X:\exists\text{ neighbourhoods }N\ni x,N'\ni y:N\cap N'=\emptyset\,.
        \end{gather}
        The points are said to be \textbf{separated by neighbourhoods}. It can be shown that this definition is equivalent to requiring that the diagonal $\Delta_X$ is closed in the product space $X\times X$.
    }
    \begin{property}[Closed points]
        Every singleton and, by extension, every finite subset is closed in a Hausdorff space.
    \end{property}

    \newdef{Urysohn space}{\index{Urysohn!space}
        A topological space is called an Urysohn space or $T_{2\nicefrac{1}{2}}$-space if every two distinct points can be separated by closed neighbourhoods.
    }

    \newdef{Regular space}{\index{regular}\label{topology:regular}
        A topological space such that for every closed subset $V$ and every point $x\not\in V$ there exist disjoint open subsets $U,U'$ such that $x\in U$ and $V\subset U'$.
    }
    \newdef{$T_3$-space}{
        A space that is both regular and $T_0$.
    }

    \newdef{Normal space}{\index{normal}\label{topology:normal}
        A topological space such that every two closed subsets have disjoint neighbourhoods.
    }
    \newdef{$T_4$-space}{
        A space that is both normal and $T_1$.
    }

    \begin{property}[Nesting of axioms]
        A space satisfying the separation axiom $T_k$ also satisfies all separation axioms $T_{i\leq k}$.
    \end{property}

\subsection{Convergence}

    \newdef{Convergence}{\index{convergence}
        A sequence $\seq{x}$ in $X$ is said to converge to a point $x\in X$ if
        \begin{gather}
            \forall\text{ neighbourhoods }U\text{ of }x:\exists N\in\mathbb{N}_0:\forall n>N:x_n\in U\,.
        \end{gather}
    }

    \begin{property}[Uniqueness]\label{topology:hausdorff_limit}
        The limit of a convergent sequence in a Hausdorff space is unique.
    \end{property}

    \begin{property}[Subsequences]
        Every subsequence of a convergent sequence converges to the same point.
    \end{property}

\section{Morphisms}
\subsection{Continuity}

    \newdef{Continuous function}{\index{continuity}
        A function between topological spaces for which the preimage of every open set is also open.
    }
    \begin{notation}
        \nomenclature[S_Cont]{$C(X,Y)$}{set of continuous functions between two topological spaces $X$ and $Y$}
        The set of all continuous functions between two topological spaces $X,Y$ is often denoted by $C(X,Y)$.
    \end{notation}

    \newdef{Initial topology}{\index{topology!initial}\label{topology:initial_topology}
        Consider a collection of functions $\{f_i:X\rightarrow Y_i\}_{i\in I}$ between topological spaces. The initial topology on $X$ with respect to this family is the coarsest topology on $X$ for which all maps $f_i$ are continuous.
    }
    \newdef{Final topology}{\index{topology!final}
        Consider a collection of functions $\{f_i:Y_i\rightarrow X\}_{i\in I}$ between topological spaces. The final topology on $X$ with respect to this family is the finest topology on $X$ for which all maps $f_i$ are continuous.
    }

    \begin{property}[Sequantial continuity]\label{topology:sequential_continuity}
        Consider a function $f:X\rightarrow Y$ of topologial spaces, where $X$ is first-countable. The following statements are equivalent:
        \begin{itemize}
            \item $f$ is continuous.
            \item The sequence $(f(x_n))_{n\in\mathbb{N}}$ converges to $f(a)\in Y$ whenever the sequence $\seq{x}$ converges to $a\in X$.
        \end{itemize}
    \end{property}
    \begin{result}
       If the space $Y$ in the previous theorem is Hausdorff, the limit is irrelevant since it is unique by \cref{topology:hausdorff_limit} above.
    \end{result}
    \begin{remark}
        If the space $X$ is not first-countable, one has to consider the convergence of nets (\cref{set:net}).
    \end{remark}

    \begin{theorem}[Urysohn's lemma]\index{Urysohn!lemma}\label{topology:urysohns_lemma}
        A topological space X is normal if and only if every two closed disjoint subsets $A, B\subset X$ can be separated by a continuous function $f:X\rightarrow [0,1]$, i.e.~$\forall a\in A,b\in B$ there exists a continuous function $f:X\rightarrow [0,1]$ such that
        \begin{gather}
            f(a) = 0\qquad\text{and}\qquad f(b) = 1\,.
        \end{gather}
    \end{theorem}
    The following, seemingly unrelated, theorem is actually equivalent to Urysohn's lemma.
    \begin{theorem}[Tietze extension theorem]\index{Tietze extension theorem}
        Consider a continuous function $f:V\rightarrow\mathbb{R}$, where $V$ is a closed subset of a normal space $X$. There exists a continuous function $F:X\rightarrow\mathbb{R}$ such that $\forall x\in V:F(x) = f(x)$. Furthermore, if the function $f$ is bounded, then $F$ can be chosen to be bounded by the same number.
    \end{theorem}

    \newdef{Limit superior and inferior}{\index{limit}\label{topology:limit_superior}
        Consider a continuous function $f:X\rightarrow\mathbb{R}$. The limit inferior and superior of $f$ at $x_0\in(X,\tau)$ are defined as follows:
        \begin{align}
            \liminf_{x\rightarrow x_0}f(x) &:= \sup\bigl\{\inf\{f(x)\mid x\in U\backslash\{x_0\}\}\bigm\vert U\in\tau,x_0\in U,U\backslash\{x_0\}\neq\emptyset\}\bigr\}\,,\\
            \limsup_{x\rightarrow x_0}f(x) &:= \inf\bigl\{\sup\{f(x)\mid x\in U\backslash\{x_0\}\}\bigm\vert U\in\tau,x_0\in U,U\backslash\{x_0\}\neq\emptyset\}\bigr\}\,.
        \end{align}
        For functions on a subspace $Y\subset X$ one should look at the values $\{f(x)\mid x\in Y\cap U\backslash\{x_0\}\}$. For the definition to make sense, $x_0$ should be a limit point of $Y$.
    }
    \newdef{Semicontinuity}{\index{semi-!continuity}\label{topology:semicontinuity}
        Consider a function $f:X\rightarrow\overline{\mathbb{R}}$ on a topological space. It is said to be \textbf{upper semicontinuous} at a point $x_0\in X$ if for all $y\in\mathbb{R}$ with $y>f(x_0)$ there exists a neighbourhood of $x_0$ such that $f(x)<y$ for all $x\in U$. Analogously, the function is said to be \textbf{lower semicontinuous} at a point $x_0\in X$ if for all $y\in\mathbb{R}$ with $y<f(x_0)$ there exists a neighbourhood of $x_0$ such that $f(x)>y$ for all $x\in U$.

        These definitions can be expressed more succintly as follows:
        \begin{itemize}
            \item Upper semicontinuous: $\limsup_{x\rightarrow x_0}f(x)\leq f(x_0)$ for all $x_0\in X$, and
            \item Lower semicontinuous: $\liminf_{x\rightarrow x_0}f(x)\geq f(x_0)$ for all $x_0\in X$.
        \end{itemize}
    }

\subsection{Homeomorphisms}

    \newdef{Homeomorphism}{\index{homeomorphism}
        A bijection $f$ such that both $f$ and $f^{-1}$ are continuous.
    }

    \newdef{Embedding}{\index{embedding}\label{topology:embedding}
        A continuous function that is a homeomorphism onto its image.
    }
    \newdef{Local homeomorphism}{\index{local!homeomorphism}\index{etale!morphism}\label{topology:etale_morphism}
        A continuous function $f:X\rightarrow Y$ is a local homeomorphism if every point $x\in X$ has an open neighbourhood $U$ such that $f|_U$ is an embedding. Local homeomorphisms are also called \textbf{\'etale morphisms}.
    }

    \newdef{Covering space}{\index{covering!space}\label{topology:covering_space}
        Consider two topological spaces $C,X$ and a continuous surjection $p:C\rightarrow X$. $C$ is said to be a covering space of $X$ (and $p$ is called a \textbf{covering map}) if for all points $x\in X$ there exists an open neighbourhood $U$ of $x$ such that $p^{-1}(U)$ can be written as a disjoint union of open sets in $C$ where every set is mapped homeomorphically onto $U$. The neighbourhoods $U$ are sometimes said to be \textbf{evenly covered}.
    }
    \begin{notation}
        Because the covering map $p:C\rightarrow M$ is surjective, the space $M$ can be left implicit. Therefore, covering spaces are often just denoted by the couple $(C,p)$.
    \end{notation}

    \newdef{Covering transformation}{\label{topology:covering_transformation}
        Consider two covering spaces $(C,p)$ and $(C',p')$. A continuous function $f:C\rightarrow C'$ is called a covering transformation if $p'\circ f=p$.
    }

    \newdef{Deck transformation}{\index{deck transformation}\label{topology:deck_transformation}
        Let $p:C\rightarrow X$ be a covering map. The automorphism group of $(C,p)$ in the category of covering spaces (over $X$) is given by all homeomorphisms $\varphi$ satisfying $p\circ\varphi=p$. These automorphisms are called deck transformations.
    }

    \newdef{\'Etale space}{\index{etale!space}\index{stalk}\label{topology:etale_space}
        Let $X$ be a topological space. A topological space $Y$ is called an \'etale space over $X$ if there exists a surjective local homeomorphism $\pi:Y\rightarrow X$. The preimage $\pi^{-1}(x)$ of a point $x\in X$ is called the \textbf{stalk} of $\pi$ over $x$.
    }
    \begin{example}
        Every covering space is an \'etale space.
    \end{example}

    \newdef{\difficult{Pseudogroup}}{\index{pseudo-!group}\label{topology:pseudogroup}
        Let $(X,\tau)$ be a topological space. A pseudogroup is a collection $\mathcal{G}$ of homeomorphisms $\phi:U\rightarrow V$ between open subsets of $X$ such that:
        \begin{enumerate}
            \item $\mathbbm{1}_U\in\mathcal{G}$ for all $U\in\tau$.
            \item If $\phi\in\mathcal{G}$, then $\phi^{-1}\in\mathcal{G}$.
            \item If $V\subset U\in\tau$, then $\phi|_V\in\mathcal{G}$.
            \item If $U=\bigcup_{i\in I}U_i$ and $\phi|_{U_i}:U_i\rightarrow V$ is an element of $\mathcal{G}$ for all $i\in I$, then $\phi\in\mathcal{G}$.
            \item If $\phi:U\rightarrow V$ and $\psi:U'\rightarrow V'$ are elements of $\mathcal{G}$ and $V\cap U'\neq\emptyset$, then $\psi\circ\phi|_{\phi^{-1}(V\cap U')}\in\mathcal{G}$.
        \end{enumerate}
    }

\section{Associated constructions}

    \begin{construct}[Product topology]\index{product!topology}\index{Tychonoff!topology}\label{topology:tychonoff_topology}
        First, consider the case with only a finite number of spaces $\{(X_i,\tau_i)\}_{i\in I}$. The Cartesian product $X:=\prod_{i\in I}X_i$ can be turned into a topological space by equipping it with the topology generated by the following basis:
        \begin{gather}
            \mathcal{B} := \left\{\prod_{i\in I}U_i\,\middle\vert\,U_i\in\tau_i\right\}.
        \end{gather}
        In the general case the topology can be defined using the canonical projections $\pi_i:X\rightarrow X_i$. The general product topology, called the \textbf{Tychonoff topology}, is the initial topology with respect to the projections $\pi_i$.
    \end{construct}

    \begin{construct}[Disjoint union]\index{disjoint union}\label{topology:disjoint_union}
        Let $\{(X_i,\tau_i)\}_{i\in I}$ be a family of topological spaces and consider the disjoint union
        \begin{gather}
            X := \bigsqcup_{i\in I}X_i
        \end{gather}
        together with the canonical inclusion maps $\phi_i:X_i\rightarrow X:x_i\mapsto(i,x_i)$. The set $X$ can be turned into a topological space by equipping it with the following topology:
        \begin{gather}
            \tau_X := \bigl\{U\subseteq X\bigm\vert\forall i\in I:\phi_i^{-1}(U)\in\tau_i\bigr\}\,.
        \end{gather}
    \end{construct}

    \begin{construct}[Quotient space]\index{quotient!space}\label{topology:quotient_space}
        Consider a topological space $(X,\tau)$ and a subset $Y\subseteq X$. The quotient $X/Y$ is defined as the set $X\backslash Y\sqcup\{\ast\}$ where the point $\ast$ can be regarded as the result of identifying all points in $Y$. This canonically turns the quotient space into a pointed space.

        Let $\pi$ be the canonical projection $X\rightarrow X/Y$. The quotient space can be turned into a topological space by equipping it with the following topology:
        \begin{gather}
            \tau_Q := \bigl\{U\subseteq X/Y\bigm\vert\pi^{-1}(U)\in\tau\bigr\}\,.
        \end{gather}
    \end{construct}
    \begin{remark}[Degenerate quotient]
        For the degenerate case $Y=\emptyset$ one can also apply the above definition. However, this has the awkward effect that it adjoins a new point to the space $X$ instead of a collapsing it:
        \begin{gather}
            \label{topology:empty_quotient}
            X/\emptyset = X\sqcup\ast\,.
        \end{gather}
    \end{remark}

    \begin{construct}[Wedge sum]\index{wedge!sum}
        Consider two pointed spaces $(X,x_0),(Y,y_0)$. The wedge sum $X\vee Y$ is defined as the quotient of the disjoint union $X\sqcup Y$ obtained by identifying the basepoints $x_0\sim y_0$.
    \end{construct}
    \newdef{Smash product}{\index{smash product}
        Consider two pointed topological spaces $(X,x_0),(Y,y_0)$. The smash product $X\wedge Y$ is defined as the quotient
        \begin{gather}
            X\wedge Y := (X\times Y)/(X\vee Y)\,,
        \end{gather}
        where $X\vee Y$ sits inside the product as the union of $X\times\{y_0\}$ and $\{x_0\}\times Y$.
    }

    \begin{construct}[Suspension]\index{suspension}\label{topology:suspension}
        Let $X$ be a topological space. The suspension of $X$ is defined as the following quotient space:
        \begin{gather}
            SX := (X\times [0,1])/\bigl\{(x,0)\sim (y,0)\text{ and }(x,1)\sim (y,1)\bigm\vert x,y\in X\bigr\}\,.
        \end{gather}
        By the remark about degenerate quotients the suspension of the empty set is in fact not empty, but equal to the two-point space $S^0$.

        An often more interesting construction is the \textbf{reduced suspension} $\Sigma X$. This is obtained by taking the ordinary suspension $SX$ of a pointed space $(X,x_0)$ and identifying all copies of $x_0$:
        \begin{gather}
            \Sigma X := SX/(x_0\times[0,1])\,.
        \end{gather}
        An equivalent definition of the reduced suspension can be given in terms of the smash product:
        \begin{gather}
            \Sigma X\cong X\wedge S^1\,.
        \end{gather}
    \end{construct}
    \begin{example}[Spheres]\label{topology:sphere_suspension}
        Up to homeomorphism, the spheres are related by (reduced) suspensions:
        \begin{gather}
            SS^n\cong S^{n+1}\cong\Sigma S^n\,.
        \end{gather}
        If one identifies the empty set with the $(-1)$-sphere, this relation can be continued to the case $n=-1$.
    \end{example}

    \begin{construct}[Attaching space]\index{attaching space}\label{topology:attaching_space}
        Let $X,Y$ be two topological spaces and consider a subspace $A\subseteq X$. For every continuous function $f:A\rightarrow Y$, called the \textbf{attaching map}, one can construct the attaching space or \textbf{adjunction space} $X\cup_f Y$ in the following way:
        \begin{gather}
            X\cup_f Y := (X\sqcup Y)/\{A\sim f(A)\}\,.
        \end{gather}
        In categorical terms, it is the pushout (\cref{cat:pushout}) in $\mathbf{Top}$ of the inclusion $\iota:A\hookrightarrow X$ along $f:A\rightarrow Y$.
    \end{construct}

    \begin{construct}[Join]\index{join}\label{topology:join}
        Let $\{X_i\}_{i\leq n}$ be a finite collection of topological spaces. The join $X:=X_1\circ\cdots\circ X_n$ is defined as follows. Every point of $X$ is defined by the following data:
        \begin{enumerate}
            \item An element of the standard $n$-simplex (\cref{topology:simplex}), i.e.~an $n$-tuple of nonnegative numbers $\{t_i\}_{i\leq n}$ satisfying $\sum_it_i=1$.
            \item For each index $i$ such that $t_i\neq 0$, a point $x_i\in X_i$.
        \end{enumerate}
        This point in $X$ is denoted by $t_1x_1\oplus\cdots\oplus t_nx_n$.

        In the case of two spaces, there exists a more intuitive construction. Let $X,Y$ be two topological spaces. The join $X\circ Y$ is equal to the quotient space $(X\times Y\times[0,1])/\sim$, where the equivalence relation is defined as follows:
        \begin{itemize}
            \item For all $x\in X$ and $y,y'\in Y$: $(x,y,0)\sim(x,y',0)$.
            \item For all $x,x'\in X$ and $y\in Y$: $(x,y,1)\sim(x',y,1)$.
        \end{itemize}
        This can be interpreted as collapsing one end of the cylinder $(X\times Y)\times[0,1]$ to $X$ and the other end to $Y$.
    \end{construct}
    \begin{property}[\difficult{Monoidal structure}]
        The join induces a monoidal structure on the category $\mathbf{Top}$, where the tensor unit is given by the empty space $\emptyset$.
    \end{property}

    \newdef{Profinite group}{\index{group!profinite}\index{Stone space}\label{topology:profinite_group}
        An inverse limit (\cref{set:inverse_limit}) of finite groups, equipped with the induced product topology. These can equivalently be defined as \textit{Stonean} topological groups.\footnote{\textbf{Stone spaces} are actually pro-objects in the category of topological spaces with respect to discrete spaces. Equivalently, these are totally-disconnected compact Hausdorff spaces (see the next sections).}
    }

\section{Connected spaces}

    \newdef{Connected space}{\index{connected!space}\label{topology:connected}
        A topological space that cannot be written as the disjoint union of two nonempty open sets. Equivalently, a space is said to be connected if the only clopen sets are the empty set and the space itself.
    }

    \begin{property}[Locally constant implies constant]
        Let $X$ be a connected space and let $f$ be a function on $X$. If $f$ is locally constant, i.e.~for every $x\in X$ there exists a neighbourhood U on which $f$ is constant, then $f$ is constant on all of $X$.
    \end{property}

    \begin{theorem}[Intermediate value theorem]\index{intermediate value theorem}\label{topology:intermediate_value_theorem}
        Let $X$ be a connected space and let $f:X\rightarrow\mathbb{R}$ be a continuous function. If $a,b\in f(X)$, then $\forall c\in\ ]a,b[\ :c\in f(X)$.
    \end{theorem}

    \newdef{Path-connected space\protect\footnotemark}{\index{arc!connected|see{path-connected}}\index{path!connected}
        \footnotetext{A similar notion is that of \textbf{arcwise-connectedness} where the function $\varphi$ is required to be a homeomorphism.}
        Let $X$ be a topological space. If for every two points $x,y\in X$ there exists a continuous function $\varphi:[0,1]\rightarrow X$ (i.e.~a \textbf{path}) such that $\varphi(0)=x$ and $\varphi(1)=y$, the space is said to be path-connected.
    }

    \begin{property}[Path-connected implies connected]
        Every path-connected space is connected. The converse does not hold. A connected and locally path-connected space is path-connected.
    \end{property}

    \begin{remark}[Connected components]\label{topology:connected_components}
        (Path-)connectedness defines an equivalence relation on the space $X$. The equivalence classes are closed in $X$ and form a cover of $X$. The set of path components of $X$ is often denoted by $\pi_0(X)$.
    \end{remark}

\section{Compact spaces}\label{section:compact}
\subsection{Compactness}

    \newdef{Sequentially compact space}{
        A topological space in which every sequence has a convergent subsequence (the sequence itself does not have to be convergent).
    }

    \begin{property}[First-countable spaces]
        A first-countable space is sequentially compact if and only if every countable open cover has a finite subcover.
    \end{property}

    \newdef{Lindel\"of space}{\index{Lindel\"of!space}
        A space for which every open cover has a countable subcover.
    }
    \begin{property}
        Every second-countable space is a Lindel\"of space.
    \end{property}

    \newdef{Finite intersection property}{\index{finite!intersection property}
        \nomenclature[A_FIP]{FIP}{finite intersection property}
        A collection $\mathcal{F}\subseteq2^X$ of subsets has the finite intersection property (FIP) if
        \begin{gather}
            \bigcap_{V\in\mathcal{F}'}V\neq\emptyset
        \end{gather}
        for all finite $\mathcal{F}'\subset\mathcal{F}$.
    }

    \newdef{Locally finite cover}{
        An open cover of a topological space such that every point has a neighbourhood that intersects only finitely many sets in the cover.
    }

    \newdef{Paracompact space}{\index{para-!compact}\label{topology:paracompact}
        A topological space for which every open cover has a locally finite open refinement.
    }
    \begin{property}
        Every paracompact Hausdorff space is normal.
    \end{property}

    \newdef{Partition of unity}{\index{partition!of unity}\label{topology:partition_of_unity}
        A collection $\{f_i:X\rightarrow[0,1]\}_{i\in I}$ of continuous functions such that for every $x\in X$ the following conditions hold:
        \begin{enumerate}
            \item\textbf{Locally finite}: For every neighbourhood $U$ of $x$, the set $\{f_i\mid\mathrm{supp}f_i\cap U\neq\emptyset\}$ is finite.
            \item\textbf{Normalization}: $\sum_{i\in I}f_i(x)=1$.
        \end{enumerate}
        Consider an open cover $\{V_i\}_{i\in I}$ of $X$. If there exists a partition of unity, also indexed by $I$, such that $\text{supp}(\varphi_i)\subseteq U_i$, this partition of unity is said to be \textbf{subordinate} to the given cover.
    }

    \begin{property}[Hausdorff spaces]\label{topology:paracompact_partition_unity}
        A paracompact space is Hausdorff if and only if it admits a partition of unity subordinate to any open cover.
    \end{property}

    \newdef{Numerable open cover}{\index{numerable}\label{topology:numerable_cover}
        An open cover of a topological space is said to be numerable if the space admits a partition of unity subordinate to the given cover.
    }

    \newdef{Compact space}{\index{compact}
        A topological space for which every open cover has a finite subcover.
    }

    \begin{theorem}[Heine--Borel\footnotemark]\index{Heine--Borel}\index{Borel--Lebesgue}\label{topology:heine_borel}
        \footnotetext{Also called the \textbf{Borel--Lebesgue} theorem.}
        If a topological space $X$ is sequentially compact and second-countable, every open cover has a finite subcover and, therefore, $X$ is compact.
    \end{theorem}
    \begin{result}[Real numbers]
        A subset of $\mathbb{R}^n$ is compact if and only if it is closed and bounded.
    \end{result}

    \begin{theorem}[Tychonoff's theorem]\index{Tychonoff!compactness theorem}
        Any product of compact topological spaces is compact under the (Tychonoff) product topology (\cref{topology:tychonoff_topology}).
    \end{theorem}

    \newdef{Relatively compact space}{\label{topology:relatively_compact}
        A topological space for which its closure is compact.
    }

    \newdef{Locally compact space}{
        A topological space in which every point has a compact neighbourhood.
    }

    \begin{theorem}[Dini]\index{Dini}
        Let $(X,\tau)$ be a compact space and let $\seq{f}$ be a monotone sequence of continuous functions $f_n:X\rightarrow\mathbb{R}$. If $\seq{f}$ converges pointwise to a continuous function, the convergence is uniform.
    \end{theorem}

    \newdef{$\omega$-bounded space}{\index{bounded!space}
        A topological space in which the closure of every countable subset is compact.
    }

    \newdef{Compact-open topology}{\index{topology!compact-open}\label{topology:compact_open_topology}
        Consider the mapping space $C(X,Y)$ between two topological spaces. This space is often endowed with a topology generated by the subbasis of subsets of the form
        \begin{gather}
            U^K := \{f:X\rightarrow Y\mid K\text{ compact}, U\text{ open and } f(K)\subseteq U\}\,.
        \end{gather}
    }
    \begin{property}[Internal hom]\index{mapping!space}\label{topology:internal_hom}
        Consider two topological spaces $X,Y$ with $X$ locally compact and equip the mapping space $C(X,Y)$ with the compact-open topology. The following relation is satisfied for all topological spaces $Z$:
        \begin{gather}
            C(Z\times X,Y)\cong C\bigl(Z,C(X,Y)\bigr)\,,
        \end{gather}
        i.e.~the mapping space $C(X,Y)$ is an internal hom (\cref{cat:internal_hom}) in the category $\mathbf{Top}$ and, because the topological product is the product in $\mathbf{Top}$, $C(X,Y)$ is even an exponential object (\cref{cat:exponential_object}). For this reason the mapping spaces $C(X,Y)$ are also sometimes denoted by $Y^X$.
    \end{property}

\subsection{Compactifications}

    \newdef{Dense}{\index{dense}
        A subset $V\subseteq X$ is said to be dense in a topological space $X$ if $\overline{V}=X$.
    }

    \newdef{Generic point}{\index{generic!point}\label{topology:generic_point}
        A point whose closure is the whole space.
    }

    \newdef{Separable space}{\index{separable!space}\label{topology:separable}
        A topological space that contains a countable, dense subset.
    }
    \begin{property}
        Every second-countable space is separable.
    \end{property}

    \newdef{Compactification}{\index{compactification}
        A compact topological space $(X',\tau')$ is a compactification of a topological space $(X,\tau)$ if $X$ is a dense subspace of $X'$.
    }

    \begin{example}
        Standard examples of compactifications are the extended real line $\mathbb{R}\cup\{-\infty,+\infty\}$ and the extended complex plane $\mathbb{C}\cup\{\infty\}$ for the real line and the complex plane, respectively.
    \end{example}
    \begin{remark*}
        It is important to note that compactifications are not necessarily unique.
    \end{remark*}

    \newdef{One-point compactification}{\index{Alexandrov!compactification}\index{compactification}\label{topology:alexandrov_compactification}
        Let $X$ be a Hausdorff space. A one-point compactification or \textbf{Alexandrov compactification} is a compactification $\widehat{X}$ such that $\widehat{X}\backslash X$ is a singleton.
    }
    \begin{example}[Real line]\index{projection!stereographic}
        The classic example of a (one-point) compactification is that of the real line. By adjoining the points $\pm\infty$ and identifying them, the circle $S^1$ is obtained. In general one can obtain the $n$-dimensional sphere $S^n$ as the one-point compactification of $\mathbb{R}^n$. This can be regarded as an inverse \textit{stereographic projection}.
    \end{example}

\section{Uniform spaces}

    \newdef{Uniform structure}{\index{uniform!space}
        A uniform structure on a set $X$ consists of a collection $\mathfrak{U}$ of subsets $U\subseteq X\times X$ that satisfy the following properties:
        \begin{enumerate}
            \item\textbf{Monotonicity}: If $U\in\mathfrak{U}$ and $U\subset V$, then $V\in\mathfrak{U}$.
            \item\textbf{Intersections}: If $U, V\in\mathfrak{U}$, then $U\cap V\in\mathfrak{U}$.
            \item\textbf{Diagonals}: If $U\in\mathfrak{U}$, then $\Delta_X\subset U$.
            \item\textbf{Halving}: If $U\in\mathfrak{U}$, there exists $V\in\mathfrak{U}$ such that $V\circ V\subseteq U$.
            \item\textbf{Transposition}: If $U\in\mathfrak{U}$, then $U^t\in\mathfrak{U}$.
        \end{enumerate}
        $U^t$ denotes the converse (\cref{set:converse}) of $U$ and $\circ$ is the relational composition (\cref{set:relational_composition}) of $V$ and $V$. The elements of the uniformity $\mathfrak{U}$ are called \textbf{entourages}. If $(x,y)\in U$ for some entourage $U\in\mathfrak{U}$, then $x$ and $y$ are said to be \textbf{$U$-close}.
    }
    \remark{The first three conditions imply that a uniform structure is in particular a filter (\cref{set:filter}).}

    @@ COMPLETE (Bourbaki) @@

\section{Bornological spaces}

    \newdef{Bornology}{\index{bornology}\label{topology:bornology}
        A bornology on a set $X$ is a collection $\mathfrak{b}\subset P(X)$ such that:
        \begin{enumerate}
            \item\textbf{Singletons}: $\forall x\in X:\{x\}\in\mathfrak{b}$.
            \item\textbf{Monotonicity}: $A\in\mathfrak{b},B\subset A\implies B\in\mathfrak{b}$.
            \item\textbf{Union}: $A,B\in\mathfrak{b}\implies A\cup B\in\mathfrak{b}$.
        \end{enumerate}
        The sets in $\mathfrak{b}$ are said to be \textbf{bounded} and the pair $(X,\mathfrak{b})$ is called a \textbf{bornological space}.
    }

    @@ COMPLETE @@

\section{\texorpdfstring{Locales $\clubsuit$}{Locales}}

    \begin{property}[Opens form a frame]
        Consider the poset $\mathbf{Open}(X)$ of opens of a topological space $X$. This set is closed under finite intersections (limits) and arbitrary unions (colimits). Furthermore, arbitrary unions distribute over finite intersections:
        \begin{gather}
            V\cap\left(\bigcup_{i\in I}U_i\right) = \bigcup_{i\in I}\left(V\cap U_i\right)\,.
        \end{gather}
        This implies that the poset $\mathbf{Open}(X)$ is a frame (\cref{set:frame}).
    \end{property}

    \newdef{Locale}{\index{locale}\label{topology:locale}
        The previous property can be used to generalize the notion of topological spaces to include ``pointless spaces''. Let \textbf{Frame} denote the category of frames together with frame homomorphisms. The category of locales is defined as its opposite category:
        \begin{gather}
            \mathbf{Loc} := \mathbf{Frame}^{op}\,.
        \end{gather}
    }
    \begin{construct}[From locale to topological space]
        There exists an adjunction
        \begin{gather}
            \mathbf{Loc}\adj{\iota}{\mathrm{Point}}\mathbf{Top}\,,
        \end{gather}
        where the right adjoint is defined as follows. Let $L$ be a locale. For a topological space the points are given by continuous functions $\ast\rightarrow X$ and, hence, by frame morphisms $\mathbf{Open}(X)\rightarrow1\equiv\Omega_\mathrm{Frame}=\{0, 1\}$. Generalizing this to locales, one defines the set of points of $L$ as the $\Omega_\mathrm{Loc}$-elements: \[\mathrm{Point}(L) := \mathbf{Loc}(1,L)\,.\] This set can be given a topology by declaring for every $U\in L$ the set $\{p\in\mathrm{Point}(L)\mid p^{-1}(U) = 1\}$ to be open.
    \end{construct}

    \newdef{Sober space}{\index{sober}\label{topology:sober_space}
        A topological space $X$ such that the map $X\rightarrow\mathrm{Point}(X)$ is a homeomorphism, i.e.~the points of $X$ are precisely determined by its frame of opens. Equivalently, a topological space such that every irreducible closed subset is the closure of a unique point, i.e.~has a unique generic point (\cref{topology:generic_point}). Important examples are Hausdorff spaces.
    }

    @@ COMPLETE @@