\chapter{Electricity and Magnetism}

    \newdef{Capacitance}{\index{capacitance}\label{em:capacitance}
        The capacitance is a (geometrical) value that reflects the amount of charge an object can store:
        \begin{gather}
            C := \frac{q}{U}.
        \end{gather}
    }

\section{Resistance}

    \newdef{Drift velocity}{
        The average speed of independent charge carriers is the drift velocity $\vector{v}_d$. It is important to remark that $v_d$ is not equal to the propagation speed of the electric signal, which is the result of particle collisions. In fact, it is several orders of magnitude smaller.
    }

    \newformula{Mobility}{\label{em:mobility}
        \begin{gather}
            \mu := \frac{v_d}{E},
        \end{gather}
        where $E$ is the external electric field.
    }
    \newformula{Conductivity}{\label{em:conductivity}
        \begin{gather}
            \sigma := nq\mu
        \end{gather}
    }
    \newformula{Resistivity}{\label{em:resistivity}
        \begin{gather}
            \rho := \frac{1}{\sigma}
        \end{gather}
    }

    \begin{formula}[Pouillet's law]\index{Pouillet}\label{em:pouillet}
        The resistance of a wire is given by the following formula:
        \begin{gather}
            R = \frac{l}{A}\rho,
        \end{gather}
        where
        \begin{itemize}
            \item $\rho$ is the resistivity of the material,
            \item $l$ is the length of the resistor, and
            \item $A$ is the cross-sectional area of the resistor.
        \end{itemize}
    \end{formula}

\section{Ohm's law}

    \newformula{Free current density}{\label{em:free_current_density}
        The current density generated by free charges is given by
        \begin{gather}
            \vector{J} = nq\vector{v}_d.
        \end{gather}
    }

    \newformula{Ohm's law}{\index{Ohm}\label{em:ohms_law}
        \begin{gather}
            \vector{J} = \sigma\cdot\vector{E},
        \end{gather}
        where $\sigma$ is the conductivity tensor. (Compare this to equations \ref{em:mobility} and \ref{em:conductivity} for 1D systems.)
    }

    \newformula{Ohm's law in wires}{\label{em:ohms_law_linear}
        The following formula can be found by combining equations \ref{em:conductivity}, \ref{em:resistivity} and \ref{em:ohms_law} and by assuming that the conductivity tensor is a scalar (this follows from the isotropic behaviour of most resistors):
        \begin{gather}
            U = RI.
        \end{gather}
        ?? CHECK THIS ARGUMENT ??
    }

\section{Electric dipoles}

    \begin{formula}[Electric dipole]\index{dipole}\label{em:dipole}
        \begin{gather}
            \vector{p} := q\vector{a}
        \end{gather}
        where
        \begin{itemize}
            \item $q$ is the positive charge, and
            \item $\vector{a}$ is the vector pointing from the negative charge to the positive charge.
        \end{itemize}
    \end{formula}

    \begin{formula}[Energy]
        If an electric dipole is placed in an external electric field, its potential energy is given by
        \begin{gather}
            \label{em:dipole_energy}
            U = -\vector{p}\cdot\vector{E}.
        \end{gather}
    \end{formula}

    \begin{formula}[Torque]
        If an electric dipole is placed in an external electric field, the torque on this system is given by
        \begin{gather}
            \label{em:dipole_torque}
            \vector{\tau} = \vector{p}\times\vector{E}.
        \end{gather}
    \end{formula}

\section{Magnetic fields}\index{magnetization}\index{induction}

    The \textbf{magnetizing field} $\vector{H}$ is the field generated by all external sources. When applying an external (magnetic) field, some materials will try to oppose this external influence. Similar to polarization in the case of electricity, one can define the \textbf{magnetization}:
    \begin{gather}
        \label{em:M}
        \vector{M} := \chi\vector{H},
    \end{gather}
    where $\chi$ is the magnetic susceptibility.

    The \textbf{magnetic induction} $\vector{B}$ is the field generated by both the external sources and the internal magnetization. It is only this field that can be measured. In vacuum, the following relation between the magnetic induction, the magnetizing field and the magnetization exists:
    \begin{gather}
        \label{em:B}
        \vector{B} = \mu_0\left(\vector{H} + \vector{M}\right).
    \end{gather}
    By combining the previous two formulas one obtains (this equation is only valid in linear media)
    \begin{gather}
        \label{em:B_with_only_H}
        \vector{B} = \mu_0\left(1 + \chi\right)\vector{H}.
    \end{gather}
    The proportionality constant in this formula is called the \textbf{magnetic permeability}:\index{permeability}
    \begin{gather}
        \label{em:relative_permeability}
        \mu := \mu_0(1 + \chi),
    \end{gather}
    where $\mu_0$ is the magnetic permeability of the vacuum. The factor $1+\chi$ is called the \textbf{relative permeability} and it is often denoted by $\mu_r$.

    \begin{remark}[Tensorial formulation]
        In anisotropic materials one has to use a tensorial formulation:
        \begin{align}
            \label{em:B_tensor}
            B_i &= \sum_j\mu_{ij}H_j\\
            \label{em:M_tensor}
            M_i &= \sum_j\chi_{ij}H_j.
        \end{align}
        In this case both $\mu$ and $\chi$ are $(1,1)$-tensors, i.e. linear maps.
    \end{remark}

\subsection{Electric charges in a magnetic field}

    \newformula{Gyroradius}{
        A charge $q$ moving in a magnetic field $B$ with velocity $\vector{v}$ will follow a circular path with radius
        \begin{gather}
            \label{em:gyroradius}
            r = \frac{mv_\perp}{|q|B}.
        \end{gather}
    }
    \newformula{Gyrofrequency\footnotemark}{\index{cyclotron}\index{Larmor frequency}
        \footnotetext{Also called the \textbf{Larmor} or \textbf{cyclotron frequency}.}
        \begin{gather}
            \label{em:gyrofrequency}
            \omega = \frac{|q|B}{m}
        \end{gather}
    }