\chapter{\difficult{Higher topos theory}}

    \minitoc

\section{Higher category theory}\label{cat:higher_category_theory}

    This section gives an introduction to the theory of higher categories, in particular $n$-categories for finite $n>0$. For the notion of $\infty$-categories, see \cref{section:infinity_categories}.

\subsection{\texorpdfstring{$n$-categories}{n-categories}}

    \newdef{$n$-category}{\index{n-category}
        A (strict) $n$-category consists of:
        \begin{itemize}
            \item objects (0-morphisms),
            \item 1-morphisms going between 0-morphisms,
            \item ...
            \item $n$-morphisms going between $(n-1)$-morphisms,
        \end{itemize}
        such that the composition of $k$-morphisms ($k\leq n$) is associative and satisfies the unit laws as required in an ordinary category. By generalizing this definition to arbitrary $n\in\mathbb{N}$, one can define the notion of a (strict) $\infty$-category.

        If one relaxes the associativity and unit laws up to higher coherent morphisms, one obtains the notion of a \textbf{weak $n$-category}. Explicit definitions for such categories have been constructed up to \textbf{tetracategories} $(n=4)$. However, this construction by \textit{Trimble} takes about 50 pages of diagrams.
    }
    \sremark{$n$-morphisms are also called \textbf{$n$-cells}. This makes their relation to topological spaces (and, in particular, simplicial spaces) more apparent.}

    \begin{example}
        The classical examples of a 1-category and 2-category are $\symbfsf{Set}$ and $\symbfsf{Cat}$, respectively.
    \end{example}

    \begin{property}[Composition in 2-categories]\index{Godement product}\index{interchange law}
        2-morphisms can be composed in two different ways:
        \begin{itemize}
            \item\textbf{Horizontal composition}:
                Consider two 2-morphisms $\alpha:f\Rightarrow g$ and $\beta:f'\Rightarrow g'$, where $f'\circ f$ and $g'\circ g$ are well-defined. These 2-morphisms can be composed as
                \begin{gather}
                    \beta\circ\alpha: f'\circ f\Rightarrow g'\circ g\,.
                \end{gather}
                This is sometimes called the \textbf{Godement product}.
            \item\textbf{Vertical composition}:
                Consider two 2-morphisms $\alpha:f\Rightarrow g$ and $\beta:g\Rightarrow h$, where $f,g$ and $h$ have the same domain and codomain. These 2-morphisms can be composed as
                \begin{gather}
                    \beta\cdot\alpha:f\Rightarrow h\,.
                \end{gather}
        \end{itemize}
        As a consistency condition, the horizontal and vertical composition are required to satisfy the following \textbf{interchange law}:
        \begin{gather}
            (\alpha\cdot\beta)\circ(\gamma\cdot\delta) = (\alpha\circ\gamma)\cdot(\beta\circ\delta).
        \end{gather}
    \end{property}
    \newdef{\texorpdfstring{$(n,r)$-category}{(n,r)-Category}}{
        A higher ($\infty$-)category for which
        \begin{itemize}
            \item all parallel $k$-morphisms with $k>n$ are equivalent and, hence, trivial.
            \item all $k$-morphisms with $k>r$ are invertible (or equivalences in the fully weak $\infty$-sense).
        \end{itemize}
    }

    \newdef{Weak inverse}{\index{weak!inverse}
        Let $\symbfsf{C}$ be a 2-category. A 1-morphism $f:x\rightarrow y$ is weakly invertible if there exist a 1-morphism $g:y\rightarrow x$ and 2-isomorphisms $g\circ f\Rightarrow\mathbbm{1}_x$ and $f\circ g\Rightarrow\mathbbm{1}_y$.
    }

    At this point, it should be obvious that the definition of a unit-counit adjunction (\cref{cat:unit_counit_adjunction}) can be generalized to general 2-categories:
    \newdef{Adjunction in 2-category}{
        Let $\symbfsf{C}$ be a 2-category. An adjunction in $\symbfsf{C}$ is a pair of 1-morphisms $F:x\rightarrow y$ and $G:y\rightarrow x$ together with 2-morphisms $\varepsilon:F\circ G\Rightarrow\mathbbm{1}_y$ and $\eta:\mathbbm{1}_x\Rightarrow G\circ F$ that satisfy the zig-zag identities.
    }

    By looking at the defining relations of duals in a rigid monoidal category (\cref{section:duality}), it should be clear that these are in fact the same as the defining relations of the unit and counit of an adjunction. This is a consequence of the fact that a 2-category with a single object can be regarded as a (strict) monoidal category where the composition in the 2-category becomes the tensor product in the monoidal category. Similarly, adjoint 1-morphisms in the 2-category become duals in the monoidal category. This is formalized as follows.
    \begin{property}[Monoidal categories]\index{monoidal!category}\index{delooping!of monoidal categories}\label{cat:monoidal_or_2}
        Consider a monoidal category $(\symbfsf{C},\otimes,\symbfsf{1})$. From this monoidal category, one can construct the so-called \textbf{delooping} (bicategory) $\symbfsf{BC}$ in the following way:
        \begin{itemize}
            \item There is a single object $\ast$.
            \item The 1-morphisms in $\symbfsf{BC}$ are the objects in $\symbfsf{C}$.
            \item The 2-morphisms in $\symbfsf{BC}$ are the morphisms in $\symbfsf{C}$.
            \item Horizontal composition in $\symbfsf{BC}$ is the tensor product in $\symbfsf{C}$.
            \item Vertical composition in $\symbfsf{BC}$ is composition in $\symbfsf{C}$.
        \end{itemize}
        Conversely, every 2-category with a single object comes from a monoidal category. Hence, the 2-category of (pointed) 2-categories with a single object and the 2-category of monoidal categories are equivalent. (This property and its generalizations are the content of the \textit{delooping hypothesis}.)

        In the same way one can deloop a braided monoidal category twice and find an identification with a one-object tricategory with one 1-morphism. However, this identification is not a trivial one as it makes use of the Eckmann--Hilton argument (\cref{cat:eckmann_hilton}) to identify different monoidal structures on this tricategory. (See also \cref{section:monoidal_n_cat}.)
    \end{property}

\subsection{\texorpdfstring{$n$-functors}{n-functors}}

    \newdef{2-functor}{\index{pseudo-!functor}\label{cat:pseudofunctor}
        A 2-functor $\func{F}{C}{D}$ (often called a \textbf{pseudofunctor}) is a morphism between bicategories. It consists of the following data:
        \begin{itemize}
            \item a function $F_0:\ob{C}\rightarrow\ob{D}$, and
            \item for every two objects $x,y\in\ob{C}$, a functor $F_{x,y}:\symbfsf{C}(x,y)\rightarrow\symbfsf{D}(Fx,Fy)$.
        \end{itemize}
        The function $F_0$ and the functors $F_{x,y}$ are also often denoted by $F$ by abuse of notation. This data is required to satisfy some coherence conditions. These are specified by the following data:
        \begin{enumerate}
            \item\textbf{Associator}: For every pair of composable 1-morphisms $f\circ g$ in $\mathrm{hom}(\symbfsf{C})$, a 2-isomorphism $\gamma_{f,g}:Ff\circ Fg\Rightarrow F(f\circ g)$ such that for every triple of composable morphisms $f\circ g\circ h$ in $\mathrm{hom}(\symbfsf{C})$, the following identity holds:
            \begin{gather}
                \gamma_{f\circ g,h}\circ(\gamma_{f,g}\cdot\mathbbm{1}_{Fh}) = \gamma_{f,g\circ h}\circ(\mathbbm{1}_{Ff}\cdot\gamma_{g,h}).
            \end{gather}
            \item\textbf{Unitor}: For every object $x\in\ob{C}$, a 2-isomorphism $\iota_x:\mathbbm{1}_{Fx}\Rightarrow F\mathbbm{1}_x$ such that for every morphism $f:x\rightarrow y$ in $\mathrm{hom}(\symbfsf{C})$, the following identities hold:
            \begin{gather}
                \begin{aligned}
                    \iota_y\cdot\mathbbm{1}_{Ff} &= \gamma_{\mathbbm{1}_y,f}\\
                \mathbbm{1}_{Ff}\cdot\iota_x &= \gamma_{f,\mathbbm{1}_x}\,.
                \end{aligned}
            \end{gather}
        \end{enumerate}
        Note that to be completely formal, one should have inserted the unitors and associators of the bicategories $\symbfsf{C},\symbfsf{D}$.
    }
    \newdef{Lax natural transformation}{
        Consider two 2-functors $\func{F,G}{C}{D}$. A lax natural transformation $\eta:F\Rightarrow G$ consists of the following data:
        \begin{enumerate}
            \item For every object $x\in\ob{C}$, a 1-morphism $\eta_x:Fx\rightarrow Gx$.
            \item For every 1-morphism $f:x\rightarrow y$ in $\mathrm{hom}(\symbfsf{C})$, a 2-morphism $\eta_f:Gf\circ\eta_x\Rightarrow\eta_y\circ Ff$ such that the $\eta_f$ are the components of a natural transformation $(\eta_x)^*\circ G\Rightarrow(\eta_y)_*\circ F$ and such that the assignment $f\mapsto\eta_f$ satisfies the `obvious' identity and composition axioms.
        \end{enumerate}
    }
    \begin{remark}\index{pseudo-!natural transformation}
        As usual in the context of higher category theory, one can speak of lax 2-functors if the associator and unitors are merely required to be 2-morphisms, and of strict 2-functors if these morphisms are required to be identities. If the natural transformations between morphism categories in the definition of a lax natural transformation are all isomorphisms, this is called a \textbf{pseudonatural transformation}. If the 1-morphisms $\eta$ are equivalences, they are called lax natural equivalences.
    \end{remark}

    \newdef{Modification}{\index{modification}\label{cat:modification}
        Consider two 2-functors $\func{F,G}{C}{D}$ and two parallel (lax) natural transformations $\alpha,\beta:F\Rightarrow G$. A modification $\mathfrak{m}:\alpha\Rrightarrow\beta$ maps every object $x\in\ob{C}$ to a 2-morphism $\mathfrak{m}_x:\alpha_x\Rightarrow\beta_x$ such that $\beta_f\circ(\mathbbm{1}_{Gf}\cdot\mathfrak{m}_x) = (\mathfrak{m}_y\cdot\mathbbm{1}_{Ff})\circ\alpha_f$.
    }
    This is generalized as follows.
    \newdef{Transfor}{\index{transfor}\label{cat:transfor}
        A $k$-transfor\footnote{This name was first introduced by~\citet{crans_localizations_1998}. A different name that is sometimes used is \textbf{$(n,k)$-transformation}, but this should not be confused with the natural transformations in the context of $(n,r)$-categories.} between two $n$-categories maps $j$-morphisms to $(j+k)$-morphisms (in a coherent way).
    }
    \begin{example}\index{perturbation}
        The definitions for operations in bicategories above lead us to the following `explicit' expressions for $k$-transfors (for small $k$):
        \begin{itemize}
            \item $k=0$: $n$-functors,
            \item $k=1$: ($n$-)natural transformations,
            \item $k=2$: modifications, and
            \item $k=3$: \textbf{perturbations}.
        \end{itemize}
    \end{example}

    The following definition generalizes the notion of essential surjectivity (\cref{cat:essentialy_surjective}) to higher category theory.
    \newdef{$n$-surjective functor}{\index{surjective}
        An $\infty$-functor $\func{F}{C}{D}$ is said to be $n$-surjective if for any two parallel $(n-1)$-morphisms $f,g$ in $\symbfsf{C}$ and $n$-morphism $\alpha:Ff\rightarrow Fg$ in $\symbfsf{D}$, there exists an $n$-morphism $\widetilde{\alpha}$ in $\symbfsf{C}$ such that $F\widetilde{\alpha}\cong\alpha$.
    }

    \newdef{Indexed category}{\index{category!indexed}\label{cat:indexed_category}
        Consider a category $\symbfsf{I}$. An $\symbfsf{I}$-indexed category is a pseudofunctor $\symbfsf{C}:\symbfsf{I}^{\text{op}}\rightarrow\symbfsf{Cat}$, i.e.~a 2-presheaf on $\symbfsf{I}$. Indexed functors and natural transformations are defined analogously.
    }

    \Cref{cat:universal_morphism} can be generalized as follows.
    \newdef{Universal morphism}{\index{universal!morphism}\label{cat:higher_universal_morphism}
        Consider an $n$-functor $\func{F}{C}{D}$ and an object $x\in\ob{D}$. A universal morphism from $x$ to $F$ is a pair $(d,f:x\rightarrow Fd)$ such that:
        \begin{enumerate}
            \item Every morphism $x\rightarrow Fd'$ factors through $f$, and
            \item The precomposition $f^*:\symbfsf{D}(Fd,Fd')\rightarrow\symbfsf{D}(x,Fd')$ is fully faithful.
        \end{enumerate}
    }

\subsection{Higher (co)limits}\label{section:higher_limits}

    \newdef{Weighted 2-limit}{\index{limit!weighted}
        Consider 2-categories $\symbfsf{I,C}$ together with 2-functors $\func{W}{I}{Cat}$ and $\func{F}{I}{C}$. By direct generalization of the ordinary definition of weighted limits (\cref{section:weighted_limits}), one says that $\wlim{W}F$ is the $W$-weighted (2-)limit of $F$ if there exists a pseudonatural equivalence
        \begin{gather}
            \symbfsf{C}(x,\wlim{W}F)\cong\funccat{I}{Cat}\bigl(W,\symbfsf{C}(x,F-)\bigr).
        \end{gather}
        By restricting to the 2-category of strict 2-categories, strict 2-functors and strict natural transformations the resulting notion of a weighted 2-limit coincides with that of an ordinary weighted limit enriched in $\symbfsf{Cat}$ (since strict 2-categories are simply $\symbfsf{Cat}$-enriched 1-categories.)
    }

\section{\texorpdfstring{$\infty$-Categories}{Infinity-categories}}\label{section:infinity_categories}
\subsection{Simplicial approach}

    The first approach to $\infty$-category theory is the simplicial one. The motivation is \cref{model:horn_filler}, which relates the categorical structure to the existence of certain horn fillers. The generalization is then given by the notion of quasicategories (\cref{model:quasicategory}).

    \begin{theorem}[Lurie]\index{Lurie}\label{model:lurie_presentation}
        An $\infty$-category is presentable if and only if it is equivalent to  the coherent nerve of the fibrant-cofibrant subcategory of a combinatorial model category and, hence by Dugger's theorem~\ref{model:dugger}, can be presented by simplicial presheaves.
     \end{theorem}

\subsection{Quasicategories}

    \newdef{Overcategory}{
        Consider a simplicial morphism $F:X\rightarrow\symbfsf{C}$ where $\symbfsf{C}$ is a quasicategory. The overcategory $\symbfsf{C}_{/F}$, generalizing the comma category $\Delta\downarrow F$, is characterized by a natural isomorphism
        \begin{gather}
            \hom_{\symbfsf{sSet}}(Y,\symbfsf{C}_{/F})\cong\hom_{X_{/\symbfsf{sSet}}}(\iota_Y,F)\,,
        \end{gather}
        where $\iota_Y:X\hookrightarrow Y\star X$ is the join inclusion.
    }
    \begin{remark}[1-categorical interpretation]
        Consider the case where $X$ is the simplicial nerve of the one-object groupoid $\{\ast\}$. The functor $F$ then singles out an object $c\in\ob{C}$. Since the image of $F$ is 0-truncated, morphisms on the right-hand side are also 0-truncated.
    \end{remark}

    \newdef{Terminal object}{
        An object $1\in\ob{C}$ such that
        \begin{enumerate}
            \item The projection $\symbfsf{C}_{/1}\rightarrow\symbfsf{C}$ is an acyclic fibration, and
            \item for every object $d\in\ob{C}$ the right hom-Kan complex $\hom^R(d,1):=\symbfsf{C}_{/1}\times_{\symbfsf{C}}\{d\}$ is contractible.
        \end{enumerate}
    }

    \Cref{cat:limit} can be generalized as follows.
    \newdef{Limit}{
        Consider an $(\infty,1)$-functor $\func{F}{C}{D}$ of quasicategories. The limit of $F$ is defined as the terminal object in the overcategory $\symbfsf{C}_{/F}$.
    }

\subsection{Categorical notions}

    \newdef{Truncated object}{
        An $\infty$-groupoid is said to be $n$-truncated for $n\in\mathbb{N}$ if it is an $n$-groupoid. An object $x\in\ob{C}$ of an $(\infty,1)$-category is said to be $n$-truncated if for all objects $y\in\ob{C}$, the hom-groupoid $\symbfsf{C}(y,x)$ is $n$-truncated.
    }

    \newdef{Monomorphism}{\index{monomorphism}
        A morphism $f:x\rightarrow y$ in an $(\infty,1)$-category $\symbfsf{C}$ that is $(-1)$-truncated as an object of the slice category $\symbfsf{C}_{/y}$. Equivalently, the projection $\symbfsf{C}_{/f}\rightarrow\symbfsf{C}_{/y}$ is fully faithful.
    }

    \newdef{Epimorphism}{\index{epimorphism}
        A morphism $f:x\rightarrow y$ in an $(\infty,1)$-category $\symbfsf{C}$ such that the induced morphism $\symbfsf{C}(f,z):\symbfsf{C}(y,z)\rightarrow(x,z)$ is monic for all objects $z\in\ob{C}$.
    }

    \section{Stacks}\index{stack}\label{section:stacks}
    \subsection{2-sheaves}
    
        An important subject, especially in the context of gauge theories in physics, is that of groupoid-valued (pre)sheaves. To this end, sites are generalized to higher category theory.
        \newdef{2-presheaf}{\index{presheaf}\index{pre-!stack}
            Consider a 2-category $\symbfsf{C}$. A 2-presheaf on $\symbfsf{C}$ is a pseudofunctor $\cfunc{F}{C}{Cat}$. When $\symbfsf{C}$ is the categorification of a 1-category, i.e.~when it has discrete hom-categories, 2-presheaves are better known as \textbf{prestacks}.
        }
        \newdef{2-coverage}{\index{coverage}\index{site}
            This is virtually the same as an ordinary coverage (\cref{topos:coverage}), but factorization is only required to exist up to an isomorphism. A 2-category equipped with a 2-coverage is called a \textbf{2-site}.
    
            As for 1-sites, every coverage generates a unique sieve. It is the full subcategory on those morphisms that factor through a covering map in the given coverage (again up to isomorphism).
        }
    
        As in the case of ordinary categories (\cref{topos:local_object_sheaf}), one can define 2-sheaves through a descent condition.
        \newdef{2-sheaf}{\label{topos:2_sheaf}
            A 2-presheaf $\cfunc{F}{C}{Cat}$ on a 2-site $(\symbfsf{C},J)$ is said to be a 2-sheaf with respect to $J$ if for all sieves $S\in J$ the following functor is an equivalence:
            \begin{gather}
                Fc\cong\symbfsf{Psh}_2(\mathcal{Y}c,F)\rightarrow\symbfsf{Psh}_2(S,F)\,,
            \end{gather}
            where the fist equivalence is just the 2-Yoneda lemma.
        }
        \begin{remark}
            It should be noted that 2-(pre)sheaves can also be defined on ordinary (1-)sites. Sieves, regarded as subfunctors of the Yoneda embedding, take values in $\symbfsf{Set}$. By composing these with the embedding $\symbfsf{Set}\hookrightarrow\symbfsf{Cat}$ of sets as (discrete) categories, one obtains 2-subfunctors of the 2-Yoneda embedding. Often 2-sheaves over 1-sites are called \textbf{stacks} (although this terminology is also used for general 2-sites).
        \end{remark}
    
        \newdef{Prestack of groupoids}{
            Consider a category $\symbfsf{C}$. A prestack of groupoids on $\symbfsf{C}$ is a $\symbfsf{Grpd}$-valued prestack on $\symbfsf{C}$.
    
            The category of (groupoid-valued) prestacks becomes $\symbfsf{Grpd}$-enriched if one takes the hom-category between two prestacks $F,G$ to consist of the following data:
            \begin{itemize}
                \item\textbf{Objects}: Natural transformations $\alpha:F\Rightarrow G$ (note that the components are themselves functors).
                \item\textbf{Morphisms}: `strict modifications' in the sense that they map objects in $\symbfsf{C}$ to natural transformations satisfying the whiskering condition (see also Definition \ref{cat:modification})
                \begin{gather}
                    \mathbbm{1}_{Ff}\cdot\mathfrak{m}_b = \mathfrak{m}_a\cdot\mathbbm{1}_{Gf}\,.
                \end{gather}
            \end{itemize}
        }
    
        For ordinary sites and presheaves, descent was defined in terms of matching families. Since presheaves are now taking values in a 2-category, the matching families are a bit more complex. However, this structure is already familiar from differential geometry and algebraic topology, where it is known under the name of the \textit{\v{C}ech nerve}.
        \newdef{\v{C}ech groupoid}{\index{Cech!groupoid}
            Consider a site $(\symbfsf{C},J)$. To every covering family $\mathcal{U}=\{f_i:x_i\rightarrow x\}_{i\in I}$, one can assign an internal groupoid in presheaves $C(\mathcal{U})$ consisting of the following data:
            \begin{itemize}
                \item\textbf{Objects}: $\bigsqcup_i\mathcal{Y}x_i$, and
                \item\textbf{Morphisms}: $\bigsqcup_{i,j}\mathcal{Y}x_i\times_{\mathcal{Y}x}\mathcal{Y}x_j$.
            \end{itemize}
            This is equivalent to the ($\symbfsf{Grpd}$-valued) presheaf that assigns to every object $y\in\ob{C}$ the groupoid consisting of the following data:
            \begin{itemize}
                \item\textbf{Objects}: Pairs $(i,g_i:y\rightarrow x_i)$ where $x_i\in\mathcal{U}$, and
                \item\textbf{Morphisms}: A unique arrow $(i,g_i)\rightarrow(j,g_j)$ if and only if $f_i\circ g_i = f_j\circ g_j$.
            \end{itemize}
        }
        Comparing the definition of morphisms in the \v{C}ech groupoid to the condition for matching families in \cref{topos:matching_family}, shows that one could presume that the \v{C}ech groupoid is related to the matching families. This intuition is indeed correct as explained by the following property.
        \begin{property}[Matching families]\label{topos:cech_matching_families}
            Any ordinary presheaf $F$ can be considered to be $\symbfsf{Grpd}$-valued by postcomposing with the embedding $\symbfsf{Set}\hookrightarrow\symbfsf{Grpd}$. For any covering family $\mathcal{U}$, there exists an isomorphism
            \begin{gather}
                \cfunccat{C}{Grpd}\bigl(C(\mathcal{U}),F\bigr)\cong\symbfsf{Psh}_2(\mathcal{U},F)\,.
            \end{gather}
            Because the \v{C}ech groupoid (co)represents a descent object, it is sometimes called a \textbf{codescent object}.
        \end{property}
        It is exactly this (co)descent property of the \v{C}ech groupoid that will be used in \labelref{chapter:hdg} to define (higher) smooth groupoids. Readers with some experience in algebraic topology will also notice that the \v{C}ech groupoid only contains the first degrees of the \v{C}ech complex. The full \v{C}ech complex can be obtained from the following construction.
        \newdef{\v{C}ech nerve}{\index{Cech!nerve}
            Consider a morphism $f:y\rightarrow x$ in a category $\symbfsf{C}$. The \v{C}ech nerve $C_\bullet(f)$ is the simplicial object (\cref{model:simplicial_object}) that contains, in degree $k\in\mathbb{N}$, the $(k+1)$-fold pullback of $f$ along itself. For a covering family $\mathcal{U}\equiv\{f_i:x_i\rightarrow x\}$, the \v{C}ech nerve is defined as
            \begin{gather}
                C_\bullet(\mathcal{U}):=C_\bullet\left(\bigsqcup_ix_i\rightarrow x\right)\,.
            \end{gather}
        }
        For $\infty$-sheaves, the full \v{C}ech nerve will be used. However, for 2-sheaves and, in particular, stacks, only its 3-coskeleton is necessary. The extra information will encode the \textit{cocycle condition}~\eqref{bundle:G_cocycle_condition}, well-known from the study of \textit{fibre bundles}.
    
    \subsection{Stacks on a 1-site}
    
        For the definition of stacks, one needs the notions of fibred categories or, equivalently, pseudofunctors as defined in \cref{section:fibred_categories}. The definitions are recalled here.
        \begin{quote}
            Consider a functor $\func{\Pi}{A}{B}$. A morphism $f$ in $\symbfsf{A}$ is said to be $\Pi$-Cartesian if, for every morphism $\varphi$ in $\symbfsf{A}$ and factorization of $\Pi\varphi$ through $\Pi f$ in $\symbfsf{B}$, there exists a unique factorization of $\varphi$ through $f$. $f$ is called the inverse image of $\Pi f$.
    
            A fibred category consists of a functor $\func{\Pi}{A}{B}$ such that for each morphism $f:c\rightarrow d$ in $\symbfsf{B}$ with $d\in\im(\Pi)$ and each lift $y\in\symbfsf{A}_d$ there exists at least one inverse image in $(\widetilde{f}:x\rightarrow y)\in\symbfsf{A}$ of $f$. By the Grothendieck construction every fibred category gives rise to a pseudofunctor $\cfunc{F}{B}{Cat}$ by sending objects to their fibres under $\Pi$ and sending morphisms $f$ to their pullback functors $f^*$.
        \end{quote}
    
        \newdef{Descent datum}{\index{descent}
            Consider a category $\symbfsf{C}$ with a covering family $\mathcal{U}\equiv\{f_i:x_i\rightarrow x\}$ and a pseudofunctor $\cfunc{F}{C}{Cat}$. The projections associated to the pullback $x_i\cap x_j:=x_i\times_xx_j$ will be denoted by $\pi_i$ and $\pi_j$, respectively (and analogously for iterated pullbacks). A descent datum for $\mathcal{U}$ with respect to $F$ is a pair of families $(\{g_i\},\{f_{ij}\})_{i,j\in I}$, where $\{g_i\}$ is a matching family for $\mathcal{U}$ with respect to $F$ and every $f_{ij}$ is an isomorphism $\pi_i^*x_i\cong \pi_j^*x_j$. This data is required to satisify the following \textbf{cocycle condition}:
            \begin{gather}
                \pi_{ik}^*f_{ik} = \pi_{ij}^*f_{ij}\circ\pi_{jk}^*f_{jk}\,.
            \end{gather}
            Morphisms $(\{g_i\},\{f_{ij}\})\rightarrow(\{g'_i\},\{f'_{ij}\})$ between descent data are families of morphisms $\{\phi_i:g_i\rightarrow g'_i\}$ that satisfy
            \begin{gather}
                \pi_i^*\phi_i\circ f_{ij} = f'_{ij}\circ\pi_j^*\phi_j\,.
            \end{gather}
            The category of descent data for $\mathcal{U}$ with respect to $F$ will be denoted by $\mathrm{Descent}(\mathcal{U},F)$.
        }
        \begin{construct}
            Consider an object $\xi$ in $Fx$. From this object, one can construct a descent datum as follows. The morphisms $g_i$ are the pullbacks $f_i^*\xi$ and the isomorphisms $f_{ij}:\pi_i^*f_i^*\xi\cong\pi_j^*f_j^*\xi$ are obtained from the fact that both these objects are (Cartesian) pullbacks of the same morphisms. Arrows in $Fx$ induce morphisms of descent data by (Cartesian) pullbacks along the covering maps. This construction defines a functor $Fx\rightarrow\mathrm{Descent}(\mathcal{U},F)$. It can be shown that this construction is independent of a choice of cleavage up to equivalence.
        \end{construct}
    
        \newdef{Stack}{\index{pre-!stack}\index{stack}
            Consider a fibred category $F$ over a site $(\symbfsf{C},J)$.
            \begin{itemize}
                \item $F$ is called a \textbf{separated prestack} if for each covering family $\mathcal{U}$ on $x\in\ob{C}$, the functor $Fx\rightarrow\mathrm{Descent}(\mathcal{U},F)$ is fully faithful.
                \item $F$ is called a \textbf{stack} if for each covering family $\mathcal{U}$ on $x\in\ob{C}$ the functor $Fx\rightarrow\mathrm{Descent}(\mathcal{U},F)$ is an equivalence.
            \end{itemize}
            This is a generalization of the descent condition in \cref{topos:local_object_sheaf}. This can be seen by observing that $\mathrm{Descent}(\mathcal{U},F)\cong\symbfsf{Psh}_2(S(\mathcal{U}),F)$, where $S(\mathcal{U})$ is the sieve generated by $\mathcal{U}$ regarded as a fibred category. When $F$ is fibred over groupoids, it is called a \textbf{stack of groupoids}. This forms the category $\symbfsf{Sh}_{(2,1)}(\symbfsf{C})$ of $(2,1)$-sheaves. In fact, it is this subcategory that is usually meant when considering stacks.
        }
    
        A more conceptual (although completely equivalent) generalization from (1-)sheaves to 2-sheaves can be obtained by starting from \cref{topos:cech_matching_families}. There, it was shown that matching families for (1-)presheaves can be obtained as natural transformations from the \v{C}ech groupoid.
        \begin{property}[Descent data and \v{C}ech nerve]
            Let $C(\mathcal{U})$ denote the 3-coskeleton of the \v{C}ech nerve $C_\bullet(\mathcal{U})$. Pseudonatural transformations $C(\mathcal{U})\Rightarrow F$ can be shown to be equivalent to tuples $(c,\{c_i\},\{c_{ij}\},\{c_{ijk}\})$, with $c_i\in Fx_i$, that fit into cubes lying in the image of $C_2(\mathcal{U})$ in which all edges consist of Cartesian morphisms. Arrows between such cubes are given by arrows between the vertices that make the `obvious' diagrams commute.
    
            By comparing these cubes to the previous definition of descent data, one obtains the following equivalence:
            \begin{gather}
                \mathrm{Descent}(\mathcal{U},F)\cong\cfunccat{C}{Cat}\bigl(C(\mathcal{U}),F\bigr)\,.
            \end{gather}
        \end{property}
    
        \begin{remark}[1-sheaves]
            Although most of the above statements and constructions might seem very abstract and complex compared to ordinary sheaf theory, it is not quite so. In fact, when restricting to pseudofunctors of the form $\op{\symbfsf{C}}\rightarrow\symbfsf{Set}$, where the embedding $\symbfsf{Set}\hookrightarrow\symbfsf{Cat}$ sends sets to discrete categories, one obtains ordinary sheaves as a subcategory of stacks. For example, by the equivalence between pseudofunctors and Grothendieck fibrations, it is known that the Cartesian pullbacks $f^*$ are in fact just the images of $f$ under the pseudofunctor $F$. This way, the condition $\pi_1^*c_i\cong\pi^*_2c_j$ can be rewritten as $Ff'_i(c_i)=Ff'_j(c_j)$, which is nothing but the matching family condition~\eqref{topos:matching_family_condition}.
        \end{remark}
    
    \section{Higher topos theory}
    
        In this section, the notion of topos is generalized from ordinary category theory to higher category theory. In particular, $\infty$-sheaves will be defined. This will require a suitable foundation for $\infty$-category theory. To this end, the language of (simplicial) model categories as introduced in \labelref{chapter:model_theory} will be used.
    
        \newdef{\texorpdfstring{$\infty$-groupoid}{Infinity-groupoid}}{\index{groupoid}\label{topos:infty_groupoid}
            Objects of the full simplicial subcategory of $\symbfsf{sSet}_\text{Quillen}$ on Kan complexes. From \cref{model:horn_filler}, it is immediately clear how this generalizes the definition of ordinary groupoids. For groupoids one needs unique horn fillers (composition in ordinary categories is unique), while for $\infty$-groupoids this is allowed to be unique up to higher coherence.
        }
        \newdef{\texorpdfstring{$(\infty,1)$-category}{(Infinity,1)-category}}{\index{category}
            An $\infty\symbfsf{Grpd}$-enriched category or, equivalently, a simplicially enriched category for which all hom-objects are Kan complexes. The functor category between $(\infty,1)$-categories is defined through the (simplicial) nerve and realization functors (\cref{model:nerve}):
            \begin{gather}
                \funccat{C}{D} := |\symbfsf{sSet}(N\symbfsf{C},N\symbfsf{D})|\,.
            \end{gather}
        }
    
        \begin{property}[\v{C}ech model structure]\index{Cech!model structure}\label{topos:cech_model_structure}
            For any small category $\symbfsf{C}$, the $\infty$-category of $\infty\symbfsf{Grpd}$-valued $\infty$-sheaves can be represented by the category $\cfunccat{C}{sSet}$ of simplicial presheaves on $\symbfsf{C}$ by a theorem of \textit{Lurie} (\cref{model:lurie_presentation}), i.e.~there exists an $\infty$-equivalence between $\symbfsf{Sh}_{(\infty,1)}(\symbfsf{C})$ and the full subcategory on fibrant-cofibrant objects of the (left Bousfield) localization of $\cfunccat{C}{sSet}$ at the \v{C}ech nerve projections. The resulting model structure is called the \textbf{\v{C}ech model structure}.
    
            A presheaf $X$ is fibrant in this model structure if the map
            \begin{gather}
                \hom(M,X)\rightarrow\hom\bigl(\mathcal{C}(\mathcal{U}),X\bigr)
            \end{gather}
            is a weak equivalence for all open covers $\mathcal{U}$, i.e.~exactly if $X$ satisfies the descent condition and, hence, is an $\infty$-stack.
        \end{property}
    
        The most straightforward definition of an $\infty$-sheaf generalizes \cref{topos:local_object_sheaf}.
        \newdef{\texorpdfstring{$\infty$-sheaf}{Infinity-sheaf}}{\index{sheaf}
            Consider an $\infty$-site $(\symbfsf{C},J)$ and let $S$ denote the collection of monomorphisms in $\symbfsf{Psh}_\infty(\symbfsf{C})$ induced by the covering sieves. An $\infty$-presheaf on $\symbfsf{C}$ is called a $J$-sheaf if it is $S$-local. A presheaf with values in an $\infty$-category $\symbfsf{D}$ is called a sheaf if the representable presheaf $\symbfsf{D}(x,F-)$ is a $J$-sheaf for all $x\in\ob{D}$.
    
            In terms of the \v{C}ech nerve $\mathcal{C}$, the descent condition can be written as follows:
            \begin{gather}
                Fx\simeq\symbfsf{Psh}_\infty\bigl(\mathcal{C}(\mathcal{U}),F\bigr)
            \end{gather}
            for all covers $\mathcal{U}$ of $x$, where $\simeq$ denotes a weak equivalence.
        }
        \newdef{\texorpdfstring{$\infty$-stack}{Infinity-stack}}{\index{stack}
            An $(\infty,1)$-sheaf taking values in $\infty\symbfsf{Grpd}$.
        }
    
        \Cref{topos:global_sections} can be generalized as follows.
        \begin{property}
            For every $\infty$-topos $\symbfsf{H}$, there exists a geometric morphism $(\mathrm{Disc}\dashv\Gamma):\symbfsf{H}\leftrightarrows\infty\symbfsf{Grpd}$. Any morphism into a discrete object $\mathrm{Disc}(X)$ is constant.
    
            The left adjoint is sometimes called the \textbf{discrete object functor}. This terminology stems from the case of the forgetful functor $\func{\Gamma}{Top}{Set}$, where the (fully faithful) left adjoint equips a set with the discrete topology.
        \end{property}
        \begin{example}[Sheaves on manifolds]
            One of the archetypal examples of $\infty$-topoi is the topos of sheaves over smooth manifolds. By the Yoneda embedding, one can regard a manifold as a sheaf and the global sections functor maps this representable sheaf to the manifold itself: $\Gamma(M)=M$. For a Lie group, one can construct the classifying stack $\mathcal{B}G$. The global sections functor maps this stack to the delooping groupoid $\symbfsf{B}G$.
        \end{example}
    
        \newdef{Mapping stack}{\label{topos:mapping_stack}
            Consider two $\infty$-stacks $X,Y\in\symbfsf{Sh}_{(\infty,1)}(\symbfsf{C})$. The mapping stack is defined as follows:
            \begin{gather}
                [X,Y](U):=\symbfsf{Sh}_{(\infty,1)}(\symbfsf{C})(X\times U,Y)\,,
            \end{gather}
            where on the right-hand side, $U$ denotes the representable $\infty$-stack.
        }
    
        \todo{FINISH (PERHAPS MOVE infinity-CATEGORY STUFF TO CHAPTER ON MODEL THEORY)}
    
    \section{Cohomology}\index{cohomology}
    
        In this section, cohomology will be generalized to the $\infty$-categorical setting.
    
        First, take a topological space $X$ and an $\infty$-groupoid $G$. Geometric realization (\cref{model:geometric_realization}) gives an equivalence $\infty\symbfsf{Grpd}\cong\symbfsf{Top}$ and, therefore, one can define the intrinsic cohomology of $X$ with coefficients in $G$ as follows:
        \begin{gather}
            H(X;G) := \pi_0\symbfsf{Top}(X,|G|)\,.
        \end{gather}
        $X$ can also be identified with its petit ($\infty$-)topos $\symbfsf{Sh}_{(\infty,1)}(X)$, in which $X$ sits as the terminal object. From this point of view, the intrinsic cohomology of $X$ with coefficients in $G$ is
        \begin{gather}
            \overline{H}(X;G) := \pi_0\symbfsf{Sh}_{(\infty,1)}(X)(X,\mathrm{LConst}\,G)\cong\pi_0\circ\Gamma\circ\mathrm{LConst}(G)\,.
        \end{gather}
        This is the \textbf{cohomology with constant coefficients} of $X$ with coefficients in $G$. If $X$ is paracompact, the two cohomologies coincide: $H(X;G)\cong\overline{H}(X;G)$.
    
        Now, it is time to pass to general cohomology.
        \newdef{Intrinsic cohomology}{\index{cohomology}\label{topos:cohomology}
            Consider a $(\infty,1)$-category $\symbfsf{H}$. For every two objects $X,A\in\symbfsf{H}$, the hom-space $\symbfsf{H}(X,A)$ is an $\infty$-groupoid. Define the following notions:
            \begin{itemize}
                \item The objects in $\symbfsf{H}(X,A)$ are called \textbf{cocycles}.
                \item The morphism in $\symbfsf{H}(X,A)$ are called \textbf{coboundaries}.
                \item The set of connected components
                \begin{gather}
                    H(X;A):=\pi_0\symbfsf{H}(X,A)=\hom_{\symbfsf{Ho(H)}}(X,A)\,,
                \end{gather}
                where $\symbfsf{Ho(H)}$ is the homotopy category \ref{model:homotopy_category_2} of $\symbfsf{H}$, is called the intrinsic cohomology of $X$ with coefficients in $A$.
            \end{itemize}
            If the object $A$ admits an $n$-delooping $\symbfsf{B}^nA$, the $n^{\text{th}}$ cohomology group of $X$ is defined as
            \begin{gather}
                H^n(X;A):=H(X;\symbfsf{B}^nA)\,.
            \end{gather}
        }
    
        \begin{example}[Singular cohomology]
            Consider a topological space $X$. For every group $G$ one can define the first delooping (\cref{cat:group_delooping}), so one can also define the zeroth and first cohomology groups $H^{0,1}(X;G)$. Only when $G$ is Abelian do higher deloopings exists (in fact, if $G$ is Abelian all higher deloopings exist), and so in this case higher cohomology groups $H^{\geq 2}(X;G)$ can be defined. It can be shown that these coincide with the singular cohomology groups of $X$.
        \end{example}
    
        \begin{example}[Group cohomology]
            Consider a (discrete) group $G$ together with its delooping groupoid $\symbfsf{B}G$. The cohomology of a group with coefficients in an Abelian group $A$ (\cref{section:group_cohomology}) is given by the intrinsic cohomology of $\infty\symbfsf{Grpd}$ of the delooping groupoids:
            \begin{gather}
                H(G;A)\cong\pi_0\infty\symbfsf{Grpd}(\symbfsf{B}G,\symbfsf{B}A)\,.
            \end{gather}
        \end{example}
    
        By replacing the topos $\symbfsf{H}$ by a slice topos $\symbfsf{H}_{/X}$, one obtains twisted cohomology.
        \newdef{Twisted cohomology}{\index{cohomology!twisted}
            Consider a $(\infty,1)$-topos $\symbfsf{H}$ with some object $X\in\ob{H}$. The mapping space $\symbfsf{H}(X,A)$, the cocycles of $X$ with coefficients in $A$, is easily seen to be isomorphic to the mapping space $\symbfsf{H}_{/X}(X,X\times A)$, where the second argument is equipped with the canonical projection morphism. Morphisms in this space are just sections of the trivial $A$-$\infty$-bundle over $X$. General twisted cohomology can then be defined as the space of sections of an arbitrary $A$-$\infty$-bundle over $X$.
    
            By passing to classifying morphisms of bundles, one obtains the twist $\chi:X\rightarrow\symbfsf{BAut}(A)$ and the universal bundle $\rho_A:A/\!\!/\symbfsf{Aut}(A)\rightarrow\symbfsf{BAut}(A)$. $\chi$-twisted cohomology is then given by (the connected components of) the following mapping space:
            \begin{gather}
                \symbfsf{H}_{/\symbfsf{BAut}(A)}(\chi,\rho_A)\,.
            \end{gather}
        }