\chapter{Algebraic Topology}

    References for this chapter are \cite{massey, merry_io}.

\section{Homotopy theory}\label{section:homotopy}
\subsection{Homotopy}

    \newdef{Retraction}{\index{retraction}
        Let $X$ be a topological space and let $A\subset X$ be a subspace. A continuous function $f:X\rightarrow A$ is called a retraction (and $A$ is called a \textbf{retract} of $X$) if it satisfies $f(a)=a$ for all $a\in A$.
    }
    \newdef{Homotopy}{\index{homotopy}
        Let $f, g\in C(X,Y)$ where $X, Y$ are topological spaces. If there exists a continuous function $H:X\times [0, 1]\rightarrow Y$ such that $f(x) = H(x, 0)$ and $g(x) = H(x, 1)$ then $f$ and $g$ are said to be homotopic. This relation induces an equivalence relation on $C(X, Y)$.
    }
    \newdef{Deformation retraction}{
        Let $X$ be a topological space and let $A\subseteq X$ be a subspace. $A$ is called a deformation retract if there exists a homotopy between the identity map on $X$ and a retraction $f:X\rightarrow A$.
    }

    \newdef{Homotopy type}{
        Two topological spaces $X$ and $Y$ are said to be homotopy equivalent, or of the same homotopy type, if there exist continuous functions $f:X\rightarrow Y$ and $g:Y\rightarrow X$ such that $f\circ g$ is homotopic to $\mathbbm{1}_Y$ and $g\circ f$ is homotopic to $\mathbbm{1}_X$. The maps $f, g$ are called \textbf{homotopy equivalences}.
    }
    \begin{property}\index{homeomorphism}
        Every homeomorphism is a homotopy equivalence.
    \end{property}

    \newdef{Null-homotopic}{
        A continuous function is null-homotopic if it is homotopic to a constant function.
    }
    \newdef{Contractible space}{\index{contractible}\label{topology:contractible_space}
        A topological space $X$ is said to be contractible if the identity map $\mathbbm{1}_X$ is null-homotopic or equivalently, if the space is homotopy-equivalent to a point.
    }

\subsection{Homotopy groups}

    In this subsection we will always assume to be working with pointed spaces \ref{topology:pointed_space}. The base point will be denoted by $\ast$.

    \newdef{Loop space}{\index{loop!space}
        \nomenclature[S_zsymOmega]{$\Omega X$}{(Based) loop space on $X$.}
        \nomenclature[S_LX]{$LX$}{Free loop space on $X$.}
        The set of all \textbf{loops} in a pointed topological space $(X, \ast)$, i.e. all continuous maps\footnote{The mapping space is equipped with the compact-open topology.} $\delta:(S^1, t_0)\rightarrow (X, \ast)$ for which $\delta(t_0) = \ast$. This space is denoted by $\Omega X$. It can be equipped with a multiplication operation corresponding to the concatenation of loops\footnote{It should be noted that the speed at which the concatenated loops are traversed is doubled because the parameter $t$ should remain an element of $S^1\cong[0, 1]/_{0\sim1}$.}.

        When one drops the requirement of based loops, i.e. one considers the space of all continuous maps $S^1\rightarrow X$, the resulting space is called the \textbf{free loop space} on $X$. This space is denoted by $LX$.
    }
    \newdef{Loop group}{\label{topology:loop_group}\index{loop!group}
        In the case of topological groups one can define a group structure on the (free) loop space. It can even be turned into a topological group by equipping it with the compact-open topology.
    }

    \newdef{Fundamental group}{\index{group!fundamental}
        The fundamental group $\pi_1(X, x_0)$ is defined as the loop space of $(X, x_0)$ modulo homotopy. As the name implies the fundamental group can be given the structure of a multiplicative group where the operation is inherited from that of the loop space.
    }
    \begin{remark}
        In general, as the notation implies, the fundamental group depends on the base point $x_0$. However when the space $X$ is path-connected, the fundamental groups belonging to different base points are isomorphic. It follows that we can speak of 'the' fundamental group in the case of path-connected spaces.
    \end{remark}

    \newdef{Fundamental groupoid $\clubsuit$}{\index{groupoid!fundamental}\index{Poincar\'e!groupoid}
        Let $X$ be a topological space. The fundamental groupoid\footnote{Sometimes called the \textbf{Poincar\'e groupoid}.} $\mathbf{\Pi}_1(X)$ is the groupoid which has the points of $X$ as objects and the endpoint-preserving homotopy classes of continuous functions $f:S^1\rightarrow X$ as morphisms. The fundamental group $\pi_1(X, a)$ can be recovered as the automorphism group of $a\in\text{ob}(\mathbf{\Pi}_1(X))$.
    }

    \newdef{Simply-connected space}{\index{simply-connected}\label{topology:simply_connected}
        A topological space is said to be simply-connected if it is path-connected and if the fundamental group is trivial.
    }

    \newdef{Universal covering space}{
        A covering space\footnote{See definition \ref{topology:covering_space}.} $C$ is said to be universal if it is simply-connected.
    }
    \begin{uproperty}
        Let $X$ be a topological space and let $\widetilde{X}$ be the universal covering space of $X$. Every other covering space $C$ of $X$ is also covered by $\widetilde{X}$.
    \end{uproperty}
    \begin{property}[Universal cover]\index{deck transformation}
        Consider a topological space $X$ and let $\widetilde{X}$ be its universal covering space. The group of deck transformations\footnote{See definition \ref{topology:deck_transformation}.} Aut$(\widetilde{X})$ is isomorphic to the fundamental group $\pi_1(X)$. Hence we obtain
        \begin{gather}
            X\cong \widetilde{X}/\pi_1(X).
        \end{gather}
    \end{property}

    The definition of a fundamental group can be generalized to arbitrary dimensions in the following way\footnote{Note however that we replace the interval $[0, 1]$ by the sphere $S^1$. This is non-restrictive as we can construct $S^n$ by identifying the boundary of $[0,1]^n$ with the basepoint $x_0$.}:
    \newdef{Homotopy group}{\index{homotopy!group}
        \nomenclature[S_zsymhos]{$\pi_n(X, x_0)$}{$n^{th}$ homotopy space over $X$ with basepoint $x_0$.}
        The homotopy group $\pi_n(X, x_0)$ is defined as the set of homotopy classes of continuous maps $f:S^n\rightarrow X$ based at $x_0\in X$. The set $\pi_0(X, x_0)$ is defined as the set of path-connected components of $X$.
    }

    \begin{property}
        For $n\geq1$ the sets $\pi_n(X, x_0)$ are groups.
    \end{property}
    \begin{property}\label{topology:abelian_homotopy_groups}
        For $n\geq2$  the homotopy groups $\pi_n(X, x_0)$ are Abelian. (This follows from an application of the Eckmann-Hilton argument.)
    \end{property}

    \begin{property}
        If $X$ is path-connected, then the homotopy groups $\pi_n(X, x_0)$ and $\pi_n(X, x_1)$ are isomorphic for all $x_0, x_1\in X$ and all $n\in\mathbb{N}$.
    \end{property}
    \begin{property}
        Homeomorphic spaces have isomorphic homotopy groups $\pi_n$.
    \end{property}

    \begin{formula}
        Let $(X, x_0)$ and $(Y, y_0)$ be pointed topological spaces with homotopy groups $\pi_n(X, x_0)$ and $\pi_n(Y, y_0)$. The homotopy groups of their product is given by
        \begin{gather}
            \pi_n(X\times Y, (x_0, y_0)) = \pi_n(X, x_0)\times\pi_n(Y, y_0).
        \end{gather}
    \end{formula}

    \newdef{$n$-connected space}{\index{connected}
        A topological space is said to be $n$-connected if its first $n$ homotopy groups are trivial.
    }

    \newdef{Weak homotopy equivalence}{\index{homotopy!equivalence}
        A continuous map which induces isomorphisms on all homotopy groups.
    }
    \begin{property}[Homotopy category $\clubsuit$]
        \nomenclature[S_hTop]{\textbf{hTop}}{Homotopy category}
        The homotopy category hTop has as objects the topological spaces and as morphisms the homotopy classes of continuous maps. It is immediately clear that there exists a functor $F:\textbf{Top}\rightarrow\textbf{hTop}$ that maps topological spaces to themselves and continuous maps to their homotopy classes.

        In fact the above definition is often too restrictive. \textit{Quillen} gave a more general construction: The homotopy category (in the sense of Quillen) is obtained as the localization\footnote{See definition \ref{cat:localization}.} of \textbf{Top} with respect to the collection of weak homotopy equivalences.
    \end{property}

    \newdef{Eilenberg-MacLane space}{\index{Eilenberg-MacLane}
        Let $G$ be a group and choose a positive integer $n\in\mathbb{N}_0$. The Eilenberg-MacLane space $K(G, n)$ is a topological space with the following property:
        \begin{gather}
            \pi_i\Big(K(G, n)\Big)=\begin{cases}
                G&i=n\\
                0&i\neq n.
            \end{cases}
        \end{gather}
        It follows from property \ref{topology:abelian_homotopy_groups} above that for $n>1$ the group $G$ has to be Abelian.
    }
    \begin{property}
        For every $G$ and $n$ the space $K(G, n)$ is unique up to weak homotopy equivalence.
    \end{property}
    \begin{property}
        The loop space $\Omega K(G, n)$ is homotopy equivalent to $K(G, n-1)$.
    \end{property}

\subsection{CW complexes}

    \newdef{$n$-cell}{\index{cell}
        An open $n$-cell is a subset of a topological space homeomorphic to the $n$-dimensional open ball. A closed $n$-cell is the image of an $n$-dimensional closed ball under an attaching map\footnote{See definition \ref{topology:attaching_space}.}.
    }

    \newdef{CW complex}{\index{CW!complex}\label{topology:cw_complex}
        A CW complex is a Hausdorff space $X$ together with a partition of $X$ in open cells satisfying following conditions:
        \begin{itemize}
            \item A subset of $X$ is closed if and only if it meets the closure of each cell in a closed set.
            \item For each open $n$-cell $C$ in the partition there exists an attaching map $f:\overline{B}_n\rightarrow X$ such that:
            \begin{itemize}
                \item $f|_{B_n}$ is homeomorphic to $C$.
                \item $f(\partial \overline{B}_n)$ is covered by a finite number of open cells in the partition, each having dimension smaller than $n$.
            \end{itemize}
        \end{itemize}
    }
    \newdef{Regular CW complex}{
        A CW complex is called regular if for every open cell $C$ the attaching map $f$ is a homeomorphism onto the closure $\overline{C}$.
    }
    \newdef{Finite type}{\index{finite!type}
        A CW complex is said to be of finite type if there are only a finite number of cells in each degree.
    }

    \begin{construct}\index{skeleton}
        Every CW complex can, up to isomorphism, be constructed inductively:

        \indent First choose a discrete space $X_0$, i.e. a topological space equipped with the discrete topology. This space forms a 0-cell. Then we can add 1-cells $C_1$ using appropriate attaching maps $f:\partial\overline{B}_1\rightarrow X_0$. This way we obtain a 1-dimensional CW complex $X_1$. Inductively one obtains a sequence of nested $n$-dimensional CW complex $X_0\subset X_1\subset\cdots\subset X_n$.

        The spaces $X_i$ are also called \textbf{$i$-skeletons}.
    \end{construct}
    \remark{Infinite-dimensional CW complexes can be obtained by taking the direct limit\footnote{See definition \ref{direct_limit}.} of the sequence above.}

    \begin{theorem}[Whitehead]\index{Whitehead}
        A continuous map between CW-complexes is a homotopy equivalence if and only if it is a weak homotopy equivalence.
    \end{theorem}
    \begin{theorem}[CW approximation theorem]\index{CW!approximation theorem}
        For every topological space $Y$ there exists a CW complex $X$ together with a weak homotopy equivalence $f:X\rightarrow Y$.
    \end{theorem}

\subsection{Fibrations}

    \newdef{Homotopy lifting property}{
        Consider a continuous map $\pi:E\rightarrow B$ between topological spaces. The map $\pi$ is said to have the homotopy lifting property with respect to a topological space $X$ if for every homotopy $f:X\times[0, 1]\rightarrow B$ and lifting $\widetilde{f}_0:X\rightarrow E$ of $f_0=f|_{X\times\{0\}}$ there exists a homotopy $\widetilde{f}:X\times[0, 1]$ lifting $f$ such that the following diagram commutes:
        \begin{figure}[ht!]
            \centering
            \begin{tikzpicture}
                \matrix (m) [matrix of math nodes,row sep=7em,column sep=7em, minimum width=2em, ampersand replacement=\&]{
                    X \& E\\
                    X\times[0, 1]\& B\\
                };
                \draw[->] (m-1-1) -- (m-1-2) node[pos=0.5, above]{$\widetilde{f}_0$};
                \draw[right hook ->] (m-1-1) -- (m-2-1) node[pos=0.5, left]{$X\times\{0\}$};
                \draw[->] (m-2-1) -- (m-2-2) node[pos=0.5, below]{$f$};
                \draw[->] (m-1-2) -- (m-2-2) node[pos=0.5, right]{$\pi$};
                \draw[dashed, ->] (m-2-1) -- (m-1-2) node[pos=0.5, above]{$\widetilde{f}$};
            \end{tikzpicture}
            \caption{Homotopy lifting property.}
            \label{fig:homotopy_lifting_property}
        \end{figure}

        where $\widetilde{f}$ denotes the lifting of $f$, i.e. $f = \pi\circ\widetilde{f}$.
    }

    \newdef{Hurewicz fibration}{\index{fibration}
        A map $\pi$ satisfying the homotopy lifting property with respect to every topological space $X$ is called a (Hurewicz) fibration.\footnote{If the homotopy lifting property only holds with respect to CW complexes (see definition \ref{topology:cw_complex}) then it is called a \textbf{Serre fibration}.}
    }
    \begin{property}
        Consider a fibration $\pi:E\rightarrow B$ with $B$ path-connected. All fibres, i.e. sets $\pi^{-1}(\{b\})$ where $b\in B$, are homotopy equivalent. Therefore a fibration is often denoted by the diagram $\prin{F}{E}{B}$.
    \end{property}

    \begin{example}[Hopf fibration]\index{Hopf!fibration}
        The Hopf fibration is given by
        \begin{gather}
            \prin{S^1}{S^3}{S^2}.
        \end{gather}
        \textit{Adam's theorem} states that this fibration can be generalized to higher dimensions as $\prin{S^n}{S^{2n+1}}{S^{2n}}$ only for $n\in\{0, 1, 3, 7\}$.
    \end{example}
    \begin{example}
        For all $n\in\mathbb{N}$ the following sequence forms a fibration:
        \begin{gather}
            \prin{\text{SO}(n)}{\text{SO}(n+1)}{S^n}.
        \end{gather}
    \end{example}

\section{Simplicial homology}\label{section:homology}
\subsection{Simplices}

    \newdef{Simplex}{\index{simplex}\index{vertex}\label{topology:simplex}
        A $k$-simplex $\sigma^k = [t_0,\ldots,t_k]$ is defined as the following set:
        \begin{gather}
            \sigma^k := \left\{\sum_{i=0}^k\lambda_it_i:\sum_{i=0}^k\lambda_i = 1\text{ and }\lambda_i\geq0\right\}.
        \end{gather}
        where the points (\textbf{vertices}) $t_i\in\mathbb{R}^n$ are \textbf{affinely independent}, i.e. the vectors $t_i-t_0$ are linearly independent. Equivalently, a simplicial $k$-simplex  is the convex hull of the $k+1$ vertices $\{t_0,\ldots,t_k\}$.
    }
    \begin{remark}[Barycentric coordinates]
        The coordinates $\lambda_i$ from previous definition are called barycentric coordinates. This follows from the fact that the point $\sum_{i=0}^k\lambda_it_i$ represents the barycenter of a gravitational system consisting of masses $\lambda_i$ placed at the points $t_i$.
    \end{remark}
    \newdef{Standard simplex}{\index{simplex!standard}
        \begin{gather}
            \Delta^k := \left\{(x_0,\ldots,x_k)\in\mathbb{R}^{k+1}:\sum_ix_i = 1 \text{ and }x_i\geq0\right\}
        \end{gather}
    }

    \newnot{Face}{\index{face}
        Consider a simplicial $k$-simplex $[v_0,\ldots,v_k]$. The face opposite to the vertex $v_i$ is the simplicial $(k-1)$-simplex $[v_0,\ldots,\hat{v}_i,\ldots,v_k]$ obtained by removing the vertex $v_i$.
    }

    \newdef{Simplicial complex}{\index{simplicial complex}
        A simplicial complex $\mathcal{K}$ is a set of simplices satisfying following conditions:
        \begin{itemize}
            \item If $\sigma$ is a simplex in $\mathcal{K}$ then so are its faces.
            \item If $\sigma_1, \sigma_2\in\mathcal{K}$ then either $\sigma_1\cap\sigma_2 = \emptyset$ or $\sigma_1\cap\sigma_2$ is a face of both $\sigma_1$ and $\sigma_2$.
        \end{itemize}
        A simplicial $k$-complex is a simplicial complex where every simplex has dimension at most $k$.
    }

    \newdef{Path-connectedness}{\index{path!connected}
        Let $\mathcal{K}$ be a simplicial complex. $\mathcal{K}$ is said to be path-connected if every two vertices in $\mathcal{K}$ are connected by edges in $\mathcal{K}$.
    }

    \newdef{Polyhedron}{\index{polyhedron}
        Let $\mathcal{K}$ be a simplicial complex. The polyhedron associated with $\mathcal{K}$ is the topological space constructed by equipping $\mathcal{K}$ with the Euclidean subspace topology.
    }


    \newdef{Triangulable spaces}{\index{triangulation}
        Let $X$ be a topological space and let $\mathcal{K}$ be a polyhedron. If there exists a homeomorphism $\varphi:\mathcal{K}\rightarrow X$ then we say that $X$ is triangulable and we call $\mathcal{K}$ a \textbf{triangulation} of $X$.
    }
    \begin{theorem}
        Let $\mathcal{K}$ be a path-connected polyhedron with basepoint $a_0$. Let $\mathcal{C}\subset\mathcal{K}$ be a contractible 1-dimensional subpolyhedron containing all vertices of $\mathcal{K}$. Let $G$ be the free group generated by the elements $g_{ij}$ corresponding to the ordered 1-simplices $[v_i,v_j]\in\mathcal{C}$.

        The group $G$ is isomorphic to the fundamental group $\pi_1(\mathcal{K}, a_0)$ if the generators $g_{ij}$ satisfy following two relations:
        \begin{itemize}
            \item $g_{ij}g_{jk} = g_{ik}$ for every ordered 2-simplex $[v_i,v_j,v_k]\in\mathcal{K}\backslash\mathcal{C}$
            \item $g_{ij} = e$ if $[v_i,v_j]\in\mathcal{C}$.
        \end{itemize}
    \end{theorem}
    \begin{result}
        From the theorem that homeomorphic spaces have the same homotopy groups it follows that the fundamental group of a triangulable space can be computed by looking at its triangulations.
    \end{result}

    \begin{remark}[Hauptvermutung $\clubsuit$]
        Although the homology invariants (which we will define below) do not depend on the choice of triangulation we should give a remark about the existence of non-equivalent triangulations. Before the construction of a counterexample it was believed, hence the terms \textit{Hauptvermutung} in German or \textit{main conjecture} in English, that every two triangulations of a topological space allowed a common refinement and hence where equivalent for many constructions. However, it was shown that this conjecture is generally false, e.g. for topological manifolds in dimensions 5 and higher there exist an infinite number of non-equivalent triangulations. In dimensions below 4 it was proven by Rad\'o and Moise that the Hauptvermutung holds for topological manifolds (see theorem \ref{manifolds:rado_moise}).
    \end{remark}

\subsection{Simplicial homology}

    \newdef{Chain group}{\index{chain!group}\label{topology:chain_group}
        Let $\mathcal{K}$ be a simplicial $n$-complex. The $k^{th}$ chain group $C_k(\mathcal{K})$ is defined as the free Abelian group generated by the $k$-simplices in $\mathcal{K}$:
        \begin{gather}
            C_k(\mathcal{K}) := \left\{\sum_ia_i\sigma_i:\sigma_i\text{ is a $k$-simplex in $\mathcal{K}$}\text{ and }a_i\in\mathbb{Z}\right\}.
        \end{gather}
        For $k>n$ we define $C_k(\mathcal{K})$ to be $\{0\}$.
    }

    \newdef{Boundary operator}{\index{boundary}
        The boundary operator $\partial_k:C_k(\mathcal{K})\rightarrow C_{k-1}(\mathcal{K})$ is the group morphism defined by following properties:
        \begin{itemize}
            \item \textbf{Linearity}:
                \begin{gather}
                    \partial_k\left(\sum_ia_i\sigma_i\right) = \sum_ia_i\partial_k\sigma_i
                \end{gather}
            \item For every oriented $k$-simplex $[v_0,\ldots,v_k]$:
                \begin{gather}
                    \partial_k[v_0,\ldots,v_k] = \sum_{i=0}^k(-1)^i[v_0,\ldots,\hat{v}_i,\ldots,v_k]
                \end{gather}
            \item The boundary of every 0-chain is the identity 0.
        \end{itemize}
    }
    \remark{The alternating sum comes down the fact that we want the \textit{oriented} boundary.}

    \begin{property}
        The boundary operators satisfy the following relation:
        \begin{gather}
            \label{topology:boundary_operator_relation}
            \partial_k\circ\partial_{k+1} = 0.
        \end{gather}
        This property turns the system $(C_k, \partial_k)$ into a chain complex\footnote{See definition \ref{homalg:chain_complex}.}.
    \end{property}

    \newdef{Cycle group}{\index{cycle}
        The $k^{th}$ cycle group $Z_k(\mathcal{K})$ is defined as the set of $k$-chains $\sigma_k$ such that $\partial_k\sigma_k = 0$. These chains are called \textbf{cycles}.
    }
    \newdef{Boundary group}{
        The $k^{th}$ boundary group $B_k(\mathcal{K})$ is defined as the set of $k$-chains $\sigma_k$ for which there exists a $(k+1)$-chain $N$ such that $\partial_{k+1}N = \sigma_k$. These chains are called \textbf{boundaries}.
    }
    \newdef{Homology group}{\index{homology!simplicial}
        From property \ref{topology:boundary_operator_relation} it follows that $B_k(\mathcal{K})\subset Z_k(\mathcal{K})$ is a subgroup. We can thus define the $k^{th}$ homology group $H_k(\mathcal{K})$ as the following quotient group:
        \begin{gather}
            H_k(\mathcal{K}) := Z_k(\mathcal{K}) / B_k(\mathcal{K}).
        \end{gather}

        Theorem \ref{group:theorem:free_group} tells us that we can write $H_k(\mathcal{K})$ as $G_k\oplus T_k$. Both of these groups tell us something about $\mathcal{K}$. The rank of $G_k$, denoted by $R_k(\mathcal{K})$, is equal to the number of $(k+1)$-dimensional holes in $\mathcal{K}$. The torsion subgroup $T_k$ tells us how the space $\mathcal{K}$ is twisted.
    }

    \begin{property}
        If two topological spaces have the same homotopy type then they have isomorphic homology groups. It follows that homeomorphic spaces have isomorphic homology groups.
    \end{property}
    \begin{result}
        As was the case for the fundamental group, it follows from the definition of a triangulation that we can construct the homology groups for a given triangulable space by looking at its triangulations.
    \end{result}

    \newdef{Betti numbers}{\index{Betti number}
        The ranks $R_k(\mathcal{K})$ are called the Betti numbers of $\mathcal{K}$.
    }
    \begin{formula}[Euler characteristic]\index{Euler!characteristic}\index{Euler-Poincar\'e formula}
        The Euler characteristic of a triangulable space $X$ is defined as follows\footnote{This formula is sometimes called the \textit{Poincar\'e} or \textit{Euler-Poincar\'e} formula.}:
        \begin{gather}
            \chi(X) := \sum_i(-1)^iR_i(X).
        \end{gather}
    \end{formula}

    \begin{construct}
        The definition of homology groups can be generalized by letting the (formal) linear combinations used in the definition of the chain group (see \ref{topology:chain_group}) be of the following form:
        \begin{gather}
            c^k = \sum_ig_i\sigma_i^k
        \end{gather}
        where $G = \{g_i\}$ is an Abelian group and $\sigma_i^k$ are $k$-simplices. The $k^{th}$ homology group of $X$ with coefficients in $G$ is denoted by $H_k(X; G)$.
    \end{construct}
    \begin{property}
        When $G$ is a field, such as $\mathbb{Q}$ or $\mathbb{R}$, the torsion subgroups $T_k$ vanish. The relation between integral homology and homology with coefficients in a group is given by the \textit{Universal coefficient theorem}.
    \end{property}

    \newformula{K\"unneth formula}{\index{K\"unneth formula}
        Let $X, Y$ be two triangulable spaces. The homology groups of the Cartesian product $X\times Y$ with coefficients in a field $F$ is given by
        \begin{gather}
            H_k(X\times Y; F) = \bigoplus_{k = i+j}H_i(X;F)\otimes H_j(Y;F).
        \end{gather}
    }
    \begin{remark*}
        When the requirement of $F$ being a field is relaxed to it merely being a group, the torsion subgroups have to be taken into account. See the literature for this general form.
    \end{remark*}

\subsection{Relative homology}

    In this section we consider a simplicial complex $K$ and a subcomplex $L$.

    \newdef{Relative chain group}{\index{chain}
        The $k$-chain group of $K$ modulo $L$ is defined as the following quotient group:
        \begin{gather}
            C_k(K, L) := C_k(K) / C_k(L).
         \end{gather}
    }
    \newdef{Relative boundary operator}{\index{boundary}
        The relative boundary operator $\overline\partial_k$ is defined as follows (denoting equivalence classes by $[c]$):
        \begin{gather}
            \overline\partial_k[c_k] := [\partial_k c_k]
        \end{gather}
        where $c_k\in C_k(K)$. Because $\partial_kC_k(L)\subseteq C_{k-1}(L)$ this is a well-defined operation.
    }
    \newdef{Relative homology groups}{\index{homology!relative}
        The relative cycle and relative boundary groups are defined analogous to their ordinary counterparts. The relative homology groups are then defined as follows:
        \begin{gather}
            H_k(K, L) := \stylefrac{\ker(\overline\partial_k)}{\im(\overline\partial_{k+1})}.
        \end{gather}
        Elements $h_k\in H_k(K, L)$ can be represented as $h_k=z_k + C_k(L)$ such that $\partial_kz_k\in C_{k-1}(L)$.
    }

    \begin{property}
        Consider two topological spaces $X, Y$ with subspaces $A\subset X, B\subset Y$. If a continuous function $f:X\rightarrow Y$ is a homotopy equivalence such that the restriction to $A$ gives a homotopy equivalence $f|_A:A\rightarrow B$ then ht relative homology groups $H_k(X, A)$ and $H_k(Y, B)$ are isomorphic.
    \end{property}

    \newdef{Homology sequence}{
        Using the relative homology groups we obtain following (long) exact sequence:
        \begin{gather}
            \cdots\rightarrow H_k(L)\xrightarrow{i_\ast}H_k(K)\xrightarrow{j_\ast}H_k(K, L)\xrightarrow{\partial_k}H_{k-1}(L)\rightarrow\cdots
        \end{gather}
        where $i_\ast$ and $j_\ast$ are the homology morphisms induced by the inclusions $i:L\rightarrow K$ and $j:K\rightarrow (K, L)$.
    }

    \begin{theorem}[Excision theorem]\index{excision theorem}\label{topology:theorem:excision}
        Let $U, V$ and $X$ be triangulable spaces such that $U\subset V\subset X$. If the closure $\overline{U}$ is contained in the interior $V\degree$ then
        \begin{gather}
            H_k(X, V) = H_k(X\backslash U, V\backslash U).
        \end{gather}
    \end{theorem}

\subsection{Examples}

    \begin{example}
        Let $X$ be a contractible space.
        \begin{gather}
            H_k(X) =
            \begin{cases}
                \mathbb{Z}&k=0\\
                0&\text{otherwise}
            \end{cases}
        \end{gather}
    \end{example}
    \begin{example}
        Let $P$ be a path-connected polyhedron (or path-connected triangulable space).
        \begin{gather}
            H_0(P) = \mathbb{Z}
        \end{gather}
        Furthermore, every point $p\in P$ determines a generator $\langle p \rangle\in H_0(P)$.
    \end{example}

    \begin{example}
        The homology groups of the $n$-sphere $S^n$ are given by
        \begin{gather}
            H_k(S^n)=
            \begin{cases}
                \mathbb{Z}&k=0\text{ or }k=n\\
                0&\text{otherwise}.
            \end{cases}
        \end{gather}
    \end{example}
    \begin{example}
        Consider the closed (or open) disks $D_n$.
        \begin{gather}
            H_k(D_n, \partial D_n) =
            \begin{cases}
                \mathbb{Z}&k=n\\
                0&\text{otherwise}
            \end{cases}
        \end{gather}
        This holds for any $n$-skeleton $X^n$ of a CW-complex.
    \end{example}

    \newdef{Homology sphere}{\index{homology!sphere}
        A $n$-dimensional manifold having the same homology groups as the $n$-sphere.
    }
    \newdef{Degree}{\index{degree}
        From the example above we know that $H_n(S^n)=\mathbb{Z}$. Given a map $f:S^n\rightarrow S^n$ the induced map $f_*$ on homology is an endomorphism of $\mathbb{Z}$ and hence is of the form $f(x)=dx$ where $d\in\mathbb{Z}$. This coefficient is called the degree of $f$.
    }
    \begin{property}
        Two maps $f:S^n\rightarrow S^n$ have the same degree if and only if they are homotopic.
    \end{property}

    \begin{example}
        Consider a closed, connected and orientable manifold $M$ with $\dim(M) = n$.
        \begin{gather}
            H_n(M) = \mathbb{Z}
        \end{gather}
    \end{example}
    \begin{result}[Orientation]\index{orientation}\index{fundamental!class}
        A choice of orientation of $M$ coincides with a choice of generator for $H_n(M)$. This generator is called the \textbf{fundamental class}. In the case $M$ is disconnected, the fundamental class equals the direct sum of the generators of the connected components\footnote{Following the idea of the additivity axiom (see \ref{topology:eilenberg_steenrod_axioms}).}.

        ?? COMPLETE (LINK WITH DIFFGEOM, EXPAND, ...) ??
    \end{result}

\section{Singular homology}\label{section:singular_homology}

    \newdef{Singular simplex}{\index{simplex!singular}
        Consider the \textbf{standard} $k$-simplex $\Delta^k$. A singular $k$-simplex in a topological space $X$ is defined as a continuous map $\sigma^k:\Delta^k\rightarrow X$.
    }
    \sremark{The name singular comes from the fact that the maps $\sigma^k$ need not be injective.}

    \newdef{$\Delta$-complex}{\index{$\Delta$-complex}
        Let $\{\sigma_\alpha:\Delta^n\rightarrow X\}$ be a collection of morphisms from simplices to $X$ where the dimension $n$ may depend on the subscript $\alpha$. This collection forms a $\Delta$-complex on $X$ if it satisfies the following conditions:
        \begin{itemize}
            \item The restrictions $\sigma_\alpha|_{\overset{\circ}{\Delta}\,^n}$ are injective and every point in $X$ lies in the image of exactly one such restriction.
            \item The restriction of a morphism $\sigma_\alpha$ to the any one of the faces of $\Delta^n$ is equal to some other $\sigma_\beta$.
            \item A set in $X$ is open if and only if it is open in all the inverse images $\sigma_\alpha^{-1}$.
        \end{itemize}

        Similar to a CW-complex or cellular complex these conditions imply that every $\Delta$-complex can be constructed inductively from a (discrete) set of vertices by gluing and identifying edges.
    }

    \newdef{Singular chain group}{\index{chain}
        The singular chain group $S_k(X)$ with coefficients in a group $G$ is defined as the set of formal linear combinations $\sum_ig_i\sigma_i^k$. The basis of this free group is in most cases infinite as there are multiple ways to map $\Delta^k$ to $X$.
    }

    Before continuing we first need to introduce an important concept in the context of simplicial complexes:
    \newdef{Face map}{\index{face!map}
        The face maps are morphisms $\varepsilon_i^k:\Delta^{k-1}\rightarrow\Delta^k$ that map $\Delta^{k-1}$ onto the i$^{th}$ face of $\Delta^k$. They are explicitly given by
        \begin{gather}
            \varepsilon_i^k(s_0,\ldots,s_{k-1}) := (s_0,\ldots,s_{i-1},0,s_i,\ldots,s_{k-1}).
        \end{gather}
        Their defining property is the following relation:
        \begin{gather}
            \varepsilon_i^k\circ\varepsilon_j^{k-1} = \varepsilon_j^k\circ\varepsilon_{i-1}^{k-1}
        \end{gather}
        where $j\leq i$.
    }
    \sremark{Some authors (for example the authors at nLab) call these maps \textit{degeneracy maps} and call what in these notes are called \textit{degeneracy maps} face maps. This is sometimes the consuequence of working in a dual picture.}

    \newdef{Singular boundary operator}{\index{face!map}
        The singular boundary operator $\partial$ (we use the same notation as for simplicial boundary operators) is defined by its linear action on the singular chain group $S_k(X)$. It follows that we only have to know the action on the singular simplices $\sigma^k$.

        The action of the boundary operator on the singular simplex $\sigma^k$ is then given by
        \begin{gather}
            \partial_k\sigma^k = \sum_{i=0}^k(-1)^i\sigma^k\circ\varepsilon^k_i
        \end{gather}
        where the $\varepsilon^k_i$ are the face maps defined above. The singular boundary operators satisfy the same relation as the the simplicial boundary operators:
        \begin{gather}
            \partial_{k-1}\circ\partial_k = 0
        \end{gather}
        and hence define a chain complex.
    }

    \newdef{Singular homology group}{\index{homology!singular}
        The singular homology groups are defined as follows:
        \begin{gather}
            H_k(X; G) := \stylefrac{\ker(\partial_k)}{\im(\partial_{k+1})}.
        \end{gather}
    }

    \begin{theorem}
        For triangulable spaces singular homology is isomorphic to simplicial homology.
    \end{theorem}
    \begin{remark}
        When $X$ is not triangulable the previous theorem is not valid. The singular approach to homology is a more general construction, but it is often more difficult to compute the homology groups (even in the case of triangulable spaces).
    \end{remark}

    \begin{property}[Induced morphism]\index{pushforward}
        Consider a continuous map $f:X\rightarrow Y$ between topological spaces. This induces a map $f_k^S:S^k(X)\rightarrow S^k(Y)$ on the chain groups as follows:
        \begin{gather}
            f_k^S\left(\sum_\sigma c_\sigma\sigma\right) := \sum_\sigma c_\sigma f\circ\sigma.
        \end{gather}
        This map takes cycle (resp. boundary) groups to (subgroups of) cycle (resp. boundary) groups and hence induces a morphism of homology groups\footnote{These induced morphisms are also called \textbf{pushforwards}.}:
        \begin{gather}
            f_\ast:H_k(X)\rightarrow H_k(Y):\langle h \rangle\mapsto \langle f_k^S(h) \rangle.
        \end{gather}
    \end{property}
    \begin{result}
        $H_k$ is a functor $\mathbf{Top}\rightarrow\mathbf{Ab}$ that maps topological spaces to their homology groups and continuous maps $f$ to their pushforward $f_\ast$.
    \end{result}

    \begin{theorem}[Hurewicz]\index{Hurewicz}
        Let $X$ be path-connected. Let $[\cdot]$ and $\langle\cdot\rangle$ denote the equivalence classes in the homotopy and homology groups respectively. Then the map\footnote{Every path (and hence loop) is essentially a singular 1-cycle.} $h:\pi(X)\rightarrow H_1(X):[\gamma]\mapsto\langle\gamma\rangle$ defines a group morphism. Furthermore, this map induces an isomorphism $h':\pi(X)/[\pi(X), \pi(X)]\rightarrow H_1(X)$.

        More generally for every topological space $Y$ and every $n\in\mathbb{N}$ there exists a morphism $h_*:\pi_n(Y)\rightarrow H_n(Y)$. If $Y$ is $(n-1)$-connected then for every $k\leq n$ this morphism is in fact an isomorphism.
    \end{theorem}

    \begin{property}
        Let $X$ be a CW-complex. There exists an isomorphism $[X, K(G, n)]\rightarrow H^n(X; G)$ between the homotopy classes of maps $X\rightarrow K(G, n)$ and the $n^{th}$ singular cohomology of $X$ with coefficients in $G$.

        This morphism takes a map $f$ to the pullback $f^*\psi$ where $\psi$ is the inverse of the Hurewicz isomorphism $\pi_n(K)\rightarrow H_n(K;\mathbb{Z})$ and using the isomorphism $G=H^n(K;G)\cong\text{Hom}(H_n(K;\mathbb{Z}), G)$.
    \end{property}

\section{Axiomatic approach}

    \newdef{Eilenberg-Steenrod axioms}{\index{Eilenberg-Steenrod axioms}\label{topology:eilenberg_steenrod_axioms}
        All homology theories have a set of properties in common. By treating these properties as axioms we can construct homology theories as a sequence of functors $H_k: \text{Top}\times\text{Top}\rightarrow\text{Ab}$. The axioms are as follows:
        \begin{enumerate}
            \item \textbf{Homotopy}: If $f, g$ are homotopic maps then their induced homology maps are the same: \[f\cong g\implies H_k(f) = H_k(g), \forall k\in\mathbb{N}.\]
            \item \textbf{Excision}\footnote{See also theorem \ref{topology:theorem:excision}.}: If $U\subset V\subset X$ and $\overline U\subset V\degree$ then $H_k(X, V) \cong H_k(X\backslash U, V\backslash U)$.
            \item \textbf{Dimension}: If $X$ is a singleton then $H_k(X) = \{0\}$ for all $k\geq1$. The group $H_0(X)$ is called the \textbf{coefficient group} and gives the coefficients used in the linear combinations of the chain group.
            \item \textbf{Additivity}: If $X = \bigsqcup_i X_i$ then $H_k(X)\cong\bigoplus_i H_k(X_i)$.
            \item \textbf{Exactness}: Each pair $(X, A)$, where $A\subset X$, induces a long exact sequence
            \begin{gather}
                \cdots\rightarrow H_k(A)\xrightarrow{i_\ast}H_k(X)\xrightarrow{j_\ast}H_k(X, A)\xrightarrow{\partial_k}H_{k-1}(A)\rightarrow\cdots
            \end{gather}
            where $i_\ast$ and $j_\ast$ are the homology morphisms induced by the inclusions $i:A\rightarrow X$ and $j:X\rightarrow (X, A)$.
        \end{enumerate}
    }
    \begin{remark}
        If the dimension axiom is removed from the set of axioms then we obtain a so-called \textit{extraordinary homology theory}.
    \end{remark}

\section{Equivariant cohomology}

    In this section we will consider topological spaces equipped with a continuous action of a topological group $G$. These spaces will be called topological $G$-spaces or just $G$-spaces.

    \newdef{Equivariant cohomology}{
        Let $X$ be a topological $G$-space for which the $G$-action is free. The equivariant cohomology of $X$ is defined as
        \begin{gather}
            H_G^*(X) := H^*(X/G)
        \end{gather}
        where the $X/G$ is the orbit space with respect to the action of $G$ on $X$.
    }

    ?? COMPLETE (E.G. LECTURES OF TU ON YOUTUBE) ??

\section{\texorpdfstring{Rational homotopy theory $\clubsuit$}{Rational homotopy theory}}

    For the theory on differential graded algebras see chapter \ref{chapter:hda}.

    \newdef{Rational space}{\index{rational!space}
        A simply connected topological space $X$ such that the homotopy groups $\pi_n(X)$ are rational vector spaces.
    }

    \newdef{Rational homotopy equivalence}{\index{rational!homotopy equivalence}
        A continuous function $f:X\rightarrow Y$ for which the induced maps on rational homotopy groups
        \begin{gather}
            \pi_n(f)\otimes\mathbb{Q}:\pi_n(X)\otimes\mathbb{Q}\rightarrow\pi_n(Y)\otimes\mathbb{Q}
        \end{gather}
        are isomorphisms for all $n\in\mathbb{N}$.
    }

    \newdef{Rational homotopy category}{
        Consider the category \textbf{Top} of topological spaces. The rational homotopy category is obtained as the localization of \textbf{Top} with respect to the collection of rational homotopy equivalences.
    }

    \newdef{Polynomial differential forms}{
        Consider the standard $n$-simplex $\Delta^n$ (as a simplicial set \ref{sheaf:standard_simplex}). On this set (and its geoemtric realisation as a manifold with corners) one can define differential forms analogous to those from section \ref{diff:section:forms}: Let $\{t_i\}_{0\leq i\leq n}$ be $n+1$ generators in degree 0 (under geometric realisation these are identified with the barycentric coordinates of the topological $\Delta^n$). Together with $n+1$ associated generators $dt_i$ in degree 1, one can construct the free graded algebra (over $\mathbb{Q}$). To preserve the geometric structure of $\Delta^n$ (after geometric realisation) we have to quotient out the following relation:
        \begin{gather*}
            \sum_{i=0}^n t_i = 1.
        \end{gather*}
        To fully preserve the structure after geometric realisation one also has to quotient out the associated ''differential'' relation:
        \begin{gather*}
            \sum_{i=0}^n dt_i = 0.
        \end{gather*}
        The resulting graded algebra is denoted by $\Omega^\bullet_{\text{poly}}(\Delta^n)$. In degree 0 this complex can be identified with the space of polynomial functions on the topological $\Delta^n$. The ordinary differential forms can be obtained by taking the tensor product of $\Omega^\bullet_{\text{poly}}$ with the space of smooth functions on the topological $\Delta^n$. Under this isomorphism the generators $dt_i$ are identified with the de Rham differentials of the barycentric coordinates.

        Morphisms $f:[m]\rightarrow[n]$ in the simplex category $\Delta$ induce morphisms of simplices and these in turn induce morphisms $F:\Omega^\bullet_{\text{poly}}(\Delta^n)\rightarrow\Omega^\bullet_{\text{poly}}(\Delta^m)$ by the following action on generators:
        \begin{gather}
            F(t_i) := \sum_{f(j)=i}t_j.
        \end{gather}
        It can be seen that this turns the above construction into a functor $\func{\Omega^\bullet_{\text{poly}}}{\Delta^{op}}{dgcAlg}$. By going to the opposite functor and taking a left Kan extension along the Yoneda embedding $\Delta\hookrightarrow\mathbf{sSet}$ one obtains a functor $\func{\Omega^\bullet_{\text{poly}}}{sSet}{dgcAlg^{op}}$. Composition with the singular set functor $\func{\text{Sing}}{Top}{sSet}$ gives the \textbf{piecewise-polynomial differential forms} functor $\Omega^\bullet_{\text{pwpoly}}$.
    }

    \newdef{Relative Sullivan algebra}{\index{Sullivan!algebra}
        An inclusion of dgca's of the form \[(A, d)\hookrightarrow(A\otimes\wedge^\bullet V, d')\] where $A$ is any dgca and $V$ is a graded vector space such that
        \begin{itemize}
            \item there is a well-ordered set $J$ indexing a linear basis $\{e_i\}_{i\in J}$ of $V$.
            \item for all $k\in J$, $V_{<k}:=\text{span}\{e_i:i<k\}$ and all $e_k$ we have that
            \begin{gather}
                d'e_k\in A\otimes\Lambda^\bullet V_{<k}.
            \end{gather}
        \end{itemize}
        If in addition the implication
        \begin{gather}
            i\leq j\implies \deg e_i\leq \deg e_j
        \end{gather}
        holds for all $i,j\in J$ then the relative Sullivan algebra is said to be \textbf{minimal}.
    }
    \begin{remark}
        If the Sullivan algebra is defined relative to the tensor unit $(k, 0)$ (where $k$ is the underlying field) then it is just called a \textbf{Sullivan algebra}.
    \end{remark}

    \newdef{Sullivan model}{\index{Sullivan!model}
        Let $X$ be a simply-connected topological space. A (minimal) Sullivan model for $X$ is a (minimal) Sullivan algebra equipped with a quasi-isomorphism to the dga of piecewise-polynomial differential forms on $X$.
    }
    \begin{property}
        Minimal Sullivan models are unique up to isomorphism.
    \end{property}