\section{Rings}\label{section:ring}

    \newdef{Ring}{\index{ring}
        Let $R$ be a set equipped with two binary operations $+,\cdot$ (called addition and multiplication). $(R,+,\cdot)$ is a ring if it satisfies the following axioms:
        \begin{enumerate}
            \item $(R,+)$ is a commutative group.
            \item $(R,\cdot)$ is a monoid.
            \item Multiplication is distributive with respect to addition.
        \end{enumerate}
    }
    \newdef{Field}{\index{field}
        A ring $(R, +, \cdot)$ is said to be a field it the monoid $(R, \cdot)$ is multiplicative and if all nonzero elements admit a multiplicative inverse.
    }

    \newdef{Unit}{\index{unit}
        An invertible element of a ring $(R, +, \cdot)$. The set of units forms a group under multiplication.
    }

    \newdef{Integral domain}{\index{domain!integral}\label{alg:integral_domain}
        A commutative ring $R$ in which the product of two nonzero elements is again nonzero.
    }

    \newdef{Reduced ring}{\index{reduced ring}
        A ring that contains no nonzero nilpotents.
    }

    \begin{construct}[Localization]\index{localization}
        Let $R$ be a commutative ring and let $S$ be a multiplicative monoid in $R$. We first define an equivalence relation $\sim$ on $R\times S$ in the following way:
        \begin{equation}
            (r_1, s_1)\sim(r_2, s_2) \iff \exists t\in S: t(r_1s_2 - r_2s_1) = 0.
        \end{equation}

        The set $R^* := (R\times S)/\sim$, called the localization of $R$ with respect to $S$, can now be turned into a ring by defining an addition and a multiplication: By writing $(r, s)\in R^*$ as the formal fraction $\frac{r}{s}$ we define these operations by analogy with the those of ordinary fractions:
        \begin{itemize}
            \item $\displaystyle\frac{r_1}{s_1} + \frac{r_2}{s_2} = \frac{r_1s_2 + r_2s_1}{s_1s_2}$
            \item $\displaystyle\frac{r_1}{s_1}\cdot\frac{r_2}{s_2} = \frac{r_1r_2}{s_1s_2}$.
        \end{itemize}
    \end{construct}
    \remark{The localization of $R$ with respect to the monoid $S$ can be interpreted as the ring obtained by collapsing $S$ into a single unit of $R$.}

    \begin{notation}\label{alg:localization_notation}
        The localization of $R$ with respect to $S$ is often denoted by $S^{-1}R$. For specific cases different notations are sometimes used. For example choose an element $f\in R$, then $R_f$ denotes the localization of $R$ with respect to the set of powers of $f$, i.e. $S=\{f^n:n\in\mathbb{N}\}$. This is called the \textbf{localization at the element} $f$. Another example occurs when working with prime ideals: Let $P$ be a prime ideal, then it is not hard to show that its complement $R\backslash P$ is multiplicatively closed. The localization of $R$ with respect to this set is denoted by $R_P$ and called the \textbf{localization at the prime ideal} $P$.
    \end{notation}

    \newdef{Valuation}{\index{valuation}
        Let $K$ be a field and let $\Gamma$ be a totally ordered\footnote{See definition \ref{group:total_order}.} Abelian group. First we extend the group law and the order relation on $\Gamma$ to the union $\Gamma\cup\{\infty\}$ in the following way:
        \begin{itemize}
            \item $g+\infty = \infty + g=\infty$ for all $g\in\Gamma$, and
            \item $g\leq\infty$ for all $g\in\Gamma$.
        \end{itemize}
        A valuation on $K$ (with values in $\Gamma$) is a map $\nu:K\rightarrow\Gamma\cup\{\infty\}$ such that:
        \begin{enumerate}
            \item $\nu(a) = \infty \iff a = 0$,
            \item $\nu(ab) = \nu(a) + \nu(b)$, and
            \item $\min(\nu(a),\nu(b))\leq\nu(a+b)$ where the equality holds if $\nu(a)\neq\nu(b)$.
        \end{enumerate}
    }

\subsection{Ideals}

    \newdef{Ideal}{\index{ideal}\label{linalgebra:ideal}
        Let $(R,+,\cdot)$ be a ring with $(R,+)$ its additive group. A subset $I\subseteq R$ is called an ideal\footnote{More generally, this is called a \textbf{two-sided ideal}.} of $R$ if it satisfies the following conditions:
        \begin{enumerate}
            \item $(I,+)$ is a subgroup of $(R,+)$.
            \item $\forall n\in I, \forall r\in R:n\cdot r, r\cdot n\in I$.
        \end{enumerate}
    }

    \newdef{Artinian ring}{\index{Artin!ring}
        A ring is said to be Artinian (or to be an Arting ring) if it satisfies the descending chain condition on ideals, i.e. it contains no infinite descending chain\footnote{See definition \ref{set:chain}.} of ideals.
    }
    \newdef{Noetherian ring}{\index{Noether!ring}
        A ring is said to be Noetherian (or, although less often, to be a Noether ring) if it satisfies the ascending chain condition on ideals, i.e. it contains no infinite ascending chain of ideals.
    }

    \newdef{Simple ring}{\index{simple!ring}
        A ring is simple if it has no nontrivial two-sided ideals.\footnote{Some authors require the ring to be Artinian.}
    }

    \newdef{Unit ideal}{
        A ring considered as an ideal of itself.
    }
    \newdef{Proper ideal}{
        Let $R$ be a ring. A subset $I\subset R$ is said to be a proper ideal if it is an ideal of $R$ and if it is not equal to $R$.
    }
    \newdef{Prime ideal}{
        Let $R$ be a ring. A proper ideal $I$ is a prime ideal if for any $a,b\in R$ the following relation holds:
        \begin{equation}
            ab\in I\implies \text a\in I \vee b\in I.
        \end{equation}
    }
    \newdef{Maximal ideal}{
        A proper ideal $I$ such that there exists no other proper ideal $T$ such that $I\subset T$.
    }
    \newdef{Minimal ideal}{
        A proper ideal that contains no other nonzero ideals.
    }

    \begin{property}
        Every maximal ideal is prime.
    \end{property}

    \newdef{Jacobson radical}{\index{Jacobson radical}
        The Jacobson radical of a ring $R$, often denoted by $J(R)$, is the ideal obtained as the intersection of all maximal left (or right) ideals. Equivalently, one can define it as the intersection of the \textit{annihilators} of all left (or right) simple $R$-modules.
    }

    \begin{construct}[Generating set of an ideal]\index{ideal!generating set}\label{group:generating_set_ideal}
        Let $R$ be a ring and let $X$ be a subset of $R$. The two-sided ideal generated by $X$ is defined as the intersection of all two-sided ideals containing $X$. An explicit construction is given by
        \begin{equation}
            I = \left\{\sum_{i=1}^n l_ix_ir_i:n\in\mathbb{N}, \forall i\leq n:l_i, r_i\in R\land x_i\in X\right\}.
        \end{equation}
        Left and right ideals are generated in a similar fashion.
    \end{construct}
    \begin{notation}
        Let the ideal $I$ be generated by the elements $\{f_j\}_{j\in J}$ (for some index set $J$). In general this is written as
        \begin{gather}
            I = (f_1, f_2, \ldots).
        \end{gather}
    \end{notation}

    \newdef{Principal ideal}{\index{ideal!principal}
        An ideal which is generated by a single element.
    }
    \newdef{Principal ideal domain}{\index{domain!principal ideal}
        An integral domain \ref{alg:integral_domain} in which every ideal is principal.
    }

    \begin{construct}[Extension]\index{extension!ideal}
        Let $I$ be an ideal of a ring $R$ and let $\iota:R\rightarrow S$ be a ring morphism. The extension of $I$ with respect to $\iota$ is the ideal generated by the set $\iota(I)$.
    \end{construct}

    \newdef{Local ring}{\index{local!ring}\label{algebra:local_ring}
        A ring for which a unique maximal (left) ideal exists.\footnote{This also implies that there exists a unique maximal right ideal and that these ideals coincide.}
    }
    \begin{property}[Characterization by invertible complements]\label{algebra:local_ring_invertible}
        A ring $R$ is local if and only if there exists a maximal ideal $M$ such that every element in the complement $R\backslash M$ is invertible.
    \end{property}

    \begin{property}[Prime localization]\label{algebra:localization_local_ring}
        The localization of a ring $R$ with respect to a prime ideal $P$ is a local ring, where the maximal ideal is given by the extension of $P$ with respect to the ring morphism $\iota:R\rightarrow R_P$. Equivalently, this says that the maximal ideal is given by $PR_P$.
    \end{property}

    \newdef{Residue field}{\index{residue!field}
        Consider a local ring $R$ and let $I$ be its maximal ideal. The quotient ring $R/I$ forms a field, called the residue field.
    }

\subsection{Modules}

    \newdef{$R$-Module}{\index{module!over a ring}
        Let $(R, +, \cdot)$ be a ring. A set $X$ is said to be an $R$-module if it satisfies the same axioms as those of a vector space \ref{linalgebra:vector_space} but where the scalars are only elements of a ring instead of a field.
    }

    \newdef{Free module}{\index{free!module}
        A module is said to be free if it admits a basis.
    }

    \begin{property}[Division rings]\index{division ring}\label{algebra:module_basis}
        For a general $R$-module the existence of a basis is not guaranteed unless $R$ is a \textit{division ring}. (See construction \ref{linalgebra:hamel_basis} for more information.)
    \end{property}
    \begin{result}
        Since every field is in particular a division ring, the existence of a basis follows from the above property for $R$-modules.
    \end{result}

    \newdef{Projective module}{\index{projective!module}
        A module $P$ is said to be projective if $P$ can be expressed as
        \begin{equation}
            P\oplus M = F
        \end{equation}
        where $M$ is a module and $F$ is a free module.
    }

    \begin{example}[Dual numbers]\index{dual!numbers}
        Let $R$ be a ring. The $R$-algebra of dual numbers (over $R$), often denoted by $R[\varepsilon]$, is defined as the free $R$-module with basis $\{1, \varepsilon\}$ subject to the relation $\varepsilon^2 = 0$.
    \end{example}

\subsection{Semisimplicity}\index{semisimple!module}

    \newdef{Simple module}{\index{simple!module}
        A module over a ring is said to be simple if it contains no nontrivial submodules. A module is said to be \textbf{semisimple} if it admits a direct sum decomposition in terms of simple modules.
    }
    \newdef{Semisimple ring}{
        A ring is said to be semisimple if it is semisimple as a module over itself.
    }
    \begin{property}[Condition]
        A ring is semisimple if and only if it is Artinian and if its Jacobson radical vanishes.
    \end{property}

    \begin{theorem}[Artin-Wedderburn]\index{Artin-Wedderburn}\label{algebra:artin_wedderburn}
        Every semisimple ring is isomorphic to a direct sum of matrix rings over division rings $D_i$ with multiplicity $n_i$. Furthermore, the integers $D_i$ and $n_i$ are unique\footnote{Up to a permutation of the indices.}.
    \end{theorem}

\subsection{Graded rings}

    \newdef{Graded ring}{\index{graded}\label{group:graded_ring}
        Let $R$ be a ring that can be written as a direct sum of Abelian groups $A_k$:
        \begin{equation}
            R = \bigoplus_{k\in\mathbb{N}}A_k.
        \end{equation}
        If the groups $A_k$ in this decomposition have the property that for all $i, j\in\mathbb{N}: A_i\star A_j\subseteq A_{i+j}$, where $\star$ is the ring multiplication, then $R$ is said to be a graded ring. The elements of the groups $A_k$ are said to be \textbf{homogeneous of degree $k$}.
    }

    \newformula{Graded commutativity}{\index{commutativity!graded}
        Consider two elements $v,w$ of a graded ring $R$. If
        \begin{equation}
            \label{algebra:graded_commutativity}
            vw = (-1)^{\deg(v)\deg(w)}wv
        \end{equation}
        for all elements $v, w$ of the graded ring, then the ring is said to be graded-commutative.
    }

\section{Limits of algebraic structures}

    \newdef{Direct system}{\index{direct!system}
        Let $(I,\leq)$ be a directed set \ref{set:directed_set} and let $\{A_i\}_{i\in I}$ be a family of algebraic objects (groups, rings, ...). Consider a collection of morphisms $\{f_{ij}:A_i\rightarrow A_j\}_{i,j\in I}$ between these objects with the following properties:
        \begin{itemize}
            \item for every $i\in I$: $f_{ii} = \mathbbm{1}_{A_i}$, and
            \item for every $i\leq j\leq k\in I$: $f_{ik} = f_{jk}\circ f_{ij}$.
        \end{itemize}
        The pair $(A_i,f_{ij})$ is called a direct system (over $I$).
    }

    \newdef{Direct limit\footnotemark}{\index{direct!limit}\index{inductive!limit}\label{set:direct_limit}
        \footnotetext{Also called an \textbf{inductive limit}.}
        Consider a direct system $(A_i,f_{ij})$ over a directed set $I$. The direct limit $A$ of this direct system is defined as follows:
        \begin{gather}
            \varinjlim A_i := \left.\bigsqcup_{i\in I}A_i\right/\sim
        \end{gather}
        where the equivalence relation is given by $x\in A_i\sim y\in A_j\iff\exists k\in I: f_{ik}(x) = f_{jk}(y)$. Informally put: two elements are equivalent if they eventually become the same.

        The algebraic operations on $A$ are defined such that the inclusion maps $\phi_i:A_i\rightarrow A$ are morphisms.
    }

    \newdef{Inverse system}{\index{inverse!system}
        Let $(I,\leq)$ be a directed set \ref{set:directed_set} and let $\{A_i\}_{i\in I}$ be a family of algebraic objects (groups, rings, ...). Consider a collection of morphisms $\{f_{ij}:A_j\rightarrow A_i\}_{i,j\in I}$ between these objects with the following properties:
        \begin{itemize}
            \item for every $i\in I$: $f_{ii} = \mathbbm{1}_{A_i}$, and
            \item for every $i\leq j\leq k\in I$: $f_{ik} = f_{ij}\circ f_{jk}$.
        \end{itemize}
        The pair $(A_i,f_{ij})$ is called an inverse system (over $I$).
    }

    \newdef{Inverse limit\footnotemark}{\index{inverse!limit}\index{projective!limit}\label{set:inverse_limit}
        \footnotetext{Also called a \textbf{projective limit}.}
        Consider an inverse system $(A_i,f_{ij})$ over a directed set $I$. The inverse limit $A$ of this inverse system is defined as follows:
        \begin{gather}
            \varprojlim A_k := \left\{\vec{a}\in\prod_{i\in I}A_i:a_i=f_{ij}(a_j), \forall i\leq j\right\}.
        \end{gather}
        For all $i\in I$ there exists a natural projection $\pi_i:\varprojlim A_k\rightarrow A_i$.
    }

    \begin{remark}
        The direct and inverse limit are each other's (categorical) dual. The former is a colimit while the latter is a limit in category theory.
    \end{remark}

\section{Galois theory}

    \newdef{Field extension}{\index{field!extension}
        Let $k$ be a field. A field extension of $k$ is a field $K$ such that $k\subset K$ and such that the operations of $k$ coincide with the ones of $K$ restricted to $k$.
    }
    \begin{notation}
        A field extension $K$ of $k$ is often denoted by $K/k$.
    \end{notation}

    \newdef{Degree}{\index{degree!of extension}
        If $L/K$ is a field extension then $L$ can be given the structure of a $K$-vector space \ref{linalgebra:vector_space}. The dimension of this vector space is called the degree of the extension $L$. It is often denoted by $[L:K]$.
    }