\subsection{Homomorphisms}
	\newdef{Homomorphism space}{\index{homomorphism!of vector spaces}
    	Let $V,W$ be two K-vector spaces. The set of all linear maps between $V$ and $W$ is called the homomorphism space of $V$ to $W$, or shorter: the 'hom-space' of $V$ to $W$.
    	\begin{equation}
        	\label{linalgebra:hom_space}
            Hom_K(V, W) = \{f:V\rightarrow W\ |\ \text{f is a linear map}\}
		\end{equation}
	}
    \begin{theorem}
    	\label{linalgebra:hom_dimension}
    	If $\ V,W$ are two finite-dimensional K-vector spaces we have:
        \begin{equation}
            \dim\left(Hom_K(V,W)\right) =\dim(V)\cdot\dim(W)
		\end{equation}
	\end{theorem}
        
	\newdef{Endomorphism ring}{The space $Hom_K(V, V)$ with the composition as multiplication forms a ring, the endomorphism ring. It is denoted as $\text{End}_K(V)$ or $\text{End}(V)$.}
    
    \newdef{Minimal polynomial}{\index{minimal polynomial}
    	Let $f\in\text{End}(V)$ and $V$ a finite-dimensional K-vector space. The monic polynomial $\mu_f(x)$ of the lowest order such that $\mu_f(f)=0$ is called the minimal polynomial of $f$.
	}
    \begin{property}
    	\label{linalgebra:minimal_polynomial_divisor}
		Let $f\in\text{End}(V)$. Let $\mu_f(x)$ be the minimal polynomial of $f$. Let $\varphi(x)\in K[x]$. If $\varphi(f) = 0$, then the minimal polynomial $\mu_f(x)$ divides $\varphi(x)$. 
	\end{property}
    
\subsection{Dual space}
	\newdef{Dual space}{\index{dual!space}
    	Let $V$ be a K-vector space. The dual space $V^*$ of $V$ is the following vector space:
    	\begin{equation}
			\label{linalgebra:dual_space}
            V^*:=\text{Hom}_K(V, K)=\{f:V\rightarrow K\ :\ f\text{ is a linear map}\}
		\end{equation}
    }
    \newdef{Linear form}{\index{linear!form}
    	The elements of $V^*$ are called \textit{linear forms}.
	}
    \begin{property}
    	\label{linalgebra:dual_space_dimension}
		From theorem \ref{linalgebra:hom_dimension} it follows that $\dim(V^*) = \dim(V)$.
	\end{property}
    \remark{If $V$ is infinite-dimensional, theorem \ref{linalgebra:dual_space_dimension} is not valid. In the infinite-dimensional case we \textbf{always} have $|V^*|>|V|$ (where we now use the cardinality instead of the dimension).}
    
	\newdef{Dual basis}{
    		Let $\mathcal{B} = \{e_1, e_2, ..., e_n\}$ be a basis for a finite-dimensional K-vector space $V$. We can define a basis $\mathcal{B}^* = \{\varepsilon_1, \varepsilon_2, ..., \varepsilon_n\}$ for $V^*$, called the dual basis of $\mathcal{B}$, as follows:
    		\begin{equation}
			\label{linalgebra:dual_basis}
        		\boxed{\varepsilon_i:V\rightarrow K:\sum_{j=1}^na_ie_i\mapsto a_i}
		\end{equation}
	}
    
	\newdef{Dual map}{\index{dual!map}\label{linalgebra:transpose}
		Let $f:V\rightarrow W$ be a linear map. The linear map $f^*:W^*\rightarrow V^*:\varphi\rightarrow\varphi\circ f$ is called the dual map or \textbf{transpose} of $f$.
	}
	\newnot{Transpose}{\index{transpose}
		When $V=W$ the dual map $f^*$ is often denoted by $f^T$.
	}
    
    \newdef{Natural pairing}{\index{natural pairing}
    	The natural pairing of $V$ and its dual $V^*$ is defined as the following bilinear map:
        \begin{equation}
        	\label{linalgebra:natural_pairing}
            \langle v, v^*\rangle = v^*(v)
        \end{equation}
    }

\subsection{Convex functions}
	
	\newdef{Convex function}{\index{convex!function}
		Let $X$ be a convex subset of $V$. A function $f:X\rightarrow \mathbb{R}$ is convex if for all $x, y\in X$ and $t\in[0, 1]$:
		\begin{equation}
			f(tx + (1-t)y) \leq tf(x) + (1-t)f(y)
		\end{equation}
	}
	\begin{remark}
		For a concave function we have to turn the inequality around.
	\end{remark}
	\begin{result}
		A linear map $f:X\rightarrow\mathbb{R}$ is both convex and concave.
	\end{result}
	
	\begin{theorem}[Karamata's inequality]\index{Karamata's inequality}
		Let $I\subset\mathbb{R}$ be an interval and let $f:I\rightarrow\mathbb{R}$ be a convex function. If $(x_1, ..., x_n)$ is a tuple that majorizes $(y_1, ..., y_n)$, i.e. $\forall k\leq n$
		\begin{align}
			\sum_{i=1}^nx_i &= \sum_{i=1}^ny_i\\
			x_{(1)} + ... + x_{(k)}&\geq y_{(1)} + ... + y_{(k)}
		\end{align}
		where $x_{(i)}$ denotes the ordering\footnote{In decreasing order: $x_{(1)}\geq...\geq x_{(n)}$.} of the tuple $(x_1, ..., x_n)$. Then
		\begin{equation}
			\sum_{i=1}^nf(x_i)\geq \sum_{i=1}^nf(y_i)
		\end{equation}
	\end{theorem}
