\chapter{Linear Algebra}

\input{Vector/VectorSpaces}
\input{Vector/InnerProduct}
\input{Vector/Matrices}
\input{Vector/Eigenvectors}

\section{Euclidean space \texorpdfstring{$\mathbb{R}^n$}\ }

    A finite-dimensional $\mathbb{R}$-vector space is called a \textbf{Euclidean space}.

    \begin{notation}
        When working in a Euclidean space the inner product $\langle v|w\rangle$ is often written as $v\cdot w$ or even $vw$.
    \end{notation}

    \newdef{Orientation}{\label{linalgebra:orientation}
        Let $\mathcal{B}, \mathcal{B}'$ be two ordered bases of $\mathbb{R}^n$. Let $Q$ be the transition matrix from $\mathcal{B}$ to $\mathcal{B}'$. If $\det(Q)>0$ then the bases are said to have the same orientation (or to be \textbf{consistently oriented}). If $\det(Q)<0$ then the bases are said to have an opposite orientation.
    }
    \begin{result}[Positive orientation]
        The previous definition imposes an equivalence relation on the set of bases of $\mathbb{R}^n$. The set of bases consists out of two equivalence classes. Take one class and call the bases in it \textbf{positively} (or \textbf{directly}) oriented. The bases in the other class are then said to be \textbf{negatively} (or \textbf{indirectly}) oriented.
    \end{result}
    \remark{It is convenient to take the standard basis $(e_1,\dotso,e_n)$ to be positively oriented.}

\section{Graded vector spaces}\label{section:graded_spaces}

    Similar to definition \ref{group:graded_ring} we can define the following:
    \newdef{Graded vector space}{\index{degree}\index{graded!vector space}\label{linalgebra:graded_vector_space}
        Let $V_n$ be a vector space for all $n\in\mathbb{N}$. The vector space
        \begin{gather}
        V = \bigoplus_{n\in\mathbb{N}} V_n
        \end{gather}
        is called a graded vector space. In fact one can replace $\mathbb{N}$ by any countable (finite or infinite) index set. For most operations however one requires the index set to be closed under addition operations. The index $n$ is often called the \textbf{degree} of the subspace $V_n$ in $V$.
    }

    \newdef{Graded algebra}{\index{graded!algebra}
        Let $V$ be a graded vector space with the additional structure of an algebra given by the multiplication $\star$. Then $V$ is a graded algebra if $\star$ maps $V^k\times V^l$ to $V^{k+l}$.
    }

    \newdef{Super vector space}{\index{super!vector space}
        A $\mathbb{Z}_2$-graded vector space.
    }

    \begin{example}[Superalgebra]\index{super!algebra}\label{linalgebra:superalgebra}
        A $\mathbb{Z}_2$-graded algebra, i.e. there exists a decomposition
        \begin{gather}
        A = A_0\oplus A_1
        \end{gather}
        such that for all $i, j \mod 2$:
        \begin{gather}
        A_i\star A_j \subseteq A_{i+j}.
        \end{gather}
    \end{example}

    \newdef{dg-algebra}{\index{dg-algebra}\label{linalgebra:dg_algebra}
        A differential graded algebra (often denoted by dg-algebra) is a graded algebra equipped with a differential\footnote{See definition \ref{homalg:chain_complex}.} of degree 1.
    }

    \newdef{Parity functor}{\index{parity}
        Consider the category $\mathbf{sVect}$ of super vector spaces. We can define the parity functor $\func{\Pi}{sVect}{sVect}$ as the functor which interchanges even and odd subspaces:
        \begin{align}
        (\Pi V)_0 &= V_1\\
        (\Pi V)_1 &= V_0
        \end{align}
    }
    \newdef{Symmetric tensors}{
        Using the parity functor one can write the exterior algebra\footnote{See definition \ref{tensor:exterior_algebra}.} $\Lambda^\bullet(V)$ as the symmetric algebra $\text{Sym}^\bullet(\Pi V)$. In a similar way one can write the symmetric algebra on a super vector space $V=V_0\oplus V_1$ as \[\text{Sym}^n(V)=\bigoplus_{p+q=n}\text{Sym}^p(V_0)\otimes\Lambda^q(V_1).\]
    }

\subsection{Supermatrices}

    For this section we will relax the requirement that we are working over a field $K$, instead we will be working over a supercommutative ring. This means that we will be working with (graded) modules instead of true vector spaces.

    \newdef{Supermatrix}{\index{parity}
        Every linear transformation between super vector spaces $(V_0, V_1)$ and $(W_0, W_1)$ can be decomposed as the sum of 4 linear transformations between the even/odd subspaces:
        \begin{itemize}
            \item $A:V_0\rightarrow W_0$
            \item $B:V_1\rightarrow W_0$
            \item $C:V_0\rightarrow W_1$
            \item $D:V_1\rightarrow W_1$
        \end{itemize}
        If we represent these transformations as matrices, then the full transformation can be represented as a block matrix \[X=\begin{pmatrix}A&B\\C&D\end{pmatrix}\] where we used the same notation to denote the matrices associated to the linear transformations.

        These matrices can be decomposed according to their \textbf{parity}. Not all supermatrices preserve the grading, or equivalently, not all linear transformations of super vector spaces are morphisms of super vector spaces. The ones that are, are said to have even parity and they are of the form \[X=\begin{pmatrix}\text{even}&\text{odd}\\\text{odd}&\text{even}\end{pmatrix}\] where by even/odd we mean that the entries in these blocks have even/odd parity as elements of the underlying (graded) ring. It should be clear that these matrices indeed preserve the grading, since acting with an odd scalar on an odd vector gives an even vector (and similar for the other combinations). The matrices that do not preserve the grading are said to have odd parity and are of the form \[X=\begin{pmatrix}\text{odd}&\text{even}\\\text{even}&\text{odd}\end{pmatrix}.\]
    }

    \newdef{Supertrace}{\index{trace}
        The supertrace of a supermatrix generalizes the trace of an ordinary matrix. Given the block matrix form from the previous definition we define the supertrace as follows:
        \begin{gather}
            \text{str}(X) = \text{tr}(A)-\text{tr}(D).
        \end{gather}
    }
    \begin{property}
        As was the case for the ordinary trace, the supertrace is invariant under basis transformations. Furthermore, the cyclicity property also still holds if we modify it slightly as to be compatible with the grading:
        \begin{gather}
            \text{str}(XY) = (-1)^{|X||Y|}\text{str}(YX)
        \end{gather}
        where $|X|, |Y|$ denote the parities of the supermatrices.
    \end{property}

    \newdef{Berezinian}{\label{linalgebra:berezinian}
        The Berezinian or \textbf{superdeterminant} generalizes the determinant of an ordinary matrix. It is (uniquely) defined through the following two conditions:
        \begin{enumerate}
            \item $\text{Ber}(XY) = \text{Ber}(X)\text{Ber}(Y)$
            \item $\text{Ber}(e^X) = e^{\text{str}(X)}$
        \end{enumerate}
        An explicit formula is given by
        \begin{gather}
            \text{Ber}(X) = \det(A - BD^{-1}C)\det(D)^{-1}.
        \end{gather}
        It should be noted that the Berezinian is only well-defined for invertible even matrices.
    }