\chapter{Conformal Field Theory}\index{conformal|seealso{CFT}}

	References for this chapter are \cite{CFT} and the lecture notes by Schellekens.

\section{Conformal invariance}
\subsection{Conformal transformations}
	
	\newdef{Conformal transformation}{
		Consider two (pseudo-)Riemannian manifold $(M, g)$ and $(M', g')$. A smooth map $\varphi:M\rightarrow M'$ if it leaves the metric invariant up to a scale transformation:\footnote{Compare this to definition \ref{diff:conformal_map}.}
		\begin{gather}
			\varphi^*g' = \Omega g
		\end{gather}
		for some smooth positive function $\Omega:M\rightarrow\mathbb{R}^+$.
	}
	
	Infinitesimally these maps are described by a specific type of vector field:
	\newdef{Conformal Killing vector}{\index{Killing!conformal vector}
		Consider a pseudo-Riemannian manifold $(M, g)$. A vector field $X$ is said to be conformal with conformal factor $\kappa:M\rightarrow\mathbb{R}$ if it satisfies
		\begin{equation}
			\mathcal{L}_Xg = \kappa g.
		\end{equation}
		In local coordinates this amounts to
		\begin{equation}
			\nabla_\mu X_\nu + \nabla_\nu X_\mu = \kappa g_{\mu\nu}
		\end{equation}
		where $\nabla$ is the Levi-Civita connection associated to $(M, g)$. Equivalently, a vector field is conformal if its flow is a conformal map.
	}
	
	By parametrizing an infinitesimal transformation as $x^\mu\rightarrow x^\mu+\varepsilon^\mu$ one obtains the following infinitesimal generators:
	\begin{itemize}
		\item Translations: $a^\mu\partial_\mu$
		\item Rotations (orthogonal transformations): $\omega^\mu_{\ \nu} x^\nu\partial_\mu$
		\item Dilations: $\lambda x^\mu\partial_\mu$
		\item Special conformal transformations: $x^2b^\mu\partial_\mu - 2(b\cdot x)x^\mu\partial_\mu$
	\end{itemize}
	As usual, exponentiating these generators gives the finite transformations. The conformal group in signature $(p, q)$ is isomorphic to SO$(p+1, q+1)$. One immediately notices that the Poincar\'e group is a subgroup of the conformal group.
	
	\begin{property}[Stress-energy tensor]
		Now we can consider a theory which is invariant under conformal transformations. The generator of general coordinate transformations is the stress-energy tensor $T$ (current: $\mathcal{J}_\mu = T_{\mu\nu}\varepsilon^\nu$). Requiring conformal invariance then leads to the stress-energy tensor being traceless:
		\begin{gather}
			T^\mu_{\ \mu} = 0
		\end{gather}
	\end{property}

\subsection{In dimension $d=2$}

	In dimension 2 (with Euclidean signature) something special happens. By inserting $d=2$ in the conformal Killing equation we obtain the Cauchy-Riemann equations. The scale factor can thus be written as
	\begin{gather}
		\Omega(x) = \left|\frac{\partial f}{\partial z}\right|^2
	\end{gather}
	for some analytic function $f(z)$. Because of this\footnote{and because we in general look at conformal compactifications (see further below)} we will from here one always work in complex coordinates. Switching to complex coordinates also has important consequences for the metric and stress-energy tensor:
	\begin{align}
		g_{zz} = g_{\zbar\zbar} = 0&\qquad g_{z\zbar} = \frac{1}{2}\\
		\partial_zT_{\zbar\zbar} = \partial_{\zbar} T_{zz} = 0&\qquad T_{z\zbar} = 0
	\end{align}
	The stress-energy tensor thus contains a meromorphic\footnote{The literature often just calls this holomorphic.} component $T_{zz}\equiv T(z)$ and an anti-meromorphic component $T_{\zbar\zbar}\equiv\overline{T}(\zbar)$.

	\newdef{Witt algebra}{\index{Witt algebra}\label{cft:witt_algebra}
		Infinitesimally this gives us an infinite-dimensional algebra. As generators\footnote{These generate the transformation $z\mapsto z-z^{n+1}$ and $\zbar\mapsto\zbar-\zbar^{n+1}$ respectively.} we choose
		\begin{align}
			l_n(z) &= -z^{n+1}\partial_z\\
			\overline{l}_n(\zbar)&=-\zbar^{n+1}\partial_{\zbar}
		\end{align}
		These generate isomorphic Lie algebras with the following commutation relation:
		\begin{gather}
			[l_m, l_n] = (m-n)l_{m+n}.
		\end{gather}
		This Lie algebra is called the \textbf{Witt algebra}.
	}
	
	\newdef{Conformal group}{\index{conformal!group}
		The conformal group of a pseudo-Riemannian manifold $M$ is not just the conformal diffeomorphism group of $M$. The conformal group Conf$(M)$ is defined as the connected component of the identity element in the conformal diffeomorphism group of the conformal compactification of $M$.
	}
	
	\begin{remark}
		Often one finds in the literature that the conformal group of a 2D CFT is infinite-dimensional. However this statement is not entirely true. It is true that the Witt algebra is infinite-dimensional, but one cannot globally exponentiate all generators $l_m$. First of all it should be remarked that the space of all holomorphic functions does not have a group structure, i.e. the composition of holomorphic functions does not have to be holomorphic. The correct notion of a conformal group for 2D Euclidean CFT's is the M\"obius group PSL$(2, \mathbb{C})$. This group is obtained as the Lie group generated by $l_0$ and $l_{\pm1}$, the only generators which can be globally exponentiated.
		
		What is however true is that the conformal group of 2D Minkowski space is infinite-dimensional. It can be shown that Conf$(\mathbb{R}^{1, 1})$ is isomorphic to $\text{Diff}(S^1)_+\times\text{Diff}(S^1)_+$ where the (orientation-preserving) diffeomorphism group Diff$(S^1)_+$ is indeed an infinite-dimensional Lie group (see the intermezzo further below).
		
		At last we should comment that although the conformal group of $\mathbb{R}^{2, 0}$ is finite-dimensional, the infinite-dimensionality of the Witt algebra (and of its extensions) is sufficient for all physical purposes.
	\end{remark}
	
	\newdef{Primary field}{\index{primary!field}\index{conformal!weight}
		A quasi-primary field is defined as any field which transforms tensorially under global conformal transformations;
		\begin{gather}
			\phi'(z', \zbar') = \left(\pderiv{f}{z}\right)^h\left(\pderiv{f}{\zbar}\right)^{\overline{h}}\phi(f(z), \overline{f}(\zbar)).
		\end{gather}
		Fields which satisfy this relation for all conformal transformations are called primary fields. The tuple $(h, \overline{h})$ is called the \textbf{conformal weight} of the field.
	}

\subsection{Intermezzo: Minkowski space}

	As mentioned in the remark above, the conformal group of 2D Minkowski space is infinite-dimensional. The theory of infinite-dimensional manifolds is however a bit more intricate than the theory of finite-dimensional manifolds. A little introduction is therefore in order.

	\newdef{Fr\'echet space}{\index{Fr\'echet!space}
		A locally convex topological vector space which admits a complete translation-invariant metric. Equivalently, a complete Hausdorff topological vector space for which the topology is induced by a countable family of seminorms.
	}
	\newdef{Smooth map}{\index{smooth!function}\index{G\^ateaux}
		The \textit{G\^ateaux derivative} of a continuous map of Fr\'echet spaces $f:F\rightarrow G$ is defined as follows:
		\begin{gather}
			D_hf(a) = \lim_{t\rightarrow0}\frac{f(a+th) - f(a)}{t}.
		\end{gather}
		If this limit exists and is continuous in both $a, h$ then $f$ is said to be of class $C^1$. By iterating this construction one can define $C^k$-maps and even smooth maps of Fr\'echet spaces.
	}
	\newdef{Fr\'echet manifold}{
		A Fr\'echet manifold is a Hausdorff space $M$ together with an atlas of coordinate charts $(U, \varphi)$ such that $\varphi:U\rightarrow F_U$ are homeomorphisms onto a Fr\'echet space and such that the transition functions are smooth maps (of Fr\'echet spaces).
	}
	
	Using the above definitions we can start to analyze the group Diff$(S^1)_+$. First of all we look at the space of all smooth maps $S^1\rightarrow S^1$. This space has the structure of a Fr\'echet manifold modelled on the Fr\'echet space $\mathfrak{X}(S^1)$ of vector fields on the circle\footnote{By this definition we see that $\mathfrak{X}(S^1)$ is isomorphic (as a Lie algebra) to the mapping space $C^\infty(S^1, \mathbb{R})$. Henceforth we identify vector fields $v$ with their corresponding function $\xi$.}:
	\begin{gather}
		\mathfrak{X}(S^1) = \left\{\xi(\theta)\frac{\partial}{\partial\theta}:\theta\in C^\infty(S^1)\right\}.
	\end{gather}
	Let $V_0$ be the set of vector fields which have norm $||v||\leq\pi$ and let $U_0$ be the set of smooth mappings $f\in C^\infty(S^1, S^1)$ such that $f(\theta)\neq-\theta$ for all $\theta\in S^1$. Then there exists a diffeomorphism $\psi:V_0\rightarrow U_0$ which assigns to any vector field $v$ the function $\psi_v:S^1\rightarrow S^1$ such that the arc between $\theta$ and $\psi_v(\theta)$ has length $||v(\theta)||$. If we choose an open subset $U\subset U_0$ of diffeomorphisms then we obtain a chart $(U, \psi^{-1})$ around the identity map. Charts around any diffeomorphism $f:S^1\rightarrow S^1$ are obtained by left multiplication of $U$.
	
	The Lie algebra of Diff$(S^1)_+$ is accordingly given by $\mathfrak{X}(S^1)$ but the induced Lie bracket is the commutator of vector fields with the opposite sign: \[[\cdot, \cdot]_{Lie} = -[\cdot, \cdot]_{\mathfrak{X}(S^1)}.\] Now it is interesting to note that the Witt algebra is actually a subalgebra of $\mathfrak{X}(S^1)$. Consider the maps $\xi_n(\theta):=-ie^{in\theta}$ (the minus sign is a convention). The associated vector fields then satisfy
	\begin{gather}
		\left[\xi_k(\theta)\frac{\partial}{\partial\theta}, \xi_l(\theta)\frac{\partial}{\partial\theta}\right] = -i(l-k)\xi_{k+l}(\theta)\frac{\partial}{\partial\theta}.
	\end{gather}
	These are exactly the relations for the Witt algebra.

\section{Quantization}

	\sremark{From here one we will always work in 2 dimensions.}
	
\subsection{Radial quantization}

	The charge of a conserved current $\mathcal{J}^\mu$ is generally given by the following formula\footnote{See Noether's theorem \ref{qft:noethers_theorem}.}:
	\begin{gather}
		Q = \int_\Sigma\mathcal{J}^0(x, t)
	\end{gather}
	where $\Sigma$ is a spacelike hypersurface. Often we will compactify the spatial dimension (this can be seen as some kind of regularization) and Wick rotate the time dimension. By a conformal transformation one can then go back to the plane (mapping one end of the cylinder to the origin and the other side to a circle at $+\infty$). After these transformations we obtain the following form for an operator on the plane:
	\begin{gather}
		Q_\varepsilon = \frac{1}{2\pi i}\oint dz\varepsilon(z)T(z) + \frac{1}{2\pi i}\oint d\zbar\varepsilon(\zbar)\overline{T}(\zbar).
	\end{gather}
	As usual we expect an infinitesimal transformation of a field $\phi$ to be given by the commutator $[Q_\varepsilon, \phi(w, \overline{w})]$ or in integral form by\footnote{We implicitly assumed a holomorphic split such that we can ignore (anti-holomorphic) $\overline{T}(\zbar)$ contributions.} \[\frac{1}{2\pi i}\oint dz\varepsilon(z)[T(z)\phi(w, \overline{w}) - \phi(w, \overline{w})T(z)].\] However because all the objects in this formula are operators we should take into account the operator ordering. On the plane this is given by the so-called \textbf{radial ordering}:\index{radial ordering}
	\begin{gather}
		\mathcal{R}(A(z, \zbar)B(w, \overline{w})) = \begin{cases}
			A(z, \zbar)B(w, \overline{w})&|z|>|w|\\
			B(w, \overline{w})A(z, \zbar)&|w|>|z|
		\end{cases}
	\end{gather}
	After deforming our integration contour we obtain the following general formula:
	\begin{gather}
		[Q_\varepsilon, \phi(w, \overline{w})] = \frac{1}{2\pi i}\oint dz\varepsilon(z)\mathcal{R}(T(z)\phi(w, \overline{w}))
	\end{gather}
	where the contour is now a circle around the point $w$. For primary fields with conformal weight $h$ one can also write an infinitesimal transformation as \[\delta_\varepsilon\phi(z, \zbar) = h(\partial_z\varepsilon(z))\phi(z, \zbar) + \partial_z\phi(z, \zbar).\] Comparing these two expressions leads us to the following form of the operator product:
	\begin{gather}
		\mathcal{R}(T(z)\phi(w, \overline{w})) = \frac{h}{(z-w)^2}\phi(w, \overline{w}) + \frac{1}{z-w}\partial_w\phi(w, \overline{w}) + \text{higher order in } (z-w)
	\end{gather}

\subsection{Virasoro algebra}\index{Virasoro algebra}

	When going from classical systems to quantum systems we have to replace symmetry operations by (unitary) projective representations. As in the case of ordinary groups, the projective representations of Lie groups are related to central extensions of Lie groups. As an example of the Lie group-Lie algebra correspondence it can be shown that central extensions of Lie algebras (see section \ref{section:central_extension_algebra}) are in correspondence with central extensions of Lie groups.
	
	If we now want a quantum theory equipped with an action of the ''conformal group'', then we need to construct a central extension of the Witt algebra. By applying the construction of section \ref{section:central_extension_algebra} we obtain the \textbf{Virasoro algebra} as a (universal) central extension of the Witt algebra by\footnote{The reason why we extend by $\mathbb{C}$ instead of $\mathbb{R}$ is that we are working with complexified Lie algebras.} $\mathbb{C}$ associated to the cocycle \[\Theta:(L_m, L_n)\mapsto \frac{c}{12}m(m^2-1)\delta_{m+n, 0}\]

	To obtain the Virasoro algebra from a more physical point of view we look at the stress-energy tensor. If we choose the current $\mathcal{J}^n(z) := z^{n+1}T(z)$ we expect to obtain the transformation $z\longrightarrow z-z^{n+1}$. In analogy with the generators of the Witt algebra (definition \ref{cft:witt_algebra}) we call this generator $L_n$:
	\begin{gather}
		L_n = \frac{1}{2\pi i}\oint dzz^{n+1}T(z).
	\end{gather}
	 This relation can be inverted using the residue theorem to obtain the following expression:
	\begin{gather}
		T(z) = \sum_{n\in\mathbb{Z}}z^{-n-2}L_n.
	\end{gather}
	Using the above expression and the product operator expansion of $T(z)T(w)$ one obtains exactly the commutation relation of the Virasoro algebra:
	\begin{gather}
		[L_m, L_n] = (m-n)L_{m+n} + \frac{c}{12}m(m^2-1)\delta_{m+n, 0}
	\end{gather}
	The occurence of the central charge $c$ gives a conformal anomaly on quantization (see the definition of a the vacuum further below). However the central charge does not affect the $\mathfrak{sl}(2, \mathbb{C})$ subalgebra spanned by $L_{-1}, L_0$ and $L_1$ which implies that concepts such as the conformal weight are still useful.

\subsection{Representation theory}

	\newdef{Highest weight state}{
		A state with minimal eigenvalue for $L_0$. Equivalently, a state that is annihilated by all generators $L_n$ for $n\geq1$:
		\begin{gather}
			L_n|h\rangle = 0
		\end{gather}
	}

	\newdef{Vacuum}{\index{vacuum}
		Consider the Virasoro generators $\{L_n\}_{n\in\mathbb{Z}}$. The vacuum $|0\rangle$ is defined as the maximally symmetric state. In terms of generators this means that we want to have \[L_n|0\rangle = 0\] for as many $n$ as possible. However due to the Virasoro commutation relations this is not possible for all $n\in\mathbb{Z}$. Instead one can only require the vanishing for $n\geq-1$.
	}
	
	\newdef{Descendants}{
		By acting with the generators $L_n$, $n\leq-1$ on a highest weight state $|h\rangle$ one obtains a whole family of states. These are called the descendants of $|h\rangle$ and together they span the Verma module\footnote{See definition \ref{lie:verma_module}.} (associated to $|h\rangle$).
	}
	
\section{Partition functions}