\chapter{Tensor Networks}\label{chapter:tensor_networks}

    The main references for the sections on \textit{topological order} are~\citet{sahinoglu_characterizing_2014,williamson_matrix_2016}.

    \minitoc

\section{Matrix Product States}
\subsection{Finite-dimensional lattices}

    \newdef{Matrix product state}{\index{matrix product!state}
        Let $\mathcal{H}_n$ be the local Hilbert spaces of dimension $d_n\in\mathbb{N}$, where $n\in\{1,\ldots,N\}$. A state $|\psi\rangle$ in the total Hilbert space $\bigotimes_i\mathcal{H}_i$ is called a matrix-product state (MPS) with periodic boundary conditions if there exist matrices $A^{i_n}(n)\in\mathcal{L}(\mathbb{C}^{D_n},\mathbb{C}^{D_{n-1}})$ with $i_n\leq d_n$ such that
        \begin{gather}
            |\psi\rangle = \sum_{\{i_k\}}\tr\left(\prod_n^NA^{i_n}(n)\right)\,|i_1\ldots i_N\rangle\,.
        \end{gather}
        For each lattice site $n\in\mathbb{Z}$, the set of matrices $\{A^{i_n}_{\alpha\beta}(n)\}$ can be regarded as forming a rank-3 tensor. The periodic boundary condition requires that $D_0=D_N$, otherwise the trace would be ill-defined. Different boundary conditions can be implemented by inserting an additional factor $X$ at the end of the trace.
    }

    \begin{remark}[Physical and virtual spaces]\index{ancilla}
        For each `physical' index $i_n$, one can regard the matrix $A^{i_n}(n)$ as a linear map between `virtual' (or \textbf{ancilla}) spaces $\mathbb{C}^{D_n}$.
    \end{remark}

    \newformula{MPS projector}{
        Consider an MPS defined by tensors $\{A(n)\}_{n\leq N}$. The associated MPS projector is defined as follows:
        \begin{gather}
            \mathcal{P}_A := \sum_{i,\alpha,\beta}A^i_{\alpha\beta}(n)|i\rangle\langle\alpha\beta|\,.
        \end{gather}
    }

    \newformula{Transfer operator}{\index{transfer operator}\label{tennet:transfer_operator}
        Given MPS tensors $\{A(n)\}_{n\leq N}$, one can define a transfer operator:
        \begin{gather}
            \mathbb{E}(n) := \sum_{i=1}^{d_n}A^i(n)\otimes\overline{A^i}(n)\,.
        \end{gather}
    }
    \begin{formula}[Superoperator]\index{super-!operator}
        More generally, one can define for every local observable $\widehat{O}_n$ a superoperator in $\mathcal{L}(\mathbb{C}^{D_n}\otimes\overline{\mathbb{C}^{D_n}},\mathbb{C}^{D_{n-1}}\otimes\overline{\mathbb{C}^{D_{n-1}}})$:
        \begin{gather}
            \mathbb{E}_{O_n}(n) := \sum_{i,i'=1}^{d_n}\langle i\mid \widehat{O}_n\mid i' \rangle A^{i'}(n)\otimes\overline{A^i}(n)\,.
        \end{gather}
        Comparing with the definition of the transfer operator, it can be seen that $\mathbb{E}$ is given by the superoperator associated to the unit operator. Given two sets of MPS tensors $\{A(n)\},\{B(n)\}$, one can define a generalized superoperator as follows:
        \begin{gather}
            \mathbb{E}^A_B(n) := \sum_{i=1}^{d_n}A^i(n)\otimes\overline{B^i}(n)\,.
        \end{gather}
    \end{formula}
    \begin{example}
        Using these definitions, one can rewrite the formula for expectation values more efficiently. Given a product operator $\widehat{O}=\bigotimes_i^N\widehat{O}_i$, one finds that
        \begin{gather}
            \langle\psi[A]\mid\widehat{O}\mid\psi[A]\rangle = \tr\left(\mathbb{E}_{O_1}(1)\cdots\mathbb{E}_{O_N}(N)\right)\,.
        \end{gather}
    \end{example}

    \begin{formula}
        Associated to the superoperator $\mathbb{E}_O(n)$, one can define a virtual superoperator:
        \begin{gather}
            \begin{aligned}
                \mathcal{E}^{(n)}_{O_n}(\phi) &= \sum_{i,i'=1}^{d_n}\langle s\mid \widehat{O}_n\mid s' \rangle A^{i'}(n)\phi A^i(n)^\dag\,,\\
                \widetilde{\mathcal{E}}^{(n)}_{O_n}(\phi) &= \sum_{i, i'=1}^{d_n}\langle s\mid \widehat{O}_n\mid s' \rangle A^i(n)^\dag\sigma A^{i'}(n)\,,
            \end{aligned}
        \end{gather}
        where $\phi\in\mathcal{L}(\mathbb{C}^{D_n})$ and $\sigma\in\mathcal{L}(\mathbb{C}^{D_{n-1}})$.
    \end{formula}
    \begin{property}
        The map $\mathcal{E}_{\mathbbm{1}}^{(n)}$ associated to the transfer operator is a CP map (\cref{operators:cp_map}) and the associated Kraus operators (\cref{qc:kraus}) are the MPS matrices $A^i(n)$.
    \end{property}

\subsection{Injectivity}\index{injective!MPS}

    \newdef{Injective MPS}{
        A translation-invariant MPS is said to be injective if its transfer operator has a unique maximal eigenvalue.
    }

    In the next sections, the MPSs will always be assumed to be injective unless stated otherwise.

\subsection{Gauge freedom and canonical forms}

    \begin{property}[Gauge freedom]
        As is clear from the construction of matrix product states, there exists some freedom in the representation of the MPS tensors. One can always perform a transformation of the form $A(n)\rightarrow X^{-1}(n)A(n)X(n+1)$.
    \end{property}
    \begin{remark}
        If one uses periodic boundary conditions, one must require that $X(L+1)=X(1)$, where $L$ is the lattice size.
    \end{remark}

    Using the gauge freedom in the representation of a generic MPS, one can construct certain forms which have useful properties.
    \begin{construct}[Left canonical form\footnotemark]
        \footnotetext{Also called the \textbf{left isometric form}, \textbf{left orthogonal form} or just \textbf{left gauge}.}
        This form is specified by the following property:
        \begin{gather}
            \begin{tikzpicture}[baseline]
                \node[rectangle,draw=black,minimum size=20] (A) at (0,1) {$A_L$};
                \node[rectangle,draw=black,minimum size=20] (Ad) at (0,-1) {$\overline{A_L}$};
                \draw (A) -- (Ad);
                \draw (A) to[out=200,in=160] (Ad);
                \draw (A) -- +(1,0);
                \draw (Ad) -- +(1,0);
                \node (E) at (1.5,0) {$=$};
                \node (A2) at (3,1) {};
                \node (Ad2) at (3,-1) {};
                \draw (A2) to[out=200,in=160] (Ad2);
            \end{tikzpicture}
        \end{gather}
        Any MPS can be written in this form. First, construct the transfer operator $\mathbb{E}(n)$ for every site and find its maximal eigenvector. By the Perron--Frobenius theorem~\ref{linalgebra:perron_frobenius}, this eigenvector (which is in fact itself a matrix) is positive and, hence, allows a decomposition of the form $\lambda(n)=L^\dag(n)L(n)$. The left orthogonal forms are then defined by
        \begin{gather}
            A_L(n) := L(n)A(n)L^{-1}(n+1)\,.
        \end{gather}
        In a similar manner, one can construct the right orthogonal form $A_R$.
    \end{construct}

    \begin{method}[Vidal]\index{Vidal}
        Given a general quantum state in terms of an $n$-leg tensor, there exists an efficient way of constructing the left (or right) canonical forms due to~\citet{vidal_efficient_2003}. To this end, perform a cut between the first and second site and apply a singular value decomposition to obtain a tensor of the form $U^{[1]}SV^{[2,\ldots]}$. Recursively applying this procedure to the product of the singular values $S$ and the right unitary $V$ will result in the canonical form.
    \end{method}

    \begin{construct}[Mixed canonical form]
        The left and right canonical forms can also be combined. Let $L(n)$ and $R(n)$ be the decompositions of the left and right eigenvectors of the transfer operator at site $n$, i.e.~$\lambda(n)=L^\dag(n)L(n)$ and $\rho(n)=R(n)R^\dag(n)$. The left and right canonical forms are then related by a matrix $C(n)$ in the following way: $A_L(n)C(n+1)=C(n)A_R(n)$. These matrices are given by
        \begin{gather}
            C(n)=L(n)R(n)\,.
        \end{gather}
    \end{construct}

\subsection{Translation-invariant states}

    \newdef{Uniform MPS}{\index{uniform!MPS}
        By choosing all MPS tensors $A(n)=B$ for a given tensor $B$, one obtains a translation-invariant (TI) state. These MPSs form the variational class of uniform MPS.
    }
    \begin{remark}[TIMPS]
        Not every TIMPS is uniform, there should only exist a local gauge transformation $A'(n) = U(n-1)A(n)U(n)^{-1}$ such that $A'(n)$ is uniform (in certain cases, this is only possible by enlarging the bond dimension).
    \end{remark}

\section{Matrix product operators}

    \newdef{Matrix product operator\footnotemark}{\index{matrix product!operator}
        \footnotetext{As in the case of matrix product states this will be abbreviated as \textbf{MPO}.}
        
        Starting from the general form of an MPS, one can easily construct more general objects. By replacing the rank-3 tensors $A^i(n)$ with rank-4 tensors $A^{i,j}(n)$ and $|i_1\cdots i_n\rangle$ by $|i_1\rangle\langle j_1|\otimes\cdots\otimes|i_n\rangle\langle j_n|$, one obtains the notion of a matrix-product operator:
        \begin{gather}
            \widehat{O} = \sum_{\{i_k,j_l\}}\tr\left(\prod_{m,n=1}^NO^{i_m,j_n}(n)\right)|i_1\cdots i_N\rangle\langle j_1\cdots j_n|\,.
        \end{gather}
        In terms of a basis $\{\widehat{O}_i\}$ for the space of local operators, this becomes:
        \begin{gather}
            \widehat{O} = \sum_{\{i_k\}}\tr\left(\prod_n^NA^{i_n}(n)\right)\widehat{O}_{i_n}\,.
        \end{gather}
    }

    \begin{method}[Local Hamiltonian to MPO]
        Given a local Hamiltonian $\widehat{H}=\sum_i\widehat{H}^{(i)}$, one can build an MPO that generates this Hamiltonian\footnote{In fact, one can use this procedure to turn any local operator into an MPO.}:
        \begin{gather}
            \widehat{H} := \sum_{\{i_k,j_l\}}\tr\left(\prod_{m,n=1}^NA^{i_m,j_n}(n)\right)|i_1\cdots i_N\rangle\langle j_1\cdots j_n|\,.
        \end{gather}
        To obtain this MPO, one can use the concept of a \textit{cellular automaton}. This is a set of possible states together with a set of rules that say how one can go from one state to another. To obtain the set of states in this case, look at a given site $i$. All distinct combinations of 1-site operators to the right of $i$ give rise to a distinct state $\mu$. The transition rules are obtained by looking at which operator can be placed at the site $i$ in a way consistent with the form of the given Hamiltonian.
    \end{method}
    \begin{example}
        Consider a 2-site Hamiltonian of the form
        \begin{gather}
            \widehat{A}\otimes\widehat{B}\otimes\mathbbm{1}\otimes\cdots+\mathbbm{1}\otimes\widehat{A}\otimes\widehat{B}\otimes\mathbbm{1}\otimes\cdots+\cdots\,.
        \end{gather}
        For a specific site $i$, three distinct possibilities can arise:
        \begin{itemize}
            \item There are only identity operators acting to the right of $i$.
            \item Immediately to the right there is an operator $\widehat{B}$ acting on $i+1$.
            \item Somewhere to the right there is a combination $\widehat{A}\otimes\widehat{B}$.
        \end{itemize}
        The transition rules for this automaton are then given by the following list:
        \begin{itemize}
            \item $1\rightarrow2$: $\mathbbm{1}$,
            \item $1\rightarrow2$: $\widehat{B}$,
            \item $2\rightarrow3$: $\widehat{A}$, and
            \item $3\rightarrow3$: $\mathbbm{1}$.
        \end{itemize}
        What is useful here is that this set of transition rules can be turned into a matrix:
        \begin{gather}
            T=\begin{pmatrix}\mathbbm{1}&\widehat{B}&0\\0&0&\widehat{A}\\0&0&\mathbbm{1}\end{pmatrix}\,.
        \end{gather}
        The MPO is then obtained by setting the MPO matrix $A$ equal to $T$ at every site.
    \end{example}

\subsection{MPO-injectivity}

    \newdef{MPO-injective PEPS}{\index{MPO!injectivity}
        Consider a trivalent PEPS network on a manifold $M$ and select a simply-connected subregion $\Omega\subset\Lambda$. By contracting the tensors withing this region, one obtains a linear map
        \begin{gather}
            A_\Lambda:(\mathbb{C}^D)^{\otimes|\Lambda|}\rightarrow(\mathbb{C}^d)^{\otimes|\partial\Lambda|}
        \end{gather}
        from the virtual spaces on the edges to the physical space in the bulk. This PEPS is said to be MPO-injective if there exists a linear map (four-leg tensor)
        \begin{gather}
            M:\mathbb{C}^D\otimes\mathbb{C}^m\rightarrow \mathbb{C}^D\otimes\mathbb{C}^m
        \end{gather}
        such that, for every subregion $\Omega\subset\Lambda$, the linear map $A_\Lambda$ is injective on a (maximal) subspace $S$ for which the projector onto $S$ can be written as an MPO constructed from the tensors $M$ living on the boundary $\partial\Lambda$. See \cref{fig:mpo_injectivity}, where the tensors $M$ are given by crossings of black and red lines.
    }

    \begin{figure}[ht!]
        \centering
        \begin{tikzpicture}
            \node[circle,draw=black,fill=black,minimum size=5pt,label={110:$A$}] (T) at (0,-1) {};
            \node[circle,draw=black,fill=black,minimum size=5pt,label={110:$A^\dag$}] (T2) at (1.85,1.2) {};
            \node (E) at (4,0) {$=$};
            \node[label={120:$M$}] (M1) at (7, 1.5) {};
            \node[label={-90:$M$}] (M2) at (8.2, -0.92) {};
            \node[label={-90:$M$}] (M3) at (5.8, -0.92) {};
            \draw[line width=0.5 mm] (T) -- (T2);
            \draw[<-, line width=0.5 mm] (T) -- +(0,1.2);
            \draw[<-, line width=0.5 mm] (T) -- (1, -1.6);
            \draw[<-, line width=0.5 mm] (T) -- (-1, -1.6);
            \draw[->, line width=0.5 mm] (T2) -- +(0,1.2);
            \draw[->, line width=0.5 mm] (T2) -- (2.85, 0.6);
            \draw[->, line width=0.5 mm] (T2) -- (0.85, 0.6);
            \draw[draw=red, line width=0.5 mm] (7, 0) circle (1.5);
            \draw[<-<, line width=0.5 mm] (7, 0.75) -- (7, 2.25);
            \draw[<-<, line width=0.5 mm] (7.5, -0.4) -- (9, -1.15);
            \draw[<-<, line width=0.5 mm] (6.5, -0.4) -- (5, -1.15);
        \end{tikzpicture}
        \caption{MPO-injective PEPS.}
        \label{fig:mpo_injectivity}
    \end{figure}

    \begin{axiom}[Pulling-through]\index{pulling-through condition}
        One of the key features of topological order is that it cannot be detected locally. Only a global measurement can show the existence of topologically ordered states. To this end, an axiom is introduced that allows to pull an MPO through the lattice. Graphically this is shown in \cref{fig:pulling_through}.
    \end{axiom}

    \begin{figure}[ht!]
        \centering
        \begin{tikzpicture}
            \node[circle,draw=black,fill=black,minimum size=5pt,label={110:$A$}] (T) at (0,0) {};
            \node[circle,draw=black,fill=black,minimum size=5pt,label={110:$A$}] (T2) at (4,0) {};
            \node (E) at (2,0) {$=$};
            \draw[line width=0.5 mm] (T) -- +(0.5,0.5);
            \draw[line width=0.5 mm] (T) -- +(0,1.2);
            \draw[line width=0.5 mm] (T) -- (1, -0.6);
            \draw[line width=0.5 mm] (T) -- (-1, -0.6);
            \draw[line width=0.5 mm, draw=red] (0.34, 0.95) arc (70:250:1);
            \draw[line width=0.5 mm] (T2) -- +(0.5,0.5);
            \draw[line width=0.5 mm] (T2) -- +(0,1.2);
            \draw[line width=0.5 mm] (T2) -- (5, -0.6);
            \draw[line width=0.5 mm] (T2) -- (3, -0.6);
            \draw[line width=0.5 mm, draw=red] (4.85, 0.5) arc (30:-90:1);
        \end{tikzpicture}
        \caption{Pulling-through condition.}
        \label{fig:pulling_through}
    \end{figure}

\subsection{MPO-symmetries for SPT phases}

    One can generalize the above framework to also include symmetry-protected topological order. To this end, one has to slightly modify the axioms.
    \begin{axiom}[Pulling-through for SPT phases]\index{pulling-through condition}
        When pulling a symmetry-MPO through a tensor one has to act with a unitary on the physical level. Graphically this is shown in \cref{fig:pulling_through_spt}.
    \end{axiom}

    \begin{figure}[ht!]
        \centering
        \begin{tikzpicture}
            \node[circle,draw=black,fill=black,minimum size=5pt,label={110:$A$}] (T) at (0,0) {};
            \node[circle,draw=black,fill=black,minimum size=5pt,label={110:$A$}] (T2) at (4,0) {};
            \node (E) at (2,0) {$=$};
            \draw[line width=0.5 mm] (T) -- +(0.5,0.5);
            \draw[line width=0.5 mm] (T) -- +(0,1.2);
            \draw[line width=0.5 mm] (T) -- (1, -0.6);
            \draw[line width=0.5 mm] (T) -- (-1, -0.6);
            \draw[line width=0.5 mm, draw=red] (0.34, 0.95) arc (70:250:1);
            \draw[line width=0.5 mm] (T2) -- +(0.5,0.5);
            \draw[line width=0.5 mm] (T2) -- +(0,1.2);
            \draw[line width=0.5 mm] (T2) -- (5, -0.6);
            \draw[line width=0.5 mm] (T2) -- (3, -0.6);
            \draw[line width=0.5 mm, draw=red] (4.85, 0.5) arc (30:-90:1);
            \node[rectangle, draw=black, fill=blue, minimum size=2pt, label={$U(g)$}] (U) at (4, 1.2) {};
        \end{tikzpicture}
        \caption{Pulling-through condition for SPT phases.}
        \label{fig:pulling_through_spt}
    \end{figure}