\chapter{Wave and Matrix Mechanics}

    The main reference for this chapter is \cite{bransden}. In this chapter the two basic formalisms of Quantum Mechanics are introduced: wave and matrix mechanics.

\section{Schr\"odinger picture}

    \begin{formula}[Time-independent Schr\"odinger equation]\index{Schr\"odinger!equation}\index{Hamilton!function}\label{wavematrix:TISE}
        \begin{gather}
            \hat{H}\psi(x) = E\psi(x).
        \end{gather}
        The operator $\hat{H}$ is called the \textbf{Hamiltonian} of the system and $\psi$ is an element of the vector space $L^2(\mathbb{R},\mathbb{C})\otimes\mathcal{H}$ with $\mathcal{H}$ the internal Hilbert space (describing for example the spin or charge of a particle). This is an eigenvalue equation for the energy levels of the system.
    \end{formula}

    \newprop{Orthogonality}{
        Let $\{\psi_i\}_{i\in I}\subset L^2(\mathbb{R})$ be a collection of eigenfunctions of the TISE (the internal space is hidden for convenience). These functions can be normalized with respect to the inner product \eqref{lebesgue:L2_inner_product} such that they obey the following relation:
        \begin{gather}
            \label{wavematrix:orthogonality}
            \int_\mathbb{R}\overline{\psi_i}(x)\psi_j(x)dx = \delta_{ij}.
        \end{gather}
    }

    The time evolution of a wave function is defined through the following equation:
    \newformula{TDSE}{\index{Schr\"odinger!equation}\label{wavematrix:TDSE}
        The following partial differential equation is called the \textbf{(time-dependent) Schr\"odinger equation}:
        \begin{gather}
            i\hbar\pderiv{}{t}\psi(x,t) = \hat{H}\psi(x,t).
        \end{gather}
        In case $\hat{H}$ is time-independent, the TISE can be obtained from this equation by separation of variables (see Derivation \ref{section:tdse_tise}).
    }
    \begin{example}[Massive particle in a stationary potential]
        \begin{gather}
            \label{wavematrix:TDSE_position}
            i\hbar\pderiv{}{t}\psi(x,t) = \left(\frac{\hat{p}^2}{2m} + \hat{V}(x)\right)\psi(x,t)
        \end{gather}
        \begin{proof}[Derivation of TISE from TDSE]
            \begin{mdframed}[roundcorner=10pt, linecolor=blue, linewidth=1pt]
                Starting from the one-dimensional TDSE in position space with a time-independent potential
                \begin{gather}
               		i\hbar\pderiv{}{t}\Psi(x,t) = \left(-\frac{\hbar^2}{2m}\partial_x^{\ 2} + V(x)\right)\Psi(x,t),
               	\end{gather}
                one can perform a separation of variables and assert a solution of the form $\Psi(x,t) = \psi(x)T(t)$. Inserting this in the previous equation gives
                \begin{gather}
               		i\hbar\psi(x)T'(t) = -\frac{\hbar^2}{2m}T(t)\psi''(x) + V(x)\psi(x)T(t).
               	\end{gather}
                Dividing both sides by $X(x)T(t)$ and rearranging the terms gives
                \begin{gather}
               		i\hbar\frac{T'(t)}{T(t)} = \left(-\frac{\hbar^2}{2m} + V(x)\right)\frac{\psi''(x)}{\psi(x)}.
               	\end{gather}
                Because the left side only depends on $t$ and the right side only depends on $x$, one can conclude that they are both equal to a constant $E$. This leads to the following system of differential equations:
                \begin{gather}
                    \begin{cases}
                        i\hbar T'(t) - ET(t) = 0,\\
                        \left(-\frac{\hbar^2}{2m} + V(x)\right)\psi''(x) = E\psi(x).
                    \end{cases}
                \end{gather}
                The first equation immediately gives a solution for $T$:
                \begin{gather}
                    \label{derivations_qm:exponential}
               		T(t) = Ce^{-\frac{i}{\hbar}Et}.
               	\end{gather}
                Rearranging the second solution in the system gives the TISE:
                \begin{gather}
                    \label{derivations_qm:TISE}
               		\psi''(x) = -\frac{2m}{\hbar^2}\left(E - V(x)\right)\psi(x).
               	\end{gather}\qed
            \end{mdframed}
        \end{proof}
    \end{example}

    \begin{formula}[General solution]
        A general solution of the time-dependent Schr\"odinger equation (for time-independent Hamiltonians) is given by the following formula (cf. Formula \ref{ode:first_order_general_solution}):
        \begin{gather}
            \label{wavematrix:general_solution}
            \psi(x,t) = \sum_Ec_E\psi_E(x)e^{-\frac{i}{\hbar}Et},
        \end{gather}
        where the functions $\psi_E(x)$ are the eigenfunctions of the TISE \ref{wavematrix:TISE}. The coefficients $c_E$ can be found using the orthogonality relations
        \begin{gather}
            \label{wavematrix:general_solution_coefficients}
            c_E=\left(\int_\mathbb{R}\overline{\psi_E}(x)\psi(x,t_0)dx\right)e^{\frac{i}{\hbar}Et_0}.
        \end{gather}
    \end{formula}

    ?? COMPLETE ??

\subsection{Hydrogen atom}

    Consider the hydrigen atom, i.e. a single proton (the nucleus) orbited by a single electron with only the electrostatic Coulomb force acting between them (gravity can safely be neglected):
    \begin{gather}
        \hat{H} := \frac{\hat{p}_p^2}{2m_p} + \frac{\hat{p}_e^2}{2m_e} - \frac{e^2}{4\pi\varepsilon r^2}.
    \end{gather}
    It is not hard to see that this is the quantum mechanical version of the Kepler problem (Section \ref{classic:kepler}). The special property of the Kepler problem was that it contained a ``hidden'' symmetry that gave rise to the conserved Laplace-Runge-Lenz vector \ref{classic:lrl_vector}. As is the case for all conserved charges in quantum mechanics, this symmetry induces a degeneracy of the energy eigenvalues. Degeneracy of the magnetic quantum number $m$ follows from rotational symmetry, but the energy levels of the hydrogen atom only depend on the principal quantum number $n$. The degeneracy of the total angular quantum number $l$ is that due to the ``hidden'' $\mathrm{SO}(4)$-symmetry. It is often called an ``accidental degeneracy'' for this reason.

    ?? COMPLETE ??

\subsection{Molecular dynamics}

    Consider the Hamiltonian of two interacting atoms:
    \begin{gather}
        \hat{H} := \frac{\hat{P}_1^2}{2M_1} + \frac{\hat{P}_2^2}{2M_2} + \frac{q_1q_2}{4\pi\varepsilon R^2} + \sum_i\frac{\hat{p}_i^2}{2m} - \frac{eq_1}{4\pi\varepsilon r_{i1}^2} - \frac{eq_2}{4\pi\varepsilon r_{i2}^2} + \sum_{i\neq j}\frac{e^2}{4\pi\varepsilon r_{ij}^2},
    \end{gather}
    where the indices $i,j$ indicate the electrons and capital symbols denote operators associated to the nuclei.

    Except for the most simple situations, solving the Schr\"odinger equation for this Hamiltonian becomes intractable (both analytically and numerically). However, in general one can approximate the situation. The mass of nuclei are much larger than those of the electrons and this influences their motion, they move much slower than the electrons. In essence the nuclei and electrons live on different time scales and this allows to decouple their dynamics:
    \begin{gather}
        \hat{H}_\mathrm{nucl} = \frac{\hat{P}_1^2}{2M_1} + \frac{\hat{P}_2^2}{2M_2} + \frac{Q_1Q_2}{4\pi\varepsilon R^2} + V_\mathrm{eff}(R_1,R_2).
    \end{gather}
    The electrons generate an effective potential for the nuclei and the Schr\"odinger equation decouples as follows:
    \begin{align}
        \hat{H}_\mathrm{nucl}(R)\psi(R) &= E\psi(R)\\
        \hat{H}_\mathrm{el}(r,R)\phi(r,R) &= E_\mathrm{el}\phi(r,R).
    \end{align}
    This is the so-called \textbf{Born-Oppenheimer approximation}. From a more modern physical perspective this approximation can also be seen to be a specific instance of renormalization theory, where the short time-scale (or, equivalently, the high energy-scale) degrees of freedom are integrated out of the theory.

\section{Heisenberg picture}

    This is the right place to reflect on what the wave function is and how it relates to the state of a system. At every point it gives the probability of observing a particle (or whatever object is being studied). But what if one wants to express this information in terms of momenta instead of positions? The information about the state should not depend on the chosen ``representation''. To this end, a state vector $|\psi\rangle$ that represents the state of the system as an abstract vector in some Hilbert space is introduced.
    \begin{notation}[Dirac notation]\index{bra-ket notation}\index{Dirac!notation}
        This notation is often called the \textbf{braket notation}. State vectors $|\psi\rangle$ are called \textbf{ket}'s and their duals $\langle\psi|$ are called \textbf{bra}'s. The inner product of a state $|\phi\rangle$ and a state $|\psi\rangle$ is denoted by $\langle\phi|\psi\rangle$.
    \end{notation}
    But then, how does one recover the position (configuration) representation $\psi(x)$? This is simply the projection of the state vector $|\psi\rangle$ on the ``basis function'' $\delta(x)$, i.e. $\psi(x)$ represents an expansion coefficient in terms of a ``basis'' for the physical Hilbert space. In the same way one can obtain the momentum representation $\psi(p)$ by projecting on the plane waves $e^{ipx}$.

    \begin{remark}
        It should be noted that neither the ``basis states'' $\delta(x)$, nor the plane waves $e^{ipx}$ are square-integrable and, hence, they are not elements of the Hilbert space $L^2(\mathbb{R},\mathbb{C})$. In the next chapter this issue will be resolved through the concept of \textit{rigged Hilbert spaces}.
    \end{remark}

    \begin{formula}[Matrix representation]
        The following formula gives the matrix representation of an operator $\hat{A}$ with respect to the orthonormal basis $\{|\psi_i\rangle\}_{i\leq n}$ cf. Construction \ref{linalgebra:matrix_representation}:
        \begin{gather}
            \label{wavematrix:matrix_entry}
            A_{ij} := \langle\psi_i|\hat{A}|\psi_j\rangle.
        \end{gather}
    \end{formula}
    \begin{remark}
         The basis $\big\{|\psi_i\rangle\big\}_{i\leq n}$ need not consist out of eigenfunctions of $\hat{A}$.
    \end{remark}

    ?? COMPLETE ??

\section{Uncertainty principle}

    \newdef{Compatible observables}{\index{observable!compatibility}
        Two observables are said to be compatible if they share a complete set of eigenvectors.
    }

    \newdef{Expectation value}{
        The expectation value of an operator $\hat{A}$ in a state $|\psi\rangle$ is defined as
        \begin{gather}
            \langle\hat{A}\rangle_\psi := \langle\psi|\hat{A}|\psi\rangle.
        \end{gather}
        The subscript $\psi$ is often left implicit. As in ordinary statistics \eqref{statistics:variance_without_sum}, the uncertainty or variance is defined as follows:
        \begin{gather}
            \Delta A := \langle\hat{A}^2\rangle - \langle\hat{A}\rangle^2.
        \end{gather}
    }

    \newformula{Uncertainty relation}{\index{Heisenberg!uncertainty relation}\index{Robertson|see{Heisenberg uncertainty relation}}
        Let $\hat{A},\hat{B}$ be two operator and let $\Delta A,\Delta B$ be the corresponding uncertainties. The (\textbf{Robertson}) uncertainty relation reads as follows:
        \begin{gather}
            \label{wavematrix:uncertainty_relation}
            \Delta A\Delta B\geq\frac{1}{4}\left|\left\langle\left[\hat{A},\hat{B}\right]\right\rangle\right|^2.
        \end{gather}
    }