\chapter{\difficult{Axiomatic QFT}}

    For the sections on the Haag-Kastler framework and its extensions hte reader is referred to the work of \textit{Brunetti, Fredenhagen et al}. A reference for the remaining sections on algebraic QFT is \cite{baez_aqft}. The sections on cohomological quantization and quantization from a sheaf point of view are based on \cite{nuiten}. For the sections on TQFTs and \textit{open-closed} TQFTS the reader is referred to the original papers \cite{atiyah} and \cite{open_closed}, respectively.

\section{Algebraic QFT}
\subsection{Haag-Kastler axioms}

    \begin{axiom}[Local net of observables]\index{microcausality|see{locality}}\index{Einstein!causality}\index{locality}\label{aqft:microcausality}
        To every causally closed set \ref{relativity:causal_closure} one associates a $C^*$-algebra. This assignment is required to satisfy the following conditions:
        \begin{enumerate}
            \item\textbf{Isotony}: If $O_1\subset O_2$, then $\mathcal{A}(O_1)\hookrightarrow\mathcal{A}(O_2)$.
            \item\textbf{(Causal) locality}\footnote{Also called \textbf{microcausality} or \textbf{Einstein causality}.}: If $O_1$ and $O_2$ are spacelike separated, then $[\mathcal{A}(O_1),\mathcal{A}(O_2)] = 0$ (as a graded commutator) within a larger algebra $\mathcal{A}(O)$ such that $O_1,O_2\subset O$.
        \end{enumerate}
    \end{axiom}
    \remark{The isotony condition implies that local nets of observables are modelled by copresheaves $\mathbf{Mink}\rightarrow\mathbf{C^*Alg}$ that map (mono)morphisms to monomorphisms.}

    \begin{axiom}[Poincar\'e covariance]
        For all causally closed sets $O$ and Poincar\'e transformations $\Lambda$ there exists an isomorphism $\alpha^O_\Lambda:\mathcal{A}(O)\rightarrow\mathcal{A}(\Lambda O)$ such that the following conditions are satisfied:
        \begin{enumerate}
            \item If $O_1\subset O_2$, then $\alpha_\Lambda\circ\iota_{O_1,O_2} = \iota_{\Lambda O_1,\Lambda O_2}\circ\alpha_\Lambda$.
            \item The isomorphisms satisfy a composition rule: $\alpha^{\Lambda O}_{\Lambda'}\circ\alpha^O_\Lambda = \alpha^O_{\Lambda'\Lambda}$.
        \end{enumerate}
    \end{axiom}

    \begin{axiom}[Spectrum]
        For all spacetime regions $O$ one can construct a faithful $C^*$-algebra representation $\rho_O$ of $\mathcal{A}(O)$ on a fixed Hilbert space by the GNS construction \ref{operators:gns}. The different representations should be compatible, i.e. if $O_1\subset O_2$, the restriction of $\rho_{O_2}$ to $\mathcal{A}(O_1)$ should equal $\rho_{O_1}$. Furthermore, all spacetime translations are implemented unitarily:
        \begin{gather}
            U(a)\rho_O(c)U(a)^{-1} = \rho_{O+a}(\alpha^O_a(c))
        \end{gather}
        for all $c\in\mathcal{A}(O)$, where $U$ is a unitary representation of the translation subgroup. In addition the generators of the translation subgroup are required to have a spectrum that is contained in the future light cone.
    \end{axiom}

    The following axiom is not part of the standard Haag-Kastler framework but can be added to introduce dynamics:
    \begin{axiom}[Time slice]
        Consider two spacetime regions $O_1,O_2$. If $O_1$ contains a Cauchy surface of $O_2$, the morphism $\mathcal{A}(O_1\hookrightarrow O_2)$ of $C^*$-algebras is an isomorphism.
    \end{axiom}

    \begin{axiom}[Haag duality]\index{Haag!duality}
        Let $\overline{O}$ denote the spacelike complement of $O$ and let $\mathcal{A}'$ denote the commutant of $\mathcal{A}$. Haag duality states that\footnote{Here it should be understood that $\mathcal{A}\left(\overline{O}\right)$ is the algebra generated by all algebras $\mathcal{A}(Q)$, where $Q$ ranges over the causally closed sets in $\overline{O}$.}
        \begin{gather}
            \mathcal{A}\left(\overline{O}\right)' = \mathcal{A}(O)
        \end{gather}
        for all causally closed sets $O$.
    \end{axiom}
    \remark{Haag duality is known to hold for all free theories and even for some interacting theories. However, it is also known to fail in the case of symmetry breaking \cite{Roberts_haag}.}

    To generalize the above axiom system to globally hyperbolic space times, one must enter the realm of category theory. The notation\footnote{This could potentially cause confusion with other notations used in this text. $\mathbf{Loc}$ here has nothing to do with the category of locales from Chapter \ref{chapter:topology}.} of \cite{cal_strobl} (?? AND OTHERS ??) will be adopted. Let $\mathbf{Loc}$ be the category of globally hyperbolic space times with orientation- and causal structure-preserving isometries. Let $\mathbf{Obs}$ be the category of relevant algebras (commutative algebras for classical physics and $C^*$-algebras for quantum theories) together with suitable algebra morphisms. The assignment of algebras is then given by a functor $\func{\mathfrak{U}}{Loc}{Obs}$. The Haag-Kastler framework is recovered when $\mathfrak{U}$ is restricted to the subcategory on globally hyperbolic subsets of some manifold (with inclusions as morphisms).

\subsection{Weyl systems}

    \newdef{Weyl system}{\index{Weyl!system}
        Let $(L,\omega)$ be a symplectic vector space and let $K$ be a complex vector space. Consider a map $W$ from $L$ to the space of unitary operators on $K$. The pair $(K,W)$ is called a Weyl system over $(L,\omega)$ if it satisfies
        \begin{gather}
            W(z)W(z') = e^{\frac{i}{2}\omega(z,z')}W(z+z')
        \end{gather}
        for all $z,z'\in L$.
    }
    \remark{This is a generalization of the Weyl form of the canonical commutation relations \ref{qm_formalism:CCR}.}

    \newdef{Heisenberg system}{\index{Heisenberg!system}
        The generators $\phi(z)$ of the maps $t\mapsto W(tz)$, which exist by Stone's theorem \ref{functional:stone}, are said to form a Heisenberg system. These operators satisfy the following properties:
        \begin{itemize}
            \item $\lambda\phi(z) = \phi(\lambda z)$ for all $\lambda>0$,
            \item $[\phi(z),\phi(z')] = -i\omega(z,z')$, and
            \item $\phi(z+z')$ is the closure \ref{functional:closure} of $\phi(z)+\phi(z')$.
        \end{itemize}
    }

\section{Topological QFT}

    For convenience the reader is reminded of some of the notations that are used. $\mathbf{FinVect}$ will denote the category of finite-dimensional vector spaces over $\mathbb{C}$ and $\mathbf{Bord}^d_{d-1}$ denotes the category of $d$-dimensional (framed\footnote{A framing in this context means a framing of the $d$-stabilized tangent bundles $TM\oplus\mathbb{R}^{d-\dim(M)}$.}) cobordisms \ref{manifold:cobordism}.

\subsection{Atiyah-Segal axioms}

    \begin{axiom}[Atiyah-Segal]\index{Atiyah-Segal axioms}\index{field!theory}
        \nomenclature[A_TQFT]{TQFT}{topological quantum field theory}
        A $d$-dimensional \textbf{topological quantum field theory} (TQFT) is a symmetric monoidal functor $F:\mathbf{Bord}_{d-1}^d\rightarrow\mathbf{FinVect}$. This means (among other things) that $F$ is a map satisfying the following axioms:
        \begin{enumerate}
            \item\textbf{Normalization}: $F(\emptyset)=\mathbb{C}$.
            \item\textbf{Disjoint union}: $F(M\sqcup M') = F(M)\otimes F(M')$.
            \item\textbf{Composition}: If $N=M\cup M'$, where $\partial M$ and $\partial M'$ have opposite orientations, then \[F(N) = F(M)\circ F(M').\]
            \item\textbf{Invariance}: If $f: M\rightarrow M'$ is a diffeomorphism rel boundary, then $F\circ f = F$.
        \end{enumerate}
        In the above conditions $M,M'$ are $d$-dimensional cobordisms between $(d-1)$-dimensional (closed) smooth manifolds.
    \end{axiom}

    \begin{example}[1D]
        In $d=1$ a TQFT functor gives rise to the following correspondence:
        \begin{gather*}
            \begin{array}{l|l}
                \text{point with orientation } + & \text{vector space } V\\
                \text{point with reversed orientation } - & \text{dual space }V^*\\
                \text{line between points} & \text{linear map }f:V\rightarrow V\\
                \text{cap between $\emptyset$ and points } +,- & \text{coevaluation } \mathbb{C}\rightarrow V\otimes V^*\\
                \text{cup between points $-,+$ and }\emptyset & \text{evaluation }V^*\otimes V\rightarrow\mathbb{C}
            \end{array}
        \end{gather*}
        Essentially this gives the structure of a (finite-dimensional) vector space and its dual.
    \end{example}

    \begin{example}[2D]
        In $d=2$ one can obtain a similar result by drawing all possible configurations. However, the existence and combination of ``pair of pants''-diagrams gives a richer structure. For 2D TQFTs the corresponding object is a (finite-dimensional) commutative and cocommutative Frobenius algebra \ref{linalgebra:frobenius}.
    \end{example}

    In dimensions 3 and higher the definition above is intractable. To allow the construction to be generalized to higher dimensions one considers the following (extended) definition:
    \newdef{Extended TQFT}{
        A $d$-dimensional extended TQFT is a symmetric monoidal $(\infty,d)$-functor $F:\mathbf{Bord}^d\rightarrow\mathbf{FinVect}$ satisfying the Atiyah-Segal axioms, where the invariance axiom is required only at the highest level of $k$-morphisms.
    }

    \begin{remark}
        One can replace the category $\mathbf{FinVect}$ by any other symmertric monoidal category $\mathbf{C}$.
    \end{remark}

    The following conjecture by \textit{Baez} and \textit{Dolan}, later proven by \textit{Lurie}, states that every extended TQFT is defined by a single object. This is motivated by the fact that all cobordisms between smooth manifolds can be obtained from gluing together (the geometric representation) of identity morphisms and taking disjoint unions of them.
    \begin{theorem}[Cobordism hypothesis]\index{cobordism!hypothesis}
        $\mathbf{Bord}^d$ is the free symmetric monoidal $(\infty,n)$-category with all duals on a single object $\mathrm{pt}$. In particular, every symmetric monoidal $(\infty,n)$-functor $F:\mathbf{Bord}^d\rightarrow\mathbf{C}$ is characterized by a single fully dualizable object $F(\mathrm{pt})\in\ob{C}$.
    \end{theorem}

\subsection{Open-closed TQFT}

    In the case of ordinary TQFTs as defined in the previous section, one considers cobordisms between closed manifolds. Hence, the relevant objects in this category are manifolds with boundary. A generalization is obtained by relaxing the constraint on the cobordisms and allowing for the notion of manifolds with corners (Section \ref{section:manifold_boundary}). For simplicity only the case of 2D TQFTs (as in the original definition) will be considered.

    ?? COMPLETE ??

\section{Prequantum field theory}

    Now, consider an (extended) TQFT. In full generality such a TQFT assigns to every manifold $M$ an ($\infty$-)stack of fields. For a cobordism, this induces a span \ref{cat:span} by restriction to the boundary components. Composition of cobordisms corresponds to pullback of spans. So, a TQFT induces a (monoidal) functor $\mathbf{Bord}^d\rightarrow\mathbf{Span}(\mathbf{H})$, where $\mathbf{H}$ is the $(\infty,1)$-category of $\infty\mathbf{Grpd}$-valued $\infty$-sheaves. When considering cobordisms between cobordisms, one obtains spans between spans. So, eventually, a $d$-dimensional TQFT can be seen to be equivalent to a symmetric monoidal $(\infty,n)$-functor $\mathbf{Bord}^d\rightarrow\mathbf{Corr}_d(\mathbf{H})$, where $\mathbf{Corr}_n(\mathbf{H})$ is the $(\infty,n)$-category of \textit{$n$-fold correspondences} in $\mathbf{H}$. By a theorem by \textit{Lurie}, correspondences admit a (symmetric) monoidal structure induced by that on $\mathbf{H}$. The cobordism hypothesis now immediately implies that any such functor is uniquely characterized by a (moduli) stack of fields. These theories are called \textbf{(local) prequantum field theories (pQFT)}.

    \begin{property}
        Any local pQFT is a (topological) $\sigma$-model:
        \begin{gather}
            \mathrm{Fields}(M)\cong[\smallint\!M,\mathbf{Fields}],
        \end{gather}
        where $\mathbf{Fields}$ is the classifying stack and $\smallint$ sends a manifold to its homotopy type, i.e. it is the shape modality \ref{topos:cohesive_modalities}.
    \end{property}

    Now, to obtain a reasonable quantum theory, one also needs an action principle. To a $d$-dimensional manifold it should assign a phase, i.e. a map $\mathbf{H}\rightarrow\mathrm{U}(1)$, to its boundary components it should assign a gauge field, i.e. a map $\mathbf{H}\rightarrow\mathbf{B}\mathrm{U}(1)$, etc. Combining this with the pQFT, one obtains a functor $\mathbf{Bord}^d\rightarrow\mathbf{Corr}_d(\mathbf{H}/_{\mathbf{B}^d\mathrm{U}(1)})$ to the $d$-fold correspondences in the slice topos over the classifying stack of circle $d$-bundles. Because the classiyfing stacks are \textit{$E_\infty$-algebras} in $\mathbf{H}$, i.e. commutative monoids in $\mathbf{H}$, the slice topos admits the structure of a symmetric monoidal category. Therefore one can extend the action functional to disjoint unions of cobordisms.
    \newdef{Action functional}{\index{action}
        An action functional for a pQFT $\mathbf{Bord}^d\rightarrow\mathbf{Corr}_d(\mathbf{H})$ is a lift (of monoidal functors) to a symmetric monoidal $(\infty,n)$-functor $\mathbf{Bord}^d\rightarrow\mathbf{Corr}_d(\mathbf{H}/_{\mathbf{B}^d\mathrm{U}(1)})$.
    }
    By the cobordism hypothesis, such a functional is uniquely defined by a morphism $\mathbf{Fields}\rightarrow\mathbf{B}^d\mathrm{U}(1)$, i.e. by a circle $d$-bundle on the space of fields.

    One can also rephrase this assignment in terms of (higher) holonomy functors. Any map $U\rightarrow[\smallint\!M,\mathbf{B}^d\mathrm{U}(1)]$ is given by th assignment of a flat $d$-bundle on $U\times M$, by the adjunction $\smallint\dashv\flat$. Taking the holonomy of this bundle defines a functor $[\smallint\!M,\mathbf{B}^d\mathrm{U}(1)]\rightarrow\mathbf{B}^{d-k}\mathrm{U}(1)$, where $k$ is the dimension of $U$. Composition with the classifying functor $\mathbf{Fields}\rightarrow\mathbf{B}^d\mathrm{U}(1)$ induces an assignment of $(d-k)$-bundles to the spaces of fields on $M$. This is the generalization of the (Lagrangian) action principle, where integration of Lagrangians is replaced by assigning higher holonomies.